### 2022-01-28 article/article.tex
initial commit

% \loop\iftrue%
\errmessage{This article is copyrighted by ___ and should not be TeXed}%
\repeat%

\documentclass{ifacconf}  % Comment this line out
\usepackage{natbib}        % required for bibliography

% \documentclass[a4paper, 10 pt, conference]{ieeeconf}  % Comment this line out
% \IEEEoverridecommandlockouts%
\pdfminorversion=4%

\input{preamble}
\input{math}
\input{symbols}

% \usepackage{flushend}% to equalize last page %%% use when text complete

\draft% show todos in red
% \final% give error if there is todos
% \comments% blank right column to make comments in pen & paper

% \newcommand{\id}{}
% \newcommand{\price}{31.00}
% \pubid{\begin{minipage}{\textwidth}\ \\[50pt]\centering
%   \copyright{} 20XX \_\_\_\_.\\
%   Personal use is permitted, but republication/redistribution requires permission.\\
%   % See https://www.ieee.org/publications/rights/index.html for more information.% if IEEE
% \end{minipage}
% }
% \makeglossaries

\begin{document}

\begin{frontmatter}
\title{\LARGE \bf
  Expectation-Maximization based defense mechanism for distributed Model Predictive Control
}

\author[First]{Rafael Accácio Nogueira}
\author[First]{Romain Bourdais}
\author[First]{Hervé Guéguen}
\address[First]{IETR-CentraleSupélec, 35510 Cesson-Sévigné, Ille-et-Vilaine, France\\
{\tt\small \{rafael-accacio.nogueira, romain.bourdais, herve.gueguen\}
@centralesupelec.fr}}

% \thanks{\centering The authors are with  \newline {\tt\small
% }\newline

% % <-this % stops a space
% \thanks{\centering The authors are with IETR-CentraleSupélec, 35510\newline Cesson-Sévigné, Ille-et-Vilaine, France \newline {\tt\small
% \{rafael-accacio.nogueira, romain.bourdais, herve.gueguen\}
% @centralesupelec.fr}\newline
% }%
% }

% \maketitle
% \IEEEpeerreviewmaketitle%
% \thispagestyle{empty}
% \pagestyle{empty}

\begin{abstract}% Abstract of not more than 250 words.
\end{abstract}

\begin{keyword}
  Model predictive and optimization-based control,
  Distributed control and estimation,
Large scale optimization problems
\end{keyword}

\end{frontmatter}

\section{INTRODUCTION}
% % % \mpc{} \dmpc{}

% % ~\\

\emph{Notation:} In this paper, $\norm{\cdot}$ and $\norm{\cdot}_{F}$ represent the $\ell_{2}$ and Frobenius norms. $\norm{\vec{v}}_{Y}$ is the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto set $\set{T}$.
$\card{x}$ is the number of elements in $x$.
$\kron{}{}$ represents the Kronecker product.
$n\mathbin{:}i\mathbin{:}j$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where $m=\mathrm{truncate}(\frac{j-n}{i})$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
$0_{m\times n}$ and $1_{m\times n}$ are ${m\times n}$ matrices filled with $0$ and $1$.
$\0$ and $\1$ are 0 and 1 filled vectors of adequate size.
$\preceq$ is the component-wise less or equal for vectors and matrix inequality for matrices in ${\symmetric=\setbuild{X:\R^{n\times n}}{X'=X}}$.
${\defpos=\setbuild{X:\symmetric}{X\succ 0}}$, ${\semidefpos=\setbuild{X:\symmetric}{X\succeq 0}}$.
$I_{c}$ is the identity matrix of size $c$.
$A^{\dagger}$ is the generalized inverse ${{(A^{T}A)}^{-1}A^{T}}$.
$|A|$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
$\random{a},\random{A},\randomvec{a}$ are random scalar, matrix and vector.
$\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

\section{PRELIMINARIES AND PROBLEM STATEMENT}\label{sec:PS}
\begin{equation}
\begin{matrix}
  \label{eq:systems}
\vec{x}_{i}[k+1]=A_{i}\xik + B_{i}\uik
\end{matrix}
\end{equation}

\begin{equation}
  \label{eq:upositive}
  \uik\succeq\0
\end{equation}

\begin{equation}
  \label{eq:constraint}
  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax
\end{equation}
where ${\Gamma_{i}:\R^{n_{u}\times n_{u}}}$, ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.

\begin{problem}{Global \mpc{} Problem.}\label{Pb:GOP}
\begin{equation*}
\begin{matrix}

\underset{\uiseq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:upositive}\ \mathrm{and}\ \eqref{eq:constraint}
\left\}\small
\begin{aligned}
  &\forall i\in \{1:\nsubsystems\}\\
  &\forall j\in \{1:\predhorz\}
\end{aligned}\right.

\end{matrix}
% \label{eq:dmpcModLOP}
\end{equation*}
where ${Q_{i}:\semidefpos}$ and ${R_{i}:\defpos}$, $\vik$ is a control objective.
Reference tracking can be written as ${\vik=\wik-\xik}$, where $\wik$ is a state reference.
Disturbance rejection can be written as ${\vik=\xik}$.

$\optglobobj$ denotes the optimal value of the problem~\ref{Pb:GOP}, and ${\optuiseq}$ is the optimal control sequences.
the problem~\ref{Pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
\end{problem}

\subsection{Distributed Model Predictive Control}\label{ssec:dMPC}
\todo{\dmpc{},
}
\begin{subequations}
  \begin{equation}
    \left.
      \small
      \begin{aligned}
        J_{i}^{\star}(\thetaik)&=\underset{\uiseq}{\mathrm{minimize}} \obji\\
        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:upositive}\\
        &\quad\Gamma_{i}\uik\preceq\thetaik:\lambdaik\\
      \end{aligned}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in \{1:\nsubsystems\}\\
      &\forall j\in \{1:\predhorz\}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}
  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}&=\underset{\thetaiseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\thetaik)\\
      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\thetaik\preceq\umax
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}

\begin{equation}
  \label{eq:projectedSubgradient}
\vec{\theta}^{(p+1)}=\Proj^{\set{H}}(\vec{\theta}\p-\rho\p\vec{g}\p)
\end{equation}

\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$. Where ${c=\pi_{\vec{u}_{i}(k:k+N_{p}-1|k)}=N_{p}\pi_{\uik}}$.}

\begin{equation}
  \label{eq:thetaNegot}
\thetai\pplusone=\thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
\end{equation}
\todo[verify if need to use max]{
\begin{equation}
  % \label{eq:thetaNegot}
  \thetai\pplusone=
  \left\{ \thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
  ,0\right\}
\end{equation}
}
where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$.
\todo{Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$}

\todo{
  The coordinator must also assure that ${\thetaik\succeq\0}$ so the local problems are feasible.
  But observe in~\eqref{eq:projectedSubgradient} that if ${\vec{\theta}_{i}\p=\0}$, ${(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p))}$ needs to be negative so $\vec{\theta}_{i}\pplusone$ becomes negative, which is impossible \question{is it though?} since $\lambdaik\succeq\0$, from the KKT optimality conditions, seen in~\cite[Chapter 5]{BoydVandenberghe2004}.
}

\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  $p:=0$\;
  Coordinator initializes $\vec{\theta}\p$ \;
  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
  $p:=p+1$
}
 \caption{Quantity decomposition based \acrlong{dmpc}.}\label{alg:quantityAlg}
\end{algorithm2e}

\section{Attack in DMPC scheme}\label{sec:attack}

\begin{equation}\label{eq:cheating}
\lambdaicheat=\gamma_{i}(\lambdai)
\end{equation}

\subsection{Quadratic Case --- Formal Analysis}\label{ssec:FA}
% \todo{As seen in \cite{NogueiraEtAl2021}},

\begin{equation}
\begin{matrix}
\underset{\Uik}{\mathrm{minimize}}& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
\mathrm{s.t.}&\Theta_{i}\Uik\preceq\vec{\theta}_{i}:\vec{\lambda}_{i}\\
&\Uik\succeq\0
\end{matrix}
\label{eq:quadratic_case}
\end{equation}

If we take reference tracking, for instance, we have:
\begin{equation}
\small\begin{matrix*}[l]
 H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik)\\

\end{matrix*}
\label{eq:matrices}
\end{equation}

We stack in vectors $\Uik$ and $\Wik$
the input and setpoint predictions sequences ${\uiseq}$ and ${\wiseq}$.
${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
$\bar{Q}_{i}$, $\bar{R}_{i}$, and $\Theta_{i}$ are block diagonal matrices built repeating $\predhorz$ times $Q_{i}$, $R_{i}$, and $\Gamma_i$ respectively.
\todo{
  Observe that $H_{i}:\defpos$.
}

\todo{As seen in~\cite{BemporadEtAl2002,BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
Similarly, we can get an explicit  solution for its dual variables $\vec{\lambda}_{i}$, which is also a \pwa{} function with respect to $\vec{\theta}_{i}$:}
\begin{equation}
  \begin{aligned}
    \label{eq:lambdafuntheta}
    \lambdai=
    \begin{cases}
      -P_{i}^{1}\vec{\theta}_{i}-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}\thetai \preceq z_{1} \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -P_{i}^{N}\vec{\theta}_{i}-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}\thetai \preceq z_{N} \\
    \end{cases}
  \end{aligned}
\end{equation}
The halfspaces defined by the set of pairs $(G_{j},z_{j})$ represent a combination of active constraints.
As each constraint can be active or inactive, for a group of $n$ constraints, \todo{potentially} we can have $N=2^{n}$ different combinations of active/inactive sets, consequently $N$ regions (or zones).
\todo{
  Here we say potentially because depending on the constraints, some of these combinations can be infeasible.\question{Remove??}
}

\todo{
  The values of $P_{i}^{j}$ and $\vec{s}_{i}^{j}$, depend on $H_{i}$, $\fik$ and parts of $\Theta_{i}$ which represents the active constraints.
  If we take for example when all constraints are active, we have ${P_{i}^{1}={(\Theta_{i}H_{i}^{-1}\Theta_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{1}[k]=P_{i}^{1}\Theta_{i}H_{i}^{-1}\fik}$.
  And if all constraints are inactive, ${P_{i}^{N}=0_{\card{\Theta_{i}}\times\card{\Theta_{i}}}}$ and ${\vec{s}_{i}^{N}=\0}$.
  Observe that in all functions, only $\vec{s}_{i}^{j}[k]$ depend on time $k$.
}

\subsection{Detection and mitigation}
\begin{assumption}
  $\gamma_{i}(\cdot)$ is the same during the negotiation phase for a given time $k$ (it does not depend on $p$).
\end{assumption}

\begin{assumption}
  We suppose the agent chooses a linear function such as $\tilde{\lambdai}=\gamma_{i}(\vec{x})=\Tik\vec{x}$ resulting in:
  \begin{equation}
    \label{eq:linear_cheating}
    \tilde{\lambdai}=
        \begin{cases}
      -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}\thetai \preceq z_{1} \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}\thetai \preceq z_{N} \\
    \end{cases}
  \end{equation}
and we define $\tilde{P}_{i}^{j}[k]=\Tik P_{i}^{j}$ and $\tilde{\vec{s}}_{i}^{j}[k]=\Tik\vec{s}_{i}^{j}[k]$.
\end{assumption}
\todo{
  Since $\Theta_{i}$ are diagonal we can assure\pdfnote{proof?} there exists one \question{zone?}zone where all constraints are active. We will fix $j=1$ for this zone and called it the $1$-zones.
  We can estimate the values for this partition using:
}
\begin{equation}
  \label{eq:lambdafuntheta_tilde}
\tilde{\lambdai}=\gamma_{i}(\lambdai(\thetai))=-\widehat{\tilde{P}_{i}^{1}}[k]\vec{\theta}_{i}-\widehat{\tilde{\vec{s}}_{i}^{1}}[k]
\end{equation}

\begin{assumption}\label{ass:Pnominal}
  The nominal value of $P_{i}^{1}$, denoted $\bar{P}_{i}^{1}$, is available from reliable attack-free historical data.
\end{assumption}

\todo{
  Using this strategy, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{1}}[k]-\bar{P}_{i}^{1}\|_{F}}$, where ${\|\cdot\|_{F}}$ is the Frobenius norm.
  Let ${d_{i}\in\{0,1\}}$ be an indicator that detects the attack in agent $i$.
  If the disturbance $E_{i}[k]$ respects an arbitrary bound
}
\begin{equation}
  \label{eq:2}
  E_{i}[k]\leq\epsilon_{P},
\end{equation}
\todo{
  then  ${d_{i}=0}$, and no attack is detected. Otherwise, ${d_{i}=1}$, and a change in behavior of agent $i$ is detected.
}

\todo{
  Using these assumptions, we can try to estimate the inverse of $T_{i}(k)$ as in
}
\begin{equation}\label{eq:5}
\widehat{{T_{i}(k)}^{-1}}=\bar{P}_{i}^{1}{\widehat{\tilde{P}_{i}^{1}}[k]}^{-1},
\end{equation}

and from~\eqref{eq:lambdafuntheta}, we can derive a method to reconstruct $\vec{\lambda}_{i}$:
\begin{equation}
  \label{eq:lambdareconstruction}
  {\vec{\lambda}_{i}}_{\mathrm{rec}}=\widehat{{T_{i}[k]}^{-1}} \tilde{\vec{\lambda}_{i}}
\end{equation}
\subsection{Estimating}
\todo{
  In order to estimate $\widehat{\tilde{P}_{i}^{1}}$, we must have enough observed points in the $1$-zones. We propose to generate points surrounding arbitrary $\bar{\vec{\theta}_{i}}$ in the $1$-zones, and use them to estimate $\widehat{\tilde{P}_{i}^{1}}$. Now we have to find such points.
}

% TODO(accacio): change unconstrained notation
\todo{
  Let $\optuncUik$ be the solution of the unconstrained version of~\ref{eq:quadratic_case}.
  It is known that its solution is ${ {\optuncUik}=-H_{i}^{-1}\fik}$.
  For the constrained solution we know that the constraints will be inactive if they respect the inequalities $\Theta_{i}\optuncUik\preceq\thetai$.
  If they regard the inequalities, the constrained value is calculated as is, if not it should be projected onto the viable region.
  This way, for all constraints to be active, all inequalities should be disrespected in $\Theta_{i}\optuncUik\preceq\thetai$, thus \[\thetai\prec\Theta_{i}\optuncUik.\]
  % Since $H_{i}:\defpos$, then $H_{i}^{-1}:\defpos$,
  If we choose $\bar{\theta}_{i}=\0$, the constraints will be active when the  will only  that can assure that all constraints are active, then we can estimate ${\widehat{\tilde{P}_{i}^{1}}[k]}$ around this point. Unfortunately, since we don't know the hyperplanes that separates the different zones, some points generated for the estimation can belong to other zones.
  So we use the \EM{} algorithm to estimate the parameter for the $1$-zones, as it can clusterize the points and estimate parameters at the same time.
}

\section{Expectation-Maximization Algorithm}\label{sec:expect-maxim-algo}

\todo{
  The main objective of the \EM{} algorithm is to maximize the complete-data likelihood for models with latent variables.
}
\todo{
  The \EM{} problem can be described as such:
  \begin{problem}{Expectation Maximization Problem}\label{pb:EM}

    Given a set of observed data $\random{X}$, and unknown latent variables $\random{Z}$ and parameters $\random{P}$ that generated these observed data, find estimates of $\random{Z}$ and $\random{P}$ that maximize the complete-data log likelihood
    \begin{equation}\label{eq:completedataLogLikelihood}
    \ln\probability{\random{X},\random{Z}|\random{P}} .
  \end{equation}
  Since nor $\random{Z}$ nor $\random{P}$ are given, we cannot calculate~\eqref{eq:completedataLogLikelihood}, instead we calculate the expectation of this value for a given estimate $\random{P}^{\mathrm{cur}}$:
  \begin{equation}
    \label{eq:expectedcompletedataloglikelihood}
Q(\random{P},\random{P}^{\mathrm{cur}})=\sum_{\randomvec{z}}\probability{\random{Z}|\random{X},\random{P}^{\mathrm{cur}}}\ln\probability{\random{X},\random{Z}|\random{P}}
  \end{equation}

  Then we can find a new estimate of $\random{P}$ that maximizes $Q(\random{P},\random{P}^{\mathrm{cur}})$:
  \begin{equation} \label{eq:Mstep}
    \random{P}^{\mathrm{new}}=\underset{\mathrm{P}}{\argmax}\ Q(\random{P},\random{P}^{\mathrm{cur}}).
  \end{equation}

  But as seen in~\eqref{eq:expectedcompletedataloglikelihood}, we need first to calculate the posterior distribution
  \begin{equation}
    \label{eq:responsibilities}
    \zeta(\random{X},\random{P}^{\mathrm{cur}})=\probability{\random{Z}|\random{X},\random{P}^{\mathrm{cur}}},
\end{equation}
which is also called responsibilities.
With this information we can describe the \EM{} in algorithm~\ref{alg:em} adpated from~\mbox{\cite[Chapter 9]{Bishop2006}}:
  \end{problem}
}
\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  Initialize parameters $\random{P}^{\mathrm{new}}$\;
  \Repeat{$\random{P}^{\mathrm{cur}}$ converge}{
    $\random{P}^{\mathrm{cur}}=\random{P}^{\mathrm{new}}$\;
  \textbf{E step} Evaluate responsibilities~\eqref{eq:responsibilities}\;
  \textbf{M step} Reestimate parameters~\eqref{eq:Mstep}\;
}
 \caption{Expectation Maximization}\label{alg:em}
\end{algorithm2e}
\todo{
  We can see the similarities to K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
  Instead of affecting the observed data to a cluster with 100\% of certainty, \EM{} gives a responsibility (posterior probability of point belonging to cluster) for each cluster.
  Then, when we solve the \textbf{M Step} we take into account all the data, but weighted by these responsibilities.
}

\section{Using the algorithm to estimate $P_{i}^{1}[k]$}\label{sec:using-em-algorithm}
\todo{
  Here to simplify the notation for this section we concentrate on a single agent, and in a single step $k$, so we drop the subscript $i$ and the time dependency $[k]$.

  % First we need to define our model.
}
% TODO(accacio): change subscript n
\todo{
  For any component of $\tilde{\vec{\lambda}}$, given an input $\randomvec{\theta}_{o}$,  we can define a latent variable $\random{z}_{o}$ which determine to which zone ${z=\{1:Z\}}$ the $\randomvec{\theta}_{o}$ belongs and which parameters $P_{i}^{z}$ and $\vec{s}^{z}$ were used
to generate an observation $o$ of the $c^{\mathrm{th}}$ component of $\tilde{\lambda}$, $\random{\lambda}_{c,o}$.

}
\todo{
  We propose to transform~\eqref{eq:linear_cheating} into a random \pwa{} for each component $c$ of $\lambda$:
}
  \begin{equation}
    \label{eq:linear_cheating_random}
    \random{\lambda}_{c,o}=
        \begin{cases}
      -\tilde{P}_{c}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}_{c}}^{1}&\text{with probability}\ \pi_{1} \\
      \qquad\quad \vdots&\qquad\quad \vdots\\
      -\tilde{P}_{c}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}_{c}}^{Z}&\text{with probability}\ \pi_{Z} \\
    \end{cases}
  \end{equation}
  where $\tilde{P}_{c}^{j}$ and $\tilde{\vec{s}}_{c}^{j}$ are the $c^{\mathrm{th}}$ row of  $\tilde{P}^{j}$ and $\tilde{\vec{s}}^{j}$.

\todo{
  As a consequence of this formulation, we have  \[\probability{\randomvec{z}_{o}=z}=\pi_{z},\]
  and since $\vec{\theta}$ is our input, we know \emph{almost surely} their values: \[\probability{\randomvec{\theta}_{o}}=1.\]
}
\todo{
  With this data, we can use the following graph in Fig.~\ref{fig:model} to represent the model for the $O$ observations of generated $\random{\lambda}_{c}$, where $\phi_{c}=\{P_{c}^{z},\vec{s}_{c}^{z}\}$ for ${z=\{1:Z\}}$ are the parameters we want to estimate for each zone $z$ and component $c$.

}

\usetikzlibrary{arrows.meta}
\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
    \node at (1.2,-1.3) {O};
    \node (pi) at (-2.5,.5) {$\pi$};
    \node (phi) at (2.5,-1.) {$\phi_{c}$};

    \graph [edge quotes={fill=white,inner sep=1pt},
    clockwise=3,nodes={circle,draw,rotate=-60,minimum width=1cm}] {
      a/"${\randomvec{\theta}_{o}}$"[rotate=60],b/"${\random{\lambda}_{c,o}}$"[rotate=60],c/"${\random{z}}_{o}$"[rotate=60];
      {a,c} ->[-latex] b;
      (pi) ->[thick,{Circle[length=3.pt]}-latex] c;
      (phi) ->[thick,{Circle[length=3.pt]}-latex] b;
      };

  \end{tikzpicture}
  \caption{Graph representation of generation of $\random{\lambda}_{c,o}$.}\label{fig:model}
\end{figure}
\todo{
Similar as in~\cite{FariaSoromenho2010}, we can model ${\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=z,\phi_{c}}}$ as an univariate Gaussian distribution with mean \[f_{c}^{z}(\randomvec{\theta})=-\tilde{P_{c}^{z}}\randomvec{\theta}_{o}-\tilde{\vec{s}_{c}}^{z}\] and variance ${\sigma^{z}}^2\to 0$:
}
\begin{equation}\label{eq:1}
\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=z,\phi_{c}}=\mathcal{N}(\random{y};f_{c}^{z}(\randomvec{\theta}_{o}),{\sigma^{z}}^{2}).
\end{equation}

The responsibilities \({{\zeta(z)}_{c,o}=\probability{\random{z}_{o}=z\mid\random{\lambda}_{c,o},\randomvec{\theta}_{o},\phi_{c}}}\) can be calculated as
\[\zeta(z)_{c,o}=\frac{\pi_{z}\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=z,\phi_{c}}}{\sum\limits_{j=1}^{Z}\pi_{j}\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=j,\phi_{c}} }\]
Observe that by taking
\begin{equation}\label{eq:argmaxz}
\underset{z}{\argmax} {\zeta(z)}_{c,o},
\end{equation}
we can get the most probable $z$-zone which generated the observation $\random{\lambda}_{c,o}$.

Then, as in~\cite[Chapter 9]{Bishop2006}, the expected value of the complete-data log-likelihood can be computed by
\begin{equation}
 Q(\phi_{c},\phi_{c}^{\mathrm{cur}})=\expectation[\probability{\random{z}\mid \random{\lambda}_{c},\randomvec{\theta},\phi_{c}^{\mathrm{cur}}}]{\ln\probability{\random{\lambda}_{c},\randomvec{\theta},\random{z},\phi_{c}}},
\end{equation}
which can be developed to
\begin{equation}\label{eq:completedataLogLikelihoodestimation}
  % Q(\phi_{c},\phi_{c}^{\mathrm{cur}})=
  \small
  \sum_{z=1}^{Z}\sum_{o=1}^{O}\zeta(z)_{c,o}^{\mathrm{cur}}\left(\ln \pi_{z}-\frac{\ln{\sigma^{z}}^{2}}{2}-\frac{{{\left(\random{\lambda}_{c,o}-f_{c}^{z}(\randomvec{\theta}_{o})\right)}^{2}}}{2{\sigma^{z}}^{2}}\right)+\ \mathrm{cnt},
\end{equation}
where the constant term is independent of any parameter in $\phi_{c}$ and it is not computed.

Calculating the gradient of~\eqref{eq:completedataLogLikelihoodestimation} with respect to vectors $\vec{\phi}_{c,z}={[{P_{c}^{z}}\ \vec{s}_{c}^{z}]}^{T}$ and making it vanish, it is possible to calculate an optimal new estimate $\vec{\phi}_{c,z}^{\mathrm{new}}$ using:
\begin{equation}
  \label{eq:mstepestimation}
  \vec{\phi}_{c,z}^{\mathrm{new}}=\pseudoinv{(\Xi_{c,z}\randomvec{\theta})}\Xi_{c,z}\randomvec{\lambda}_{c},
\end{equation}
where
${\randomvec{\theta}={[ {[\randomvec{\theta}_{1}^{T}\ 1]}\ \dots\ [\randomvec{\theta}_{O}^{T}\ 1] ]}^{T}}$%
,
${\randomvec{\lambda}_{c}={[\random{\lambda}_{c,1}\ \dots\ \random{\lambda}_{c,O}]}^{T}}$%
and
${\Xi_{c,z}={\diag(\sqrt{{\zeta(z)}_{c,1}},\cdots,\sqrt{{\zeta(z)}_{c,O}})}}$.

As we can see,~\eqref{eq:mstepestimation} is a weighted Least-Squares, that uses the responsibilities as weights, ignoring those values closer to $0$.

Once all $\tilde{P}_{c}^{z}$ and $\tilde{\vec{s}}_{c}^{z}$ are estimated, we can stack them creating estimates of $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$.

Observe that

% \begin{equation*}
%   p(\randomvec{x})=\sum_{i=0}^{N}\pi_{i} \mathcal{N}(\random{y};f_{i}(\randomvec{x}),\Sigma_{i})
% \end{equation*}

% \begin{equation}
%   \mathcal{N}(\random{y}; {\mu}_{i},\Sigma_{i})=\frac{1}{{\sqrt{2\pi\sigma_{i}^{2}}}}e^{-\frac{(\random{y}-{\mu}_{i})^{2}}{2\sigma^{2}}}
% \end{equation}

% for a $D-$dimensional gaussian we have:
% \begin{equation}
%   \mathcal{N}(\randomvec{y}; \vec{\mu}_{i},\Sigma_{i})=\frac{1}{{{(2\pi)}^{D/2}}{|\Sigma_{i}|^{\frac{1}{2}}}}e^{-\frac{1}{2}\norm{\randomvec{y}-\vec{\mu}_{i}}^{2}_{\Sigma_{i}^{-1}}}
% \end{equation}

% \begin{figure}[b]
%   \centering
%   \scriptsize \def\svgwidth{0.49\textwidth}
%   \includegraphics[width=\columnwidth]{pwagaussian.pdf}
%   \caption{Example of Gaussian Probability density functions used in mixture for identification of a 2D Piecewise Affine function with 2 modes, where \todo[verify colormap]{black represents 0 probability and white maximum }}\label{fig:pwagaussian}
% \end{figure}
\todo{Simulated annealing as in \cite{OzerovFevotte2010}}

\section{Conclusion}

A conclusion section is not required. Although a conclusion may review
the main points of the paper, do not replicate the abstract as the
conclusion. A conclusion might elaborate on the importance of the work
or suggest applications and extensions.
\begin{ack}
Simon Leglaive
\end{ack}

% \nocite{*}
\bibliography{bibliography}
% \todo[Change bibliography to the one used in the aux]{}
\end{document}

### 2022-01-28 code/gmmpwa.m

@@ -1,0 +1,24 @@
+clear; close all ;
+subplot(1,2,1)
+x1 = linspace(0,5,100);
+x2 = linspace(-5,10,100);
+[X1 ,X2 ] = meshgrid(x1,x2);
+Z=(1000/sqrt(2*pi).*exp(-((X2-0*X1-0).^2/2)));;
+
+surf(X1,X2,Z,'LineStyle','none')
+colormap(gca,bone)
+set_image_style(18);
+view(0,90)
+
+subplot(1,2,2)
+x1 = linspace(0,5,100);
+x2 = linspace(-5,10,100);
+[X1 ,X2 ] = meshgrid(x1,x2);
+Z=(1000/sqrt(2*pi).*exp(-((X2-2*X1+4).^2/2)));;
+surf(X1,X2,Z,'LineStyle','none')
+colormap(gca,bone)
+sgtitle('Gaussian distributions used in mixture','FontSize',18,'interpreter','latex')
+view(0,90)
+set_image_style(18);
+print_image_pdf('../img/pwagaussian',[8 3.5])
+

### 2022-01-31 article/article.tex
correct first line
add graph for dmpc exchange

@@
-% \loop\iftrue%
+\loop\iftrue%
 \errmessage{This article is copyrighted by ___ and should not be TeXed}%
@@
 }
-where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$.
-\todo{Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$}
+\no{\todo[delete probably]{
+  But observe in~\eqref{eq:projectedSubgradient} that if ${\vec{\theta}_{i}\p=\0}$, ${(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p))}$ needs to be negative so $\vec{\theta}_{i}\pplusone$ becomes negative, which is impossible \question{is it though?} since $\lambdaik\succeq\0$, from the KKT optimality conditions, seen in~\cite[Chapter 5]{BoydVandenberghe2004}.
+}}
 
-\todo{
-  The coordinator must also assure that ${\thetaik\succeq\0}$ so the local problems are feasible.
-  But observe in~\eqref{eq:projectedSubgradient} that if ${\vec{\theta}_{i}\p=\0}$, ${(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p))}$ needs to be negative so $\vec{\theta}_{i}\pplusone$ becomes negative, which is impossible \question{is it though?} since $\lambdaik\succeq\0$, from the KKT optimality conditions, seen in~\cite[Chapter 5]{BoydVandenberghe2004}.
+% TODO(accacio): correct image
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[node distance=1.7cm]
+    \tikzset{
+      agent/.style={draw,circle,minimum width=1cm,},
+      coord/.style={draw,circle,minimum width=1.7cm,very thick}
     }
+    % \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
+    % \node at (1.2,-1.3) {O};
+    % \node (pi) at (-2.5,.5) {$\pi$};
+    % \node (phi) at (2.5,-1.) {$\phi_{c}$};
+
+    \node[agent] (a1) {A$_{1}$};
+    \node (a2) [right of=a1]  {$\dots$};
+    \node[agent] (am) [right of=a2] {A$_{M}$};
+    \node[coord] (c) [above of=a2] {Coord};
+
+    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}$} (c) ;
+    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}$} (a1);
+
+    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}$} (c);
+    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}$} (am);
+
+  \end{tikzpicture}
+  \caption{Graph representation of Coordinator exchangin with $M$ agents in a \dmpc{} framework.}\label{fig:dmpc_graph}
+\end{figure}
 

### 2022-01-31 article/article.tex
add challenge and time dependency

@@
     \begin{cases}
-      -P_{i}^{1}\vec{\theta}_{i}-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}\thetai \preceq z_{1} \\
+      -P_{i}^{1}\vec{\theta}_{i}-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
       \qquad\quad \vdots&\qquad\quad \vdots\\
-      -P_{i}^{N}\vec{\theta}_{i}-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}\thetai \preceq z_{N} \\
+      -P_{i}^{N}\vec{\theta}_{i}-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
     \end{cases}
@@
 \end{equation}
-The halfspaces defined by the set of pairs $(G_{j},z_{j})$ represent a combination of active constraints.
+The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active constraints for a given time $k$.
+
+\begin{challenge}
+It is important to note that since the actual state is unknown we cannot anticipate how the space is partitioned, nor the values of $\vec{s}_{i}[k]$.
+\end{challenge}
  As each constraint can be active or inactive, for a group of $n$ constraints, \todo{potentially} we can have $N=2^{n}$ different combinations of active/inactive sets, consequently $N$ regions (or zones).

### 2022-02-02 article/article.tex
add Simon

@@
 \author[First]{Romain Bourdais}
+\author[First]{Simon Leglaive}
 \author[First]{Hervé Guéguen}
 \address[First]{IETR-CentraleSupélec, 35510 Cesson-Sévigné, Ille-et-Vilaine, France\\
-{\tt\small \{rafael-accacio.nogueira, romain.bourdais, herve.gueguen\}
+{\tt\small \{rafael-accacio.nogueira, romain.bourdais, simon.leglaive, herve.gueguen\}
 @centralesupelec.fr}}

### 2022-02-03 article/article.tex
add first draft of section about EM

@@
 $\kron{}{}$ represents the Kronecker product.
+$\vectorize{A}$ vectorizes matrix $A$.
+$\hadamard{}{}$ represents the Hadamard product.
 $n\mathbin{:}i\mathbin{:}j$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where $m=\mathrm{truncate}(\frac{j-n}{i})$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
+$[A,B]$ and $[A;B]$ are horizontal and vertical concatenation of matrices of adequate sizes.
 $0_{m\times n}$ and $1_{m\times n}$ are ${m\times n}$ matrices filled with $0$ and $1$.
 $\0$ and $\1$ are 0 and 1 filled vectors of adequate size.
+$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
 $\preceq$ is the component-wise less or equal for vectors and matrix inequality for matrices in ${\symmetric=\setbuild{X:\R^{n\times n}}{X'=X}}$.
@@
 
-\todo{
-  The main objective of the \EM{} algorithm is to maximize the complete-data likelihood for models with latent variables.
-}
-\todo{
-  The \EM{} problem can be described as such:
+The main objective of the \EM{} algorithm is to find maximum likelihood estimators for models with latent variables.
+
+We will use our estimation problem to illustrate how the algorithm works, and
+we will concentrate on the estimation of the parameters for a single agent in a single step $k$, so we drop the subscript $i$ and the time dependency $[k]$ to simplify the notation.
+
+As in~\eqref{eq:lambdafuntheta},  any response variable $\tilde{\vec{\lambda}}$, is a function of an input $\vec{\theta}$, which belongs to a unknown zone ${\set{Z}=\{1\mathbin{:}Z\}}$, Challenge~\ref{ch:zone_unknown}.
+The relationship between the input and the response is given by a set of parameters we want to estimate ${\set{P}=\setbuild{(\tilde{P}^{z},\tilde{\vec{s}}^{z})}{z\in\set{Z}}}$.
+
+For an observation $o$, we have the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$.
+As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a expansion of the model referred as \emph{mixture of switching regressions} in~\cite{QuandtRamsey1978} and \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010}, which we will call \emph{mixture of switching affine regressions}, since our regressors have a linear term (matrices $\tilde{P}^{z}$) and a constant term (vectors $\tilde{\vec{s}}^{z}$):
+% TODO(accacio): change subscript n
+  \begin{equation}
+    \label{eq:linear_cheating_random}
+    \random{\lambda}_{o}=
+        \begin{cases}
+      -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{with probability}\ \pi_{1} \\
+      \qquad\quad \vdots&\qquad\quad \vdots\\
+      -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{with probability}\ \pi_{Z} \\
+    \end{cases}
+  \end{equation}
+
+As we can see, this observation does not gives us the complete data, since we don't know which zone generated the response, but only the probability of the input being in that zone.
+We consider the complete data consisting of the input, the response, and a non-observable variable.
+Therefore, for each observation $o$ there exists a latent random variable ${\random{z}_{o}\in\set{Z}}$, with associated probabilities ${\Pi=\{\pi_{1},\dots,\pi_{Z}\}}$ such prior probability
+   \begin{equation}
+     \label{eq:prob_zo_equal_z}
+\probability{\random{z}_{o}=z}=\pi_{z}.
+\end{equation}
+Since $\vec{\theta}$ is our input, we know \emph{almost surely} their values giving
+\begin{equation}
+  \label{eq:theta_almost_surely}
+  \probability{\randomvec{\theta}_{o}}=1.
+\end{equation}
+With this information we can use the graph in Fig.~\ref{fig:model} to represent the model.
+\begin{figure}[b]
+  \centering
+  \begin{tikzpicture}
+    \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
+    \node at (1.2,-1.3) {O};
+    \node (pi) at (-2.5,.5) {$\Pi$};
+    \node (phi) at (2.5,-1.) {$\set{P}$};
+
+    \graph [edge quotes={fill=white,inner sep=1pt},
+    clockwise=3,nodes={circle,draw,rotate=-60,minimum width=1cm}] {
+      a/"${\randomvec{\theta}_{o}}$"[rotate=60],b/"${\random{\vec{\lambda}}_{o}}$"[rotate=60],c/"${\random{z}}_{o}$"[rotate=60];
+      {a,c} ->[-latex] b;
+      (pi) ->[thick,{Circle[length=3.pt]}-latex] c;
+      (phi) ->[thick,{Circle[length=3.pt]}-latex] b;
+      };
+  \end{tikzpicture}
+  \caption{Graph representation of model proposed.}\label{fig:model}
+\end{figure}
+
+As said, the objective of the \EM{} algorithm is to maximize the likelihood, more precisely the complete data log-likelihood $\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}$.
+Using the dependency graph in Fig.~\ref{fig:model} yields
+    \begin{equation}\label{eq:completedataLogLikelihood}
+      \ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}=\ln \prod_{o=1}^{O}\probability{\randomvec{\theta}_{o},\randomvec{\lambda}_{o},\random{z}_{o};\set{P}} .
+    \end{equation}
+where the probability $\probability{\randomvec{\theta}_{o},\randomvec{\lambda}_{o},\random{z}_{o};\set{P}}$ is
+\begin{equation}\label{eq:condional}
+ \prod_{z=1}^{Z}{[\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z;\set{P}^{z}}\probability{\random{z}_{o}=z}\probability{\randomvec{\theta}_{o}}]}^{\indicator{\random{z}_{o}=z}},
+\end{equation}
+and $\set{P}^{z}$ is the $z$-th tuple in $\set{P}$.
+
+For the likelihood $\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z,\set{P}^{z}}$, we use multivariate normal probability density functions
+\begin{equation}
+  \label{eq:multivariate_gaussian}
+\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}}),
+\end{equation}
+where the mean is
+\[f(\randomvec{\theta}_{o};(P,\vec{s}))=-{P}\randomvec{\theta}_{o}-{\vec{s}}\]
+and the covariance is ${\Sigma^{z}}$, which tends to $0$.
+
+Using~\eqref{eq:prob_zo_equal_z},~\eqref{eq:theta_almost_surely}, and~\eqref{eq:multivariate_gaussian} in~\eqref{eq:completedataLogLikelihood} yields
+\begin{equation}\label{eq:completedataLogLikelihood_complete}
+\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}=  \sum_{o=1}^{O}\sum_{z=1}^{Z}{\indicator{\random{z}_{o}=z}}
+  \alpha_{zo},
+\end{equation}
+where ${\alpha_{zo}=\ln{\pi_{z}}+\ln{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}$.
+
+Now we can formalize the \EM{} problem.
 \begin{problem}{Expectation Maximization Problem}\label{pb:EM}
 
-    Given a set of observed data $\random{X}$, and unknown latent variables $\random{Z}$ and parameters $\random{P}$ that generated these observed data, find estimates of $\random{Z}$ and $\random{P}$ that maximize the complete-data log likelihood
-    \begin{equation}\label{eq:completedataLogLikelihood}
-    \ln\probability{\random{X},\random{Z}|\random{P}} .
+  Given a set of observed data $(\random{\Theta},\random{\Lambda})$, estimate the unknown latent variables $\random{Z}$ and parameter set $\set{P}$ which maximize the complete-data log likelihood in~\eqref{eq:completedataLogLikelihood_complete}.
+\end{problem}
+
+But as mentioned, we do not observe the complete data $(\random{\Theta},\random{\Lambda},\random{Z})$, instead we observe only $(\random{\Theta},\random{\Lambda})$ and all information we can get from $\random{Z}$ is the posterior probability  ${\zeta(z_{zo};\set{P})=\probability{\random{z}_{o}=z|\randomvec{\lambda}_{o},\randomvec{\theta}_{o};\set{P}}}$, also called \emph{responsibility}, which can be calculated as
+\begin{align}
+  \label{eq:responsibilities}
+\zeta(z_{zo};\set{P})&=\frac{\probability{\random{z}_{o}=z}\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o};\set{P}^{z}}}{\sum\limits_{j=1}^{Z}\probability{\random{z}_{o}=j}\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o};\set{P}^{j}}}\nonumber\\
+  &=\frac{\pi_{z}{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}{\sum\limits_{j=1}^{Z}\pi_{j}\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{j}),{\Sigma^{j}})}.
+\end{align}
+
+Observe that by taking
+\begin{equation}\label{eq:argmaxz}
+\underset{z_{o}}{\argmax}\ {\zeta(z_{zo};\set{P})},
 \end{equation}
-  Since nor $\random{Z}$ nor $\random{P}$ are given, we cannot calculate~\eqref{eq:completedataLogLikelihood}, instead we calculate the expectation of this value for a given estimate $\random{P}^{\mathrm{cur}}$:
+we can get the most probable $z$-zone which generated the observation $\randomvec{\lambda}_{o}$.
+
+So, rather than using the complete data log-likelihood~\eqref{eq:completedataLogLikelihood_complete}, which we do not have, we use its expectation with respect to the prior calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$,  ${Q\left(\set{P},\set{P}_{\mathrm{cur}}\right)=\expectation[{\zeta(z_{zo};\set{P}_{\mathrm{cur}}}]{\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}}$:
 \begin{equation}
-    \label{eq:expectedcompletedataloglikelihood}
-Q(\random{P},\random{P}^{\mathrm{cur}})=\sum_{\randomvec{z}}\probability{\random{Z}|\random{X},\random{P}^{\mathrm{cur}}}\ln\probability{\random{X},\random{Z}|\random{P}}
+  \label{eq:completedataLogLikelihood_expectation}
+ Q\left(\set{P},\set{P}_{\mathrm{cur}}\right) =
+  \sum_{o=1}^{O}\sum_{z=1}^{Z}  \zeta(z_{zo};\set{P}_{\mathrm{cur}})\alpha_{zo}
+      ,
 \end{equation}
 
-  Then we can find a new estimate of $\random{P}$ that maximizes $Q(\random{P},\random{P}^{\mathrm{cur}})$:
+Then we can find a new estimate of $\set{P}$ that maximizes $Q(\set{P},\set{P}_{\mathrm{cur}})$:
 \begin{equation} \label{eq:Mstep}
-    \random{P}^{\mathrm{new}}=\underset{\mathrm{P}}{\argmax}\ Q(\random{P},\random{P}^{\mathrm{cur}}).
+  \set{P}_{\mathrm{new}}=\underset{\set{P}}{\argmax}\ Q(\set{P},\set{P}^{\mathrm{cur}}).
 \end{equation}
 
-  But as seen in~\eqref{eq:expectedcompletedataloglikelihood}, we need first to calculate the posterior distribution
-  \begin{equation}
-    \label{eq:responsibilities}
-    \zeta(\random{X},\random{P}^{\mathrm{cur}})=\probability{\random{Z}|\random{X},\random{P}^{\mathrm{cur}}},
-\end{equation}
-which is also called responsibilities.
-With this information we can describe the \EM{} in algorithm~\ref{alg:em} adpated from~\mbox{\cite[Chapter 9]{Bishop2006}}:
-  \end{problem}
-}
+With this information we can describe the \EM{} algorithm in two parts, first calculate the responsibilities $\zeta(z_{zo};\set{P}_{\mathrm{cur}})$, and then update the parameters.
+Algorithm~\ref{alg:em} adpated from~\mbox{\cite[Chapter 9]{Bishop2006}} describes the steps.
+
 \begin{algorithm2e}[h]
   \DontPrintSemicolon%
-  Initialize parameters $\random{P}^{\mathrm{new}}$\;
-  \Repeat{$\random{P}^{\mathrm{cur}}$ converge}{
-    $\random{P}^{\mathrm{cur}}=\random{P}^{\mathrm{new}}$\;
+  Initialize parameters $\set{P}_{\mathrm{new}}$\;
+  \Repeat{$\set{P}_{\mathrm{cur}}$ converge}{
+    $\set{P}_{\mathrm{cur}}=\set{P}_{\mathrm{new}}$\;
   \textbf{E step} Evaluate responsibilities~\eqref{eq:responsibilities}\;
@@
 \end{algorithm2e}
-\todo{
-  We can see the similarities to K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
-  Instead of affecting the observed data to a cluster with 100\% of certainty, \EM{} gives a responsibility (posterior probability of point belonging to cluster) for each cluster.
-  Then, when we solve the \textbf{M Step} we take into account all the data, but weighted by these responsibilities.
-}
 
-\section{Using the algorithm to estimate $P_{i}^{1}[k]$}\label{sec:using-em-algorithm}
-\todo{
-  Here to simplify the notation for this section we concentrate on a single agent, and in a single step $k$, so we drop the subscript $i$ and the time dependency $[k]$.
+We can see some similarities to the K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
+Instead of affecting the observed data to a cluster with 100\% of certainty (\emph{hard assignment}), \EM{} uses the responsibilities (\emph{soft assignment}) for each cluster.
+Then, when we solve the \textbf{M Step}, rather than using only the data assigned to update the parameters of a certain cluster, we take into account all data, whose contribution to the update is weighted by their associated responsibilities.
 
-  % First we need to define our model.
-}
-% TODO(accacio): change subscript n
-\todo{
-  For any component of $\tilde{\vec{\lambda}}$, given an input $\randomvec{\theta}_{o}$,  we can define a latent variable $\random{z}_{o}$ which determine to which zone ${z=\{1:Z\}}$ the $\randomvec{\theta}_{o}$ belongs and which parameters $P_{i}^{z}$ and $\vec{s}^{z}$ were used
-to generate an observation $o$ of the $c^{\mathrm{th}}$ component of $\tilde{\lambda}$, $\random{\lambda}_{c,o}$.
-
-}
-\todo{
-  We propose to transform~\eqref{eq:linear_cheating} into a random \pwa{} for each component $c$ of $\lambda$:
-}
-  \begin{equation}
-    \label{eq:linear_cheating_random}
-    \random{\lambda}_{c,o}=
-        \begin{cases}
-      -\tilde{P}_{c}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}_{c}}^{1}&\text{with probability}\ \pi_{1} \\
-      \qquad\quad \vdots&\qquad\quad \vdots\\
-      -\tilde{P}_{c}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}_{c}}^{Z}&\text{with probability}\ \pi_{Z} \\
-    \end{cases}
-  \end{equation}
-  where $\tilde{P}_{c}^{j}$ and $\tilde{\vec{s}}_{c}^{j}$ are the $c^{\mathrm{th}}$ row of  $\tilde{P}^{j}$ and $\tilde{\vec{s}}^{j}$.
-
-\todo{
-  As a consequence of this formulation, we have  \[\probability{\randomvec{z}_{o}=z}=\pi_{z},\]
-  and since $\vec{\theta}$ is our input, we know \emph{almost surely} their values: \[\probability{\randomvec{\theta}_{o}}=1.\]
-}
-\todo{
-  With this data, we can use the following graph in Fig.~\ref{fig:model} to represent the model for the $O$ observations of generated $\random{\lambda}_{c}$, where $\phi_{c}=\{P_{c}^{z},\vec{s}_{c}^{z}\}$ for ${z=\{1:Z\}}$ are the parameters we want to estimate for each zone $z$ and component $c$.
-
-}
-
-\usetikzlibrary{arrows.meta}
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
-    \node at (1.2,-1.3) {O};
-    \node (pi) at (-2.5,.5) {$\pi$};
-    \node (phi) at (2.5,-1.) {$\phi_{c}$};
-
-    \graph [edge quotes={fill=white,inner sep=1pt},
-    clockwise=3,nodes={circle,draw,rotate=-60,minimum width=1cm}] {
-      a/"${\randomvec{\theta}_{o}}$"[rotate=60],b/"${\random{\lambda}_{c,o}}$"[rotate=60],c/"${\random{z}}_{o}$"[rotate=60];
-      {a,c} ->[-latex] b;
-      (pi) ->[thick,{Circle[length=3.pt]}-latex] c;
-      (phi) ->[thick,{Circle[length=3.pt]}-latex] b;
-      };
-
-  \end{tikzpicture}
-  \caption{Graph representation of generation of $\random{\lambda}_{c,o}$.}\label{fig:model}
-\end{figure}
-\todo{
-Similar as in~\cite{FariaSoromenho2010}, we can model ${\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=z,\phi_{c}}}$ as an univariate Gaussian distribution with mean \[f_{c}^{z}(\randomvec{\theta})=-\tilde{P_{c}^{z}}\randomvec{\theta}_{o}-\tilde{\vec{s}_{c}}^{z}\] and variance ${\sigma^{z}}^2\to 0$:
-}
-\begin{equation}\label{eq:1}
-\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=z,\phi_{c}}=\mathcal{N}(\random{y};f_{c}^{z}(\randomvec{\theta}_{o}),{\sigma^{z}}^{2}).
-\end{equation}
-
-The responsibilities \({{\zeta(z)}_{c,o}=\probability{\random{z}_{o}=z\mid\random{\lambda}_{c,o},\randomvec{\theta}_{o},\phi_{c}}}\) can be calculated as
-\[\zeta(z)_{c,o}=\frac{\pi_{z}\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=z,\phi_{c}}}{\sum\limits_{j=1}^{Z}\pi_{j}\probability{\random{\lambda}_{c,o}\mid\randomvec{\theta}_{o},\random{z}_{o}=j,\phi_{c}} }\]
-Observe that by taking
-\begin{equation}\label{eq:argmaxz}
-\underset{z}{\argmax} {\zeta(z)}_{c,o},
-\end{equation}
-we can get the most probable $z$-zone which generated the observation $\random{\lambda}_{c,o}$.
-
-Then, as in~\cite[Chapter 9]{Bishop2006}, the expected value of the complete-data log-likelihood can be computed by
-\begin{equation}
- Q(\phi_{c},\phi_{c}^{\mathrm{cur}})=\expectation[\probability{\random{z}\mid \random{\lambda}_{c},\randomvec{\theta},\phi_{c}^{\mathrm{cur}}}]{\ln\probability{\random{\lambda}_{c},\randomvec{\theta},\random{z},\phi_{c}}},
-\end{equation}
-which can be developed to
-\begin{equation}\label{eq:completedataLogLikelihoodestimation}
-  % Q(\phi_{c},\phi_{c}^{\mathrm{cur}})=
-  \small
-  \sum_{z=1}^{Z}\sum_{o=1}^{O}\zeta(z)_{c,o}^{\mathrm{cur}}\left(\ln \pi_{z}-\frac{\ln{\sigma^{z}}^{2}}{2}-\frac{{{\left(\random{\lambda}_{c,o}-f_{c}^{z}(\randomvec{\theta}_{o})\right)}^{2}}}{2{\sigma^{z}}^{2}}\right)+\ \mathrm{cnt},
-\end{equation}
-where the constant term is independent of any parameter in $\phi_{c}$ and it is not computed.
-
-Calculating the gradient of~\eqref{eq:completedataLogLikelihoodestimation} with respect to vectors $\vec{\phi}_{c,z}={[{P_{c}^{z}}\ \vec{s}_{c}^{z}]}^{T}$ and making it vanish, it is possible to calculate an optimal new estimate $\vec{\phi}_{c,z}^{\mathrm{new}}$ using:
+Here we introduce a variable ${\phi^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$ so we can calculate the new estimates of $\set{P}$ from~\eqref{eq:Mstep}.
+Using the KKT conditions~\cite{BoydVandenberghe2004}, we can find an optimal solution for the problem in~\eqref{eq:Mstep}, by
+taking the gradients of~\eqref{eq:completedataLogLikelihood_expectation} with respect to vectors $\vec{\phi}^{z}$ and making them vanish.
+Because of the multidimensional nature of the problem, some matrix operations are needed to synthesize the results.
+After those operations, we have a matricial solution that yields the optimal estimates $\vec{\phi}_{z}^{\mathrm{new}}$:
 \begin{equation}
   \label{eq:mstepestimation}
-  \vec{\phi}_{c,z}^{\mathrm{new}}=\pseudoinv{(\Xi_{c,z}\randomvec{\theta})}\Xi_{c,z}\randomvec{\lambda}_{c},
+  \vec{\phi}_{z}^{\mathrm{new}}=\pseudoinv{(\Xi_{z}\random{\Omega})}\Xi_{z}\vectorize{\random{\Lambda}},
 \end{equation}
 where
-${\randomvec{\theta}={[ {[\randomvec{\theta}_{1}^{T}\ 1]}\ \dots\ [\randomvec{\theta}_{O}^{T}\ 1] ]}^{T}}$%
-,
-${\randomvec{\lambda}_{c}={[\random{\lambda}_{c,1}\ \dots\ \random{\lambda}_{c,O}]}^{T}}$%
+${\random{\Omega}=[\hadamard{(\Upsilon \random{\Theta}\Delta)}{Y};G]}$,
+with matrices
+${\Upsilon=\kron{\1_{n}^{T}}{I_{n}}}$,
+${\Delta=\kron{I_{N}}{\1_{n}^{T}}}$,
+${Y=\kron{G}{\1_{n}}}$,
+${G=\kron{\1_{N}^{T}}{I_{n}}}$,
 and
-${\Xi_{c,z}={\diag(\sqrt{{\zeta(z)}_{c,1}},\cdots,\sqrt{{\zeta(z)}_{c,O}})}}$.
+\[{\Xi_{z}={\diag(\sqrt{{\zeta(z_{z1};\set{P}_{\mathrm{cur}})}}I_{n},\cdots,\sqrt{{\zeta(z_{zO};\set{P}_{\mathrm{cur}})}}I_{n})}}.\]
 
@@
 
-Once all $\tilde{P}_{c}^{z}$ and $\tilde{\vec{s}}_{c}^{z}$ are estimated, we can stack them creating estimates of $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$.
+Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use in our mitigation scheme proposed in \S\ref{sec:attack}.
 
-Observe that
-
+\todo {
+  Observe that not necessarily the parameters estimated associated to a value z corresponds to the z shown in
+}
 % \begin{equation*}
@@
 \begin{ack}
-Simon Leglaive
+
 \end{ack}

### 2022-02-03 article/article.tex
change structure and add some comments

@@
-\loop\iftrue%
-\errmessage{This article is copyrighted by ___ and should not be TeXed}%
-\repeat%
+% \loop\iftrue%
+% \errmessage{This article is copyrighted by ___ and should not be TeXed}%
+% \repeat%
 
@@
 \section{INTRODUCTION}
+\todo{Some citations:
+  \cite{MaestreEtAl2021}
+
+}
 % % % \mpc{} \dmpc{}
@@
 \end{equation}
-where ${\Gamma_{i}:\R^{n_{u}\times n_{u}}}$, ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.
+where ${\Gamma_{i}:\R^{n_{u}\times n_{u}}}$\todo{$\Gamma_{i}:\semidefpos$?}, ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.
 
@@
 \begin{matrix}
-
-\underset{\uiseq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
+\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
 \mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:upositive}\ \mathrm{and}\ \eqref{eq:constraint}
@@
 
-$\optglobobj$ denotes the optimal value of the problem~\ref{Pb:GOP}, and ${\optuiseq}$ is the optimal control sequences.
-the problem~\ref{Pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
+$\optglobobj$ denotes the optimal value of the problem~\ref{Pb:GOP}, and ${\optuseq}$ is the optimal control sequences for all agents.
+The problem~\ref{Pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
 \end{problem}
@@
     \begin{aligned}
-      J^{\star}&=\underset{\thetaiseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\thetaik)\\
+      J^{\star}&=\underset{\thetaseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\thetaik)\\
       \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\thetaik\preceq\umax
@@
 
-\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$. Where ${c=\pi_{\vec{u}_{i}(k:k+N_{p}-1|k)}=N_{p}\pi_{\uik}}$.}
+Where ${\set{H} = \setbuild{\vec{\theta}}{\sum_{i=1}^{M}\vec{\theta}_{i}=\vec{u}_{\max}}}$, $\vec{\theta}$ are the allocations for all agents.
+\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$. Where ${c=N_{p}\pi_{\uik}}$.}
 
@@
 \end{equation}
+
+where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$.
+\todo{Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$}
+
+\todo{
+  But as the coordinator must also assure that ${\thetaik\succeq\0}$ so the local problems are feasible, we set 0 as lower bounds:
+  }
   \todo[verify if need to use max]{
@@
 
-\section{Attack in DMPC scheme}\label{sec:attack}
-
-\begin{equation}\label{eq:cheating}
-\lambdaicheat=\gamma_{i}(\lambdai)
-\end{equation}
-
 \subsection{Quadratic Case --- Formal Analysis}\label{ssec:FA}
@@
 
-\todo{As seen in~\cite{BemporadEtAl2002,BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
+\todo{As seen in~\cite{BemporadEtAl2002} and in~\cite{BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
 Similarly, we can get an explicit  solution for its dual variables $\vec{\lambda}_{i}$, which is also a \pwa{} function with respect to $\vec{\theta}_{i}$:}
@@
 
-\begin{challenge}
-It is important to note that since the actual state is unknown we cannot anticipate how the space is partitioned, nor the values of $\vec{s}_{i}[k]$.
+\begin{challenge}\label{ch:partition_unknown}
+  It is important to note that since the actual state is unknown we cannot anticipate the partition of the space.
 \end{challenge}
- As each constraint can be active or inactive, for a group of $n$ constraints, \todo{potentially} we can have $N=2^{n}$ different combinations of active/inactive sets, consequently $N$ regions (or zones).
-\todo{
-  Here we say potentially because depending on the constraints, some of these combinations can be infeasible.\question{Remove??}
-}
+\begin{challenge}\label{ch:zone_unknown}
+ For the same reason the values of $\vec{s}_{i}[k]$ are also unknown.
+\end{challenge}
+
+ As each constraint can be active or inactive, for a group of $n$ constraints, we can have at most $N=2^{n}$ different combinations of active/inactive sets and consequently $N$ regions (or zones).
+ \begin{remark}\label{re:N_different_zones}
+  Observe that depending on the constraints, some of these combinations can redundant or even make the optimization infeasible.
+ \end{remark}
+ \begin{assumption}
+   We assume that none of these cases in Remark~\ref{re:N_different_zones} can occur, and we have $N$ zones.
+ \end{assumption}
 
@@
 
-\subsection{Detection and mitigation}
+\section{Attack in DMPC scheme}\label{sec:attack}
+
+\begin{equation}\label{eq:cheating}
+\lambdaicheat=\gamma_{i}(\lambdai)
+\end{equation}
+
+% \subsection{Detection and mitigation}
 \begin{assumption}
@@
 \begin{assumption}
-  We suppose the agent chooses a linear function such as $\tilde{\lambdai}=\gamma_{i}(\vec{x})=\Tik\vec{x}$ resulting in:
+  equation
+  T inversivel
+  We suppose the agent chooses a linear function such as \[ {\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i}} \] resulting in:
   \begin{equation}
@@
         \begin{cases}
-      -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}\thetai \preceq z_{1} \\
+      -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
       \qquad\quad \vdots&\qquad\quad \vdots\\
-      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}\thetai \preceq z_{N} \\
+      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
     \end{cases}

### 2022-02-07 article/article.tex
correction Simon

@@
 
-The main objective of the \EM{} algorithm is to find maximum likelihood estimators for models with latent variables.
+The main objective of the \EM{} algorithm, seen in~\cite{DempsterEtAl1977}, is to find maximum likelihood estimators for models with latent variables.
 
@@
 
-For an observation $o$, we have the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$.
-As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a expansion of the model referred as \emph{mixture of switching regressions} in~\cite{QuandtRamsey1978} and \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010}, which we will call \emph{mixture of switching affine regressions}, since our regressors have a linear term (matrices $\tilde{P}^{z}$) and a constant term (vectors $\tilde{\vec{s}}^{z}$):
+For an observation ${o\in\set{O}=\{1\mathbin{:}O\}}$, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$.
+As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a expansion of the model referred as \emph{mixture of switching regressions} in~\cite{QuandtRamsey1978} and \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010}, which we will call \emph{mixture of affine regressions}, since our regressors have a linear term (matrices $\tilde{P}^{z}$) and a constant term (vectors $\tilde{\vec{s}}^{z}$):
 % TODO(accacio): change subscript n
@@
     -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{with probability}\ \pi_{Z} \\
-    \end{cases}
+  \end{cases}.
 \end{equation}
 
-As we can see, this observation does not gives us the complete data, since we don't know which zone generated the response, but only the probability of the input being in that zone.
-We consider the complete data consisting of the input, the response, and a non-observable variable.
-Therefore, for each observation $o$ there exists a latent random variable ${\random{z}_{o}\in\set{Z}}$, with associated probabilities ${\Pi=\{\pi_{1},\dots,\pi_{Z}\}}$ such prior probability
+Each couple of observed input and response variables $(    \random{\lambda}_{o}, \randomvec{\theta}_{o})$ is associated in this model to a latent unobserved random variable ${\random{z}_{o}\in\set{Z}}$ that indicates from which of the $Z$ affine regression models in~\eqref{eq:linear_cheating_random} the response variable was obtained.
+The latent variable $\random{z}_{o}$ follows a categorical prior distribution, with associated probabilities ${\Pi=\{\pi_{1},\dots,\pi_{Z}\}}$:
    \begin{equation}
      \label{eq:prob_zo_equal_z}
-\probability{\random{z}_{o}=z}=\pi_{z}.
+\probability{\random{z}_{o}=z}=\pi_{z} \in [0,1], \qquad \sum_{z=1}^{Z} \pi_{z} = 1.
 \end{equation}
-Since $\vec{\theta}$ is our input, we know \emph{almost surely} their values giving
+Since $\vec{\theta}$ is our input,
+we consider a non-informative improper probability density function~\cite{ChristensenEtAl2010}
+% SL: BTW, maybe we should use different notations for "the probability that some discrete random variable (e.g.
+z_0) is equal to something, and the probability density function of a continuous random variable (e.g.
+theta_0 and lambda_0).
+
 \begin{equation}
   \label{eq:theta_almost_surely}
-  \probability{\randomvec{\theta}_{o}}=1.
+  \probability{\randomvec{\theta}_{o}} \,\propto \,1.
 \end{equation}
-With this information we can use the graph in Fig.~\ref{fig:model} to represent the model.
+
+Given the input and latent variables, the response variable $\random{\lambda}_{o}$ is modeled as a multivariate normal random variable with probability density function
+\begin{equation}
+  \label{eq:multivariate_gaussian}
+\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z; \set{P}^{z}} = \mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}}),
+\end{equation}
+where, following \eqref{eq:linear_cheating_random}, the mean vector is defined by
+\[f(\randomvec{\theta}_{o};(P,\vec{s}))=-{P}\randomvec{\theta}_{o}-{\vec{s}},\]
+and the covariance matrix ${\Sigma^{z}}$ tends to $0$.
+
+This mixture of affine regression model is represented as a probabilistic graphical model in Fig.~\ref{fig:model}, and it corresponds to the following factorization of the complete-data (i.e., observed and latent data) likelihood, \cite{Bishop2006}:
+\begin{align}\label{eq:completedataLikelihood}
+  \probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}= \prod_{o=1}^{O}\prod_{z=1}^{Z}\big[&\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z;\set{P}^{z}} \nonumber \\
+& \times \probability{\random{z}_{o}=z}\probability{\randomvec{\theta}_{o}}\big]^{\indicator{\random{z}_{o}=z}},
+\end{align}
+
+Ideally, we would like to estimate the unknown model parameters $\set{P}$ by maximizing the log-marginal likelihood $\ln \probability{\random{\Theta},\random{\Lambda};\set{P}}$, computed by marginalizing the latent variables $\random{Z}$ in~\eqref{eq:completedataLikelihood}.
+Unfortunately, this optimization problem does not admit a closed-form analytical solution.
+Instead, we can exploit the latent-variable structure of the model to derive an \EM{} algorithm,~\cite{DempsterEtAl1977}, which is a convergence-guaranteed iterative algorithm ensuring at each iteration a monotonic increase of the log-marginal likelihood.
+
 \begin{figure}[b]
@@
 \end{figure}
+% SL: observed variables are generally represented as shaded nodes in probabilistic graphical models, and latent variables as empty nodes.
 
-As said, the objective of the \EM{} algorithm is to maximize the likelihood, more precisely the complete data log-likelihood $\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}$.
-Using the dependency graph in Fig.~\ref{fig:model} yields
-    \begin{equation}\label{eq:completedataLogLikelihood}
-      \ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}=\ln \prod_{o=1}^{O}\probability{\randomvec{\theta}_{o},\randomvec{\lambda}_{o},\random{z}_{o};\set{P}} .
-    \end{equation}
-where the probability $\probability{\randomvec{\theta}_{o},\randomvec{\lambda}_{o},\random{z}_{o};\set{P}}$ is
-\begin{equation}\label{eq:condional}
- \prod_{z=1}^{Z}{[\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z;\set{P}^{z}}\probability{\random{z}_{o}=z}\probability{\randomvec{\theta}_{o}}]}^{\indicator{\random{z}_{o}=z}},
-\end{equation}
-and $\set{P}^{z}$ is the $z$-th tuple in $\set{P}$.
+Now we can formalize the \EM{} problem.
+\begin{problem}{Expectation Maximization Problem}\label{pb:EM}
 
-For the likelihood $\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z,\set{P}^{z}}$, we use multivariate normal probability density functions
-\begin{equation}
-  \label{eq:multivariate_gaussian}
-\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}}),
-\end{equation}
-where the mean is
-\[f(\randomvec{\theta}_{o};(P,\vec{s}))=-{P}\randomvec{\theta}_{o}-{\vec{s}}\]
-and the covariance is ${\Sigma^{z}}$, which tends to $0$.
+  Given a set of observed data $(\random{\Theta},\random{\Lambda})$, estimate the unknown latent variables $\random{Z}$ and parameter set $\set{P}$ by iteratively maximizing \todo{a function of} the complete-data log likelihood $\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}$.
+\end{problem}
 
-Using~\eqref{eq:prob_zo_equal_z},~\eqref{eq:theta_almost_surely}, and~\eqref{eq:multivariate_gaussian} in~\eqref{eq:completedataLogLikelihood} yields
+Using~\eqref{eq:prob_zo_equal_z},~\eqref{eq:theta_almost_surely}, and~\eqref{eq:multivariate_gaussian} in~\eqref{eq:completedataLikelihood} yields {\color{blue}the following expression of the complete-data log-likelihood:}
 \begin{equation}\label{eq:completedataLogLikelihood_complete}
@@
 
-Now we can formalize the \EM{} problem.
-\begin{problem}{Expectation Maximization Problem}\label{pb:EM}
+But as mentioned, we do not observe the complete data $(\random{\Theta},\random{\Lambda},\random{Z})$, instead we observe only $(\random{\Theta},\random{\Lambda})$ and all the information on $\random{Z}$ given these observations is carried by the posterior probabilities  ${\zeta(z_{zo};\set{P})=\probability{\random{z}_{o}=z|\randomvec{\lambda}_{o},\randomvec{\theta}_{o};\set{P}}}$, also called \emph{responsibilities}, which can be calculated as
+\todo[Change Notation?]{Change Notation?}
+% SL: the notation z_{zo} is a bit confusing and I am not sure it was defined.
+It's actually related to the fact that you use z_0 as the latent variable and $z$ as the values it can take in \mathcal{Z}.
+Using another variable for the latter may help clarifying.
 
-  Given a set of observed data $(\random{\Theta},\random{\Lambda})$, estimate the unknown latent variables $\random{Z}$ and parameter set $\set{P}$ which maximize the complete-data log likelihood in~\eqref{eq:completedataLogLikelihood_complete}.
-\end{problem}
-
-But as mentioned, we do not observe the complete data $(\random{\Theta},\random{\Lambda},\random{Z})$, instead we observe only $(\random{\Theta},\random{\Lambda})$ and all information we can get from $\random{Z}$ is the posterior probability  ${\zeta(z_{zo};\set{P})=\probability{\random{z}_{o}=z|\randomvec{\lambda}_{o},\randomvec{\theta}_{o};\set{P}}}$, also called \emph{responsibility}, which can be calculated as
 \begin{align}
@@
 Observe that by taking
+\todo[Change Notation?]{Change Notation?}
 \begin{equation}\label{eq:argmaxz}
@@
 
-So, rather than using the complete data log-likelihood~\eqref{eq:completedataLogLikelihood_complete}, which we do not have, we use its expectation with respect to the prior calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$,  ${Q\left(\set{P},\set{P}_{\mathrm{cur}}\right)=\expectation[{\zeta(z_{zo};\set{P}_{\mathrm{cur}}}]{\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}}$:
-\begin{equation}
+So, rather than using the complete data log-likelihood~\eqref{eq:completedataLogLikelihood_complete}, which we do not have, since the latent variables are not observed by definition, the \EM{} algorithm uses its expectation with respect to the posterior calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$:
+\begin{align}
   \label{eq:completedataLogLikelihood_expectation}
- Q\left(\set{P},\set{P}_{\mathrm{cur}}\right) =
-  \sum_{o=1}^{O}\sum_{z=1}^{Z}  \zeta(z_{zo};\set{P}_{\mathrm{cur}})\alpha_{zo}
-      ,
-\end{equation}
+ Q\left(\set{P},\set{P}_{\mathrm{cur}}\right) &= \expectation[{\zeta(z_{zo};\set{P}_{\mathrm{cur}})}]{\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}
+\end{align}
+and since \[\expectation[{\zeta(z_{zo};\set{P}_{\mathrm{cur}})}]{\indicator{\random{z}_{o}=z}}=\zeta(z_{zo};\set{P}_{\mathrm{cur}}),\]
+we can rewrite~\eqref{eq:completedataLogLikelihood_expectation}:
+\begin{align}
+  \label{eq:completedataLogLikelihood_expectation_developed}
+Q\left(\set{P},\set{P}_{\mathrm{cur}}\right) &= \sum_{o=1}^{O}\sum_{z=1}^{Z}  \zeta(z_{zo};\set{P}_{\mathrm{cur}})\alpha_{zo},
+\end{align}
+\todo{\begin{remark}
+For the first iteration, the model parameters must be initialized.
+A discussion about the initialization of the parameters is beyond the scope of this article, and we refer the reader to~\cite{Bishop2006},~\cite{DempsterEtAl1977} and their references.
+\end{remark}}
 
@@
 \end{equation}
+For the next iteration of the algorithm, this new estimate $\set{P}_{\mathrm{new}}$ will be used as $\set{P}^{\mathrm{cur}}$.
 
 With this information we can describe the \EM{} algorithm in two parts, first calculate the responsibilities $\zeta(z_{zo};\set{P}_{\mathrm{cur}})$, and then update the parameters.
-Algorithm~\ref{alg:em} adpated from~\mbox{\cite[Chapter 9]{Bishop2006}} describes the steps.
+Algorithm~\ref{alg:em} adapted from~\mbox{\cite[Chapter 9]{Bishop2006}} describes the steps.
 
@@
 
-We can see some similarities to the K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
-Instead of affecting the observed data to a cluster with 100\% of certainty (\emph{hard assignment}), \EM{} uses the responsibilities (\emph{soft assignment}) for each cluster.
-Then, when we solve the \textbf{M Step}, rather than using only the data assigned to update the parameters of a certain cluster, we take into account all data, whose contribution to the update is weighted by their associated responsibilities.
-
-Here we introduce a variable ${\phi^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$ so we can calculate the new estimates of $\set{P}$ from~\eqref{eq:Mstep}.
+Here we introduce a variable ${\vec{\phi}^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$ so we can calculate the new estimates of $\set{P}$ from~\eqref{eq:Mstep}.
 Using the KKT conditions~\cite{BoydVandenberghe2004}, we can find an optimal solution for the problem in~\eqref{eq:Mstep}, by
@@
 \[{\Xi_{z}={\diag(\sqrt{{\zeta(z_{z1};\set{P}_{\mathrm{cur}})}}I_{n},\cdots,\sqrt{{\zeta(z_{zO};\set{P}_{\mathrm{cur}})}}I_{n})}}.\]
+% SL: I find it surprising that you have square roots of the responsabilities in this equation. Are you sure it is correct?
 
-As we can see,~\eqref{eq:mstepestimation} is a weighted Least-Squares, that uses the responsibilities as weights, ignoring those values closer to $0$.
+As we can see,~\eqref{eq:mstepestimation} is the solution of a weighted Least-Squares, having the responsibilities as weights, adjusting the contribution of all observations to the regression model of index $z$.
+We can see some similarities to the K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
+Instead of affecting the observed data to a zone with 100\% of certainty (\emph{hard assignment}), \EM{} uses the responsibilities (\emph{soft assignment}) for each zone.
+Then, when we solve the \textbf{M Step}, rather than using only the data assigned to update the parameters of a certain zone, we take into account all data, whose contribution to the update is weighted by their associated responsibilities.
 

### 2022-02-07 code/calculate_responsabilities.m
working mitigation algo for active sets

@@ -1,0 +1,29 @@
+function Responsabilities = calculate_responsabilities(x, y, Phi, Sigma, Pi)
+    %% modes x N responsability matrix
+    modes=size(Phi,1);
+    [n, N]=size(x);
+
+    respNum = zeros(size(x,2),modes);
+
+
+    A=kron(ones(n,1),eye(n));
+    B=kron(eye(N),ones(1,n));
+    C=kron(repmat(eye(n),1,N),ones(n,1));
+    x_lin=sparse([(A*x*B).*C;kron(repmat(eye(n),1,N),ones(1,1))]);
+    y_cell = cellfun(@(x) x',mat2cell(y',ones(1,N))','UniformOutput',0);
+    y_diag=sparse(blkdiag(y_cell{:}));
+    %= calculate numerator for each mode
+    % for i=1:modes
+    for i=1:modes
+        y_lin=reshape(x_lin'*-Phi(i,:)',n,N);
+        y_lin_cell = cellfun(@(x) x',mat2cell(y_lin',ones(1,N))','UniformOutput',0);
+        y_lin_diag=sparse(blkdiag(y_lin_cell{:}));
+        err_diag=y_diag-y_lin_diag;
+        SigmaInvMat=kron(sparse(eye(N)),sparse(Sigma(:,:,i)\eye(n)));
+
+        respNum(:,i)=sparse(Pi(1,i)*(1/(2*pi)^(n/2))*(1/sqrt(det(Sigma(:,:,i))))*exp(-1/2*max(err_diag'*SigmaInvMat*err_diag))');
+    end
+    normalizing_constant=sum(respNum,2);
+    Responsabilities =(respNum./normalizing_constant)';
+
+end

### 2022-02-07 code/decompositionNagents.m

@@ -1,0 +1,342 @@
+%= Clean variables
+close all
+clear
+
+paren = @(x, varargin) x(varargin{:}); %
+                                       % apply index in created elements
+curly = @(x, varargin) x{varargin{:}}; %
+
+%% Options
+
+%= Optimization settings
+options = optimset('Display', 'off');
+% options = [];
+rand('seed',2);
+
+%= Simulation parameters
+M=4;    %= # of systems
+Te=.25; %= Sampling
+Np=4;   %= Prediction horizon
+
+simK = 50;    %= Simulation horizon
+negotP = 100; %= max # of iteration for each negotiation
+err_theta=1e-2; %= err to test theta convergence in each negotiation
+
+chSetpoint = 0; %= Change setpoint?
+selfish = 1;    %= Selfish behaviour?
+secure = 1;     %= Secure algorithm?
+for chSetpoint=1
+for selfish=0:1
+for secure=0:1
+tic
+%= Global constraint
+Umax=4;
+% Umax=2;
+
+%= Input bounds
+u_min=0;
+% u_min=-inf;
+u_max=Umax;
+% u_max=inf;
+
+%% Define systems
+% TODO(accacio): move to a function
+
+Cair_mean=8;
+Cwalls_mean=5;
+Roaia_mean=5;
+Riwia_mean=2.5;
+Rowoa_mean=1.1;
+
+Cair  = repmat(Cair_mean,1,M)+(-.5+rand(1,M));
+Cwalls = repmat(Cwalls_mean,1,M)+(-.5+rand(1,M));
+Roaia = repmat(Roaia_mean,1,M)+(-.5+rand(1,M));
+Riwia = repmat(Riwia_mean,1,M)+(-.5+rand(1,M));
+Rowoa = repmat(Rowoa_mean,1,M)+(-.5+rand(1,M));
+
+%= Define continuos systems using 3R2C
+for i=M:-1:1 % make it backward to "preallocate"
+    csys(:,:,1,i)=model_3R2C(Roaia(i),Riwia(i),Rowoa(i),Cwalls(i),Cair(i));
+end
+ni=size(csys.B(:,:,1,1),2); %= # of inputs
+ns=size(csys.A(:,:,1,1),2); %= # of states
+n=Np*ni; % constant used everywhere
+dsys=c2d(csys,Te); %= Discretize systems
+
+%= Output prediction matrices, such $\vec{Y}=\mathcal{M}\vec{x}_k+\mathcal{C}\vec{U}$
[…]

### 2022-02-07 code/emgm_estimate.m

@@ -1,0 +1,64 @@
+function [Phi,Responsabilities,pi_new,Sigma] = emgm_estimate(X,Y,phi_init,modes,emMaxIter,maxErr)
+% EMGM_NESTIMATE - ESTIMATE N DIMENSIONAL
+
+    Pi=repmat(1/modes,1,modes);
+    OldclusterSize=zeros(1,modes);
+    [n, ~]=size(X);
+
+    if(isempty(phi_init))
+        % TODO(accacio): do some multidimensional magic
+        % indexpts=randi([1 size(x,2)-1],1,modes);
+        % dx=max(x(:,2:end)-x(:,1:end-1),[],2);
+        % dx=num2cell(dx,2);
+
+        % y_square=reshape(y,sqrt(size(y,2)),sqrt(size(y,2)));
+        % [V{1:2}]=gradient(y_square,dx{:});
+        % grad=cell2mat(cellfun(@(x) reshape(x,[],1),V,'UniformOutput',0))';
+        % C=grad(:,indexpts);
+        % x_indexed=x(:,indexpts);
+        % y_indexed=y(:,indexpts);
+        % d=y_indexed-sum(C.*x_indexed);
+        % C=reshape(C,2,1,modes);
+        % d=reshape(d,1,1,modes);
+        % C=5*rand(size(x,1),1,modes);
+        % d=5*rand(1,1,modes);
+        Phi=20*rand(modes,n^2+n);
+
+    else
+        Phi=phi_init;
+    end
+
+    OldPhi=zeros(size(Phi));
+    % eps=100;
+    eps=10000;
+    Sigma(:,:,1:modes)=repmat(eps*eye(n),1,1,modes);
+
+    for emInd=1:emMaxIter
+
+        Responsabilities=calculate_responsabilities(X,Y,Phi,Sigma,Pi);
+
+        [Phi, pi_new] = update_parameters(X, Y, Responsabilities);
+
+        [~,z_hat]=max(Responsabilities,[],1);
+        for i=1:modes
+            z_i=find(z_hat==i);
+            clusterSize(i)=size(z_i,2);
+            if OldclusterSize(i)==clusterSize(i)
+                % Sigma(:,:,i)=Sigma(:,:,i)*.9;
+                Sigma(:,:,i)=Sigma(:,:,i)*.01;
+            else
+                Sigma(:,:,i)=Sigma(:,:,i)*1.;
+            end
+        end
+
+        % Phi
[…]

### 2022-02-07 code/update_parameters.m

@@ -1,0 +1,21 @@
+function [Phi, pi_new] = update_parameters(X, Y, Responsabilities)
+% UPDATE_PARAMETERS -
+    modes=size(Responsabilities,1);
+    [n N]=size(X);
+
+    A=kron(ones(n,1),eye(n));
+    B=kron(eye(N),ones(1,n));
+    C=kron(repmat(eye(n),1,N),ones(n,1));
+    theta=sparse([(A*X*B).*C; kron(ones(1,N),eye(n))]');
+
+    N_k=sum(Responsabilities,2);
+    pi_new(:)=N_k/N;
+    for i=1:modes
+        responsabilities=Responsabilities(i,:);
+        resp2=cellfun(@(x) x*eye(n),mat2cell(responsabilities',ones(1,N)),'UniformOutput',0);
+        Gamma=sqrt(sparse(blkdiag(resp2{:})));
+
+        Phi(i,:)=-((Gamma*theta)\(Gamma*Y(:)));
+    end
+
+end

### 2022-02-07 article/article.tex
some changes after meeting

@@
   \label{eq:linear_cheating_random}
-  \random{\lambda}_{o}=
+  \randomvec{\lambda}_{o}=
   \begin{cases}
@@
 
-Each couple of observed input and response variables $(    \random{\lambda}_{o}, \randomvec{\theta}_{o})$ is associated in this model to a latent unobserved random variable ${\random{z}_{o}\in\set{Z}}$ that indicates from which of the $Z$ affine regression models in~\eqref{eq:linear_cheating_random} the response variable was obtained.
+Each couple of observed input and response variables $(    \randomvec{\lambda}_{o}, \randomvec{\theta}_{o})$ is associated in this model to a latent unobserved random variable ${\random{z}_{o}\in\set{Z}}$ that indicates from which of the $Z$ affine regression models in~\eqref{eq:linear_cheating_random} the response variable was obtained.
 The latent variable $\random{z}_{o}$ follows a categorical prior distribution, with associated probabilities ${\Pi=\{\pi_{1},\dots,\pi_{Z}\}}$:
@@
 \end{equation}
-where, following \eqref{eq:linear_cheating_random}, the mean vector is defined by
+where, following~\eqref{eq:linear_cheating_random}, the mean vector is defined by
 \[f(\randomvec{\theta}_{o};(P,\vec{s}))=-{P}\randomvec{\theta}_{o}-{\vec{s}},\]
@@
     clockwise=3,nodes={circle,draw,rotate=-60,minimum width=1cm}] {
-      a/"${\randomvec{\theta}_{o}}$"[rotate=60],b/"${\random{\vec{\lambda}}_{o}}$"[rotate=60],c/"${\random{z}}_{o}$"[rotate=60];
+      a/"${\randomvec{\theta}_{o}}$"[rotate=60,fill=lightgray],b/"${\random{\vec{\lambda}}_{o}}$"[rotate=60,fill=lightgray],c/"${\random{z}}_{o}$"[rotate=60];
       {a,c} ->[-latex] b;
@@
 \end{figure}
-% SL: observed variables are generally represented as shaded nodes in probabilistic graphical models, and latent variables as empty nodes.
 
@@
 \begin{equation}\label{eq:argmaxz}
-\underset{z_{o}}{\argmax}\ {\zeta(z_{zo};\set{P})},
+\arg\underset{z}{\max}\ {\zeta_{zo}(;\set{P})},
 \end{equation}

### 2022-02-08 code/decompositionNagents.m
delete unused

@@ -149,15 +149,6 @@ selfish_time= [ simK/2;
                 simK;
               ];
 
-identity=@(sys,x) x;
-
-if selfish
-    modify_lambda=@linear_modify;
-else
-    modify_lambda=identity;
-end
-
-
 rho_fun = @(a,b,negot) 1/(a+b*negot);
 % rho = @(a,b,negot) (1/a)/negot;
 a=100;

### 2022-02-08 article/article.tex
simulation, figures, table and some text

@@
 
-\emph{Notation:} In this paper, $\norm{\cdot}$ and $\norm{\cdot}_{F}$ represent the $\ell_{2}$ and Frobenius norms. $\norm{\vec{v}}_{Y}$ is the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
+\emph{Notation:}
+In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
 $\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto set $\set{T}$.
 $\card{x}$ is the number of elements in $x$.
-$\kron{}{}$ represents the Kronecker product.
+$\kron{}{}$, and $\hadamard{}{}$ are the Kronecker and Hadamard products.
 $\vectorize{A}$ vectorizes matrix $A$.
-$\hadamard{}{}$ represents the Hadamard product.
-$n\mathbin{:}i\mathbin{:}j$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where $m=\mathrm{truncate}(\frac{j-n}{i})$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
-$[A,B]$ and $[A;B]$ are horizontal and vertical concatenation of matrices of adequate sizes.
+${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
+$[A,B]$ and $[A;B]$ are horizontal and vertical concatenation of matrices.
 $0_{m\times n}$ and $1_{m\times n}$ are ${m\times n}$ matrices filled with $0$ and $1$.
@@
 $\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
-$\preceq$ is the component-wise less or equal for vectors and matrix inequality for matrices in ${\symmetric=\setbuild{X:\R^{n\times n}}{X'=X}}$.
-${\defpos=\setbuild{X:\symmetric}{X\succ 0}}$, ${\semidefpos=\setbuild{X:\symmetric}{X\succeq 0}}$.
+$\preceq$ is the component-wise less or equal for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X'=X}}$.
+${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
 $I_{c}$ is the identity matrix of size $c$.
@@
 A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
-$\random{a},\random{A},\randomvec{a}$ are random scalar, matrix and vector.
+$\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
 $\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
-${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.
+$\pdf{\cdot}$ is a probability density function. ${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.
 
-\section{PRELIMINARIES AND PROBLEM STATEMENT}\label{sec:PS}
+\section{Problem Statement}\label{sec:PS}
+In this section we present an algorithm to decompose a MPC problem based on the primal decomposition (\cite{BoydEtAl2015}), and a vulnerability that can be exploited by a malicious agent.
+\subsection{Model Predictive Control}\label{ssec:MPC}
+Let us consider a system composed of $M$ discrete-time linear time-invariant agents.
+The dynamics of an agent $i$ is ruled by:
 \begin{equation}
@@
   \label{eq:systems}
-\vec{x}_{i}[k+1]=A_{i}\xik + B_{i}\uik
+\vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
+\vec{y}_{i}[k+1]&=&C_{i}\xik &&\\
 \end{matrix}
 \end{equation}
+Each agent's input $\uik$ is constrained by
+\begin{equation}
+  \label{eq:u_local_constraints}
+  \0\preceq\uik\preceq\vec{u}_{\max},
+\end{equation}
+where ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.
 
+All agents are coupled by a global input constraint
 \begin{equation}
-  \label{eq:upositive}
-  \uik\succeq\0
+  \label{eq:global_constraint}
+  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax,
 \end{equation}
+where \todo{$\Gamma_{i}:\semidefpos^{n_{u}}$}.
 
-\begin{equation}
-  \label{eq:constraint}
-  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax
-\end{equation}
-where ${\Gamma_{i}:\R^{n_{u}\times n_{u}}}$\todo{$\Gamma_{i}:\semidefpos$?}, ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.
-
-\begin{problem}{Global \mpc{} Problem.}\label{Pb:GOP}
+The overall system can be controlled by a \mpc{}, by predicting the dynamics of the system for a finite prediction horizon $N_{p}$ and computing the optimal input by solving the following optimization problem:
+\begin{problem}{Global \mpc{} Problem.}\label{pb:GOP}
 \begin{equation*}
@@
 \underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
-\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:upositive}\ \mathrm{and}\ \eqref{eq:constraint}
+\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
 \left\}\small
@@
 \end{equation*}
-where ${Q_{i}:\semidefpos}$ and ${R_{i}:\defpos}$, $\vik$ is a control objective.
-Reference tracking can be written as ${\vik=\wik-\xik}$, where $\wik$ is a state reference.
-Disturbance rejection can be written as ${\vik=\xik}$.
+where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
+\todo{Reference tracking can be written as ${\vik=\wik-C_{i}\xik}$, where $\wik$ is a state reference.
+Disturbance rejection can be written as ${\vik=C_{i}\xik}$.
+}
 
-$\optglobobj$ denotes the optimal value of the problem~\ref{Pb:GOP}, and ${\optuseq}$ is the optimal control sequences for all agents.
-The problem~\ref{Pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
+$\optglobobj$ denotes the optimal value of the objective function of problem~\ref{pb:GOP}, and ${\optuseq}$ is the optimal control sequences for all agents.
+The problem~\ref{pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
 \end{problem}
 
+Due to the monolithic aspect of problem~\ref{pb:GOP}, the prediction horizon and the number of agents, solving the problem can be computationally intensive.
+One can see that if it were not for the global constraints~\eqref{eq:global_constraint}, the problem would be easily decomposed in $M$ parts and solvable in parallel.
+
 \subsection{Distributed Model Predictive Control}\label{ssec:dMPC}
+&continue from here
+The technique consists of decomposing the
 \todo{\dmpc{},
@@
 
-Where ${\set{H} = \setbuild{\vec{\theta}}{\sum_{i=1}^{M}\vec{\theta}_{i}=\vec{u}_{\max}}}$, $\vec{\theta}$ are the allocations for all agents.
-\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$. Where ${c=N_{p}\pi_{\uik}}$.}
+Where ${\set{H} = \setbuild{\vec{\theta}}{\sum_{i=1}^{M}\vec{\theta}_{i}\preceq \vec{u}_{\max}\ \&\&\ \vec{\theta}_{i}\succeq\0,i }}$.
+\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$.
+}
 
@@
 
-where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$.
-\todo{Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$}
+where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$, with ${c=N_{p}\pi_{\uik}}$.
+Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$.
 
-\todo{
-  But as the coordinator must also assure that ${\thetaik\succeq\0}$ so the local problems are feasible, we set 0 as lower bounds:
-  }
-  \todo[verify if need to use max]{
-\begin{equation}
-  % \label{eq:thetaNegot}
-  \thetai\pplusone=
-  \left\{ \thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
-  ,0\right\}
-\end{equation}
-}
 \no{\todo[delete probably]{
@@
 
-\subsection{Quadratic Case --- Formal Analysis}\label{ssec:FA}
+\subsection{Local Problems --- Formal Analysis}\label{ssec:FA}
 % \todo{As seen in \cite{NogueiraEtAl2021}},
@@
 
-If we take reference tracking, for instance, we have:
+If we choose reference tracking as our, for instance, we have:
 \begin{equation}
@@
 \todo{
-  Observe that $H_{i}:\defpos$.
+  Observe that $H_{i}:\defpos$, and that $\fik$ varies with $k$.
 }
@@
 \end{equation}
-The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active constraints for a given time $k$.
+The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active/inactive constraints of $\Theta_{i}\Uik\preceq\vec{\theta}_{i}$ for a given time $k$.
 
 \begin{challenge}\label{ch:partition_unknown}
-  It is important to note that since the actual state is unknown we cannot anticipate the partition of the space.
+  It is important to note that since the actual state is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
 \end{challenge}
 \begin{challenge}\label{ch:zone_unknown}
- For the same reason the values of $\vec{s}_{i}[k]$ are also unknown.
+ For the same reason, the values of $\vec{s}_{i}[k]$ are also unknown.
 \end{challenge}
@@
  \begin{remark}\label{re:N_different_zones}
-  Observe that depending on the constraints, some of these combinations can redundant or even make the optimization infeasible.
+  Observe that depending on the constraints, some of these combinations can be redundant or even make the optimization infeasible.
  \end{remark}
@@
 
-\section{Attack in DMPC scheme}\label{sec:attack}
+\subsection{Attack in DMPC scheme}\label{ssec:attack}
 
@@
 
+\section{Simulation}
+\begin{figure}[h]
+  \centering
+ \includegraphics[width=\columnwidth]{airtemp_roomI/__ErrorWX_command_normErrH.pdf}
+  \caption{Air temperature in room I and the Frobenius norm of the decision variable $E_{I}[k]$ for different scenarios: nominal (N), selfish behavior without correction (S),
+and selfish behavior with correction.}\label{fig:response3Scenarios}
+\end{figure}
+\begin{figure}[h]
+  \centering
+ \includegraphics[width=\columnwidth]{airtemp_roomI/control.pdf}
+  \caption{Control applied in all rooms for the 3 different scenarios: nominal, selfish behavior without correction, and selfish behavior with correction (C).}\label{fig:response3Scenarios}
+\end{figure}
+
+% TODO(accacio): Verify
+\begin{table}[t]
+  \centering
+  \caption{Comparison of costs $J_{i}$}\label{tab:costsGlobalLocal}
+  \begin{tabular}[t]{cccc} \toprule
+    Agent  & Nominal & Selfish & Selfish + Correction\\
+    \midrule
+    I & -11468 &  -9074 & -11468 \\
+    II & -6536 &  -7151 & -6536 \\
+    III & -3192 &  -4052 & -3192 \\
+    IV & -2849 &  -3694 & -2849 \\
+    Global & -24044 &  -23971 & -24044 \\
+
+    \bottomrule
+  \end{tabular}
+\end{table}
+
 \section{Conclusion}
@@
 or suggest applications and extensions.
-\begin{ack}
-
-\end{ack}
 

### 2022-02-08 code/decompositionNagents.m

@@ -22,13 +22,10 @@ simK = 50;    %= Simulation horizon
 negotP = 100; %= max # of iteration for each negotiation
 err_theta=1e-2; %= err to test theta convergence in each negotiation
 
-chSetpoint = 0; %= Change setpoint?
-selfish = 1;    %= Selfish behaviour?
-secure = 1;     %= Secure algorithm?
-for chSetpoint=1
-for selfish=0:1
-for secure=0:1
-tic
+% TODO(accacio): change name of these variables
+chSetpoint_list = 0; %= Change setpoint?
+selfish_list = 0:1;    %= Selfish behaviour?
+secure_list = 0:1;     %= Secure algorithm?
 %= Global constraint
 Umax=4;
 % Umax=2;
@@ -88,7 +85,7 @@ f_fun=@(Cmat,Mmat,Q,xt,Wt) Cmat'*Q*(Mmat*xt-Wt);
 
 %= Gains Q and R for $\sum_{j=1}^n \|v\|^2_{Q}+\|u\|^2_{R}$
 for i=M:-1:1 % make it backward to "preallocate"
-    Qbar(:,:,i)=1*eye(Np*size(dsys(:,:,1,i).C,1)); % no x no
+    Qbar(:,:,i)=10*eye(Np*size(dsys(:,:,1,i).C,1)); % no x no
     Rbar(:,:,i)=eye(n); % nc x nc
 end
 
@@ -115,7 +112,7 @@ X0(:,1,3) = [15 3.1]';
 X0(:,1,4) = [17. 5.7]';
 
 %= Setpoint
-Wt = [X0(1,1)*1.05;
+Wt = [X0(1,1)*1.5;
       X0(1,2)*1.05;
       X0(1,3)*1.05;
       X0(1,4)*1.05;
@@ -137,14 +134,14 @@ umax(1:M)=u_max;
 
 %= Selfish behavior profile
 % T(:,:,1)=4*eye(n);
-T(:,:,1)=(10*rand(n));
+T(:,:,1)=20*diag(rand(n,1));
 T(:,:,2)=10*eye(n);
 T(:,:,3)=1*eye(n);
 T(:,:,4)=1*eye(n);
 
 %= Time selfish behavior activated
 selfish_time= [ simK/2;
-                simK*2/3;
+                simK;
                 simK;
                 simK;
               ];
@@ -155,16 +152,20 @@ a=100;
 b=100;
 
 %% === Control Loop ===
[…]

### 2022-02-08 code/generate_figures.py

@@ -1,0 +1,292 @@
+#!/usr/bin/env python3
+# import scipy.io
+# mat = scipy.io.loadmat('file.mat')
+# nuitka --recurse-on --python-version=3.6 daoct;
+import sys
+import os
+import getopt
+import math
+import matplotlib.pyplot as plt
+from matplotlib.patches import Circle, Wedge, Polygon
+import matplotlib.font_manager as font_manager
+# import matplotlib
+from matplotlib import rc
+from matplotlib import rcParams
+
+
+# plt.rcParams['text.usetex'] = True
+# plt.rcParams['font.family'] = 'serif'
+# plt.rcParams['font.sans-serif'] = 'cm'
+
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('text', usetex=True)
+import tikzplotlib
+
+import scipy.io as sio
+import numpy as np
+import numpy.matlib
+
+import warnings
+warnings.filterwarnings("ignore")
+# config variables
+
+nominal = sio.loadmat("../data/dmpc4rooms_ineq_chSetpoint__0_selfish__0_secure__0_.mat")
+selfish = sio.loadmat("../data/dmpc4rooms_ineq_chSetpoint__0_selfish__1_secure__0_.mat")
+corrected = sio.loadmat("../data/dmpc4rooms_ineq_chSetpoint__0_selfish__1_secure__1_.mat")
+
+simK=nominal['simK'][0][0]
+
+nominal_Wt=nominal['Wt']
+nominal_xt=nominal['xt']
+nominal_err=nominal['norm_err']
+nominal_u=nominal['uHist'][0]
+
+
+selfish_Wt=selfish['Wt']
+selfish_xt=selfish['xt']
+selfish_err=selfish['norm_err']
+selfish_u=selfish['uHist'][0]
+
+
+corrected_Wt=corrected['Wt']
+corrected_xt=corrected['xt']
+corrected_err=corrected['norm_err']
+corrected_u=corrected['uHist'][0]
+
+# ktotal=mat['ktotal'][0][0].astype(int)
+nominal_J=nominal['J']
[…]

### 2022-02-09 article/article.tex
section 2 and most part of section 3 also adds Romain's draft for intro

@@
 \begin{abstract}% Abstract of not more than 250 words.
+  \todo[write abstract]{\blindtext}
 \end{abstract}
@@
 \section{INTRODUCTION}
+\todo{Voilà ce que je propose en introduction. Il manque les références et
+peut-être un ou deux articles (j'ai laissé un endroit « libre »). Redis moi
+ce que tu en penses.
+}
+
+The cybersecurity of critical infrastructures (energy networks, industrial
+processes, water production plants, etc.) is a worldwide and sensitive
+issue. The information collected, transmitted, and processed within these
+infrastructures impact physical systems whose proper functioning is
+essential for society.
+
+Given the complexity of these structures, it is necessary to adopt a control
+structure that is distributed rather than centralized. For these systems, a
+set of controllers are designed to ensure the safe and efficient operation
+of the facilities collectively. To do this, they exchange information with
+each other, allowing them to work collaboratively. This kind of approach can
+be the case for the voltage regulation of electrical networks [REF] or the
+production-consumption balance in these same networks [REF]. Multi-agent
+systems [REF] or distributed model predictive control [REF] are standard
+tools to design the controllers.
+
+But attacking of one of these controllers, which then leaves the cooperative
+mode, can result in the violent destruction of the corrupted element. More
+subtly, it can lead to deviant behavior that is more difficult to detect. It
+aims to de-optimize the global operation or to lead the system slowly
+towards a state of non-functioning or instability [REF].
+
+With this in mind, a few articles have focused on safety in distributed
+control structures.
+
+The recent state of the art (REF Maetre 2021) highlights the complexity of
+the problems. A modification of a local goal, the hacking of communication,
+or the failure of one of the agents are events that disrupt global behavior.
+
+[QUELQUES AUTRES REFERENCES A AJOUTER]
+
+Our previous work~\citep{NogueiraEtAl2021} proposed a framework dedicated to
+sharing resource algorithms. The attack consists of communication corruption
+between a coordinator and constraint-free local agents. In this paper, we
+propose an extension of our approach by including local constraints. This
+extension profoundly changes the complexity, as the exchanges between the
+agents and the coordinator are no longer characterized by an affine function
+but a piecewise affine function. This leads not only to a combinatorial
+explosion but also to a parametric identification challenge. To overcome
+this problem, we propose using a learning method based on the Expectation
+Maximisation algorithm (REF), which allows us to estimate the corruption
+mechanism and correct it if necessary.
+
+The paper is constructed as follows ...
+
 \todo{Some citations:
@@
 }
-% % % \mpc{} \dmpc{}
 
-% % ~\\
+\section{Problem Statement}\label{sec:PS}
+In this section we present an algorithm to decompose a \mpc{} problem based on the primal decomposition \citep{BoydEtAl2015}, and a vulnerability of the decomposition that can be exploited by a malicious agent.
 
@@
 $[A,B]$ and $[A;B]$ are horizontal and vertical concatenation of matrices.
-$0_{m\times n}$ and $1_{m\times n}$ are ${m\times n}$ matrices filled with $0$ and $1$.
-$\0$ and $\1$ are 0 and 1 filled vectors of adequate size.
+${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
+$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
 $\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
@@
 $A^{\dagger}$ is the generalized inverse ${{(A^{T}A)}^{-1}A^{T}}$.
-$|A|$ is the determinant of $A$.
+$\det{A}$ is the determinant of $A$.
 A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
@@
 
-\section{Problem Statement}\label{sec:PS}
-In this section we present an algorithm to decompose a MPC problem based on the primal decomposition (\cite{BoydEtAl2015}), and a vulnerability that can be exploited by a malicious agent.
 \subsection{Model Predictive Control}\label{ssec:MPC}
@@
 \vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
-\vec{y}_{i}[k+1]&=&C_{i}\xik &&\\
-\end{matrix}
+\vec{y}_{i}[k+1]&=&C_{i}\xik &&
+\end{matrix},
 \end{equation}
+with $\vec{x}_{i}[k]:\R^{n_{x}}$, $\vec{u}_{i}[k]:\R^{n_{u}}$ and $\vec{y}_{i}[k]:\R^{n_{y}}$.
+
 Each agent's input $\uik$ is constrained by
@@
   \label{eq:u_local_constraints}
-  \0\preceq\uik\preceq\vec{u}_{\max},
+  \0_{n_{u}}\preceq\uik\preceq\vec{u}_{\max},
 \end{equation}
-where ${\umax:\R^{n_{u}\times 1}}$ and ${n_{u}=\card{\uik}}$.
-
-All agents are coupled by a global input constraint
+with ${\umax:\R^{n_{u}}}$, and
+the $M$ agents are coupled by a global input constraint with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
 \begin{equation}
   \label{eq:global_constraint}
-  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax,
+  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax.
 \end{equation}
-where \todo{$\Gamma_{i}:\semidefpos^{n_{u}}$}.
 
-The overall system can be controlled by a \mpc{}, by predicting the dynamics of the system for a finite prediction horizon $N_{p}$ and computing the optimal input by solving the following optimization problem:
+The overall system is controlled by a \mpc{}, which predicts the dynamics for a finite prediction horizon $\predhorz$ and computes the optimal input by solving the following problem:
 \begin{problem}{Global \mpc{} Problem.}\label{pb:GOP}
@@
 where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
-\todo{Reference tracking can be written as ${\vik=\wik-C_{i}\xik}$, where $\wik$ is a state reference.
-Disturbance rejection can be written as ${\vik=C_{i}\xik}$.
-}
-
-$\optglobobj$ denotes the optimal value of the objective function of problem~\ref{pb:GOP}, and ${\optuseq}$ is the optimal control sequences for all agents.
-The problem~\ref{pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in each respective $i$ subsystem, following a receding horizon strategy.
+$\optglobobj$ denotes the optimal value of the objective function for the the optimal control sequences ${\optuiseq}$.
+The problem~\ref{pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a receding horizon strategy.
 \end{problem}
 
-Due to the monolithic aspect of problem~\ref{pb:GOP}, the prediction horizon and the number of agents, solving the problem can be computationally intensive.
+Due to the monolithic aspect of problem~\ref{pb:GOP}, solving it at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive.
+Another issue that can arise is the fact that the complete information (${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$) from all the subsystems is needed to solve the problem,
+which can be viewed negatively in a confidentiality point of view.
+
 One can see that if it were not for the global constraints~\eqref{eq:global_constraint}, the problem would be easily decomposed in $M$ parts and solvable in parallel.
+So, we use the \emph{primal decomposition} method~\citep{BoydEtAl2015}, sometimes called \emph{quantity decomposition} or \emph{resource allocation}~\citep{Cohen1978}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.
 
-\subsection{Distributed Model Predictive Control}\label{ssec:dMPC}
-&continue from here
-The technique consists of decomposing the
-\todo{\dmpc{},
-}
+\subsection{Primal Decomposition based \todo{dMPC}}\label{ssec:dMPC}
+
+The technique consists of decomposing the monolithic \mpc{} problem~\ref{pb:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
 \begin{subequations}
@@
         J_{i}^{\star}(\thetaik)&=\underset{\uiseq}{\mathrm{minimize}} \obji\\
-        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:upositive}\\
-        &\quad\Gamma_{i}\uik\preceq\thetaik:\lambdaik\\
+        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
+        &\quad\Gamma_{i}\uik\preceq\qik:\dik\\
       \end{aligned}
@@
     \begin{aligned}
-      &\forall i\in \{1:\nsubsystems\}\\
-      &\forall j\in \{1:\predhorz\}
+      &\forall i\in \{1\mathbin{:}\nsubsystems\}\\
+      &\forall j\in \{1\mathbin{:}\predhorz\}
     \end{aligned}
@@
     \begin{aligned}
-      J^{\star}&=\underset{\thetaseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\thetaik)\\
-      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\thetaik\preceq\umax
+      J^{\star}&=\underset{\qiseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\qik)\\
+      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\qik\preceq\umax,\ \forall j\in\{1\mathbin{:}\predhorz\}\\
+      &\quad \qik[k]\succeq\0_{n_{u}}\text{, } \forall i\in\{1\mathbin{:}M\},\forall j\in\{1\mathbin{:}\predhorz\}
     \end{aligned}
@@
 \end{subequations}
+The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $[k]$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}.
+This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$.
+The sequence of this variables variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.
 
+The \emph{master problem}, finds the optimal allocation sequences $\optthetai$ by using a projected sub-gradient method~\citep{BoydVandenberghe2004}:
 \begin{equation}
   \label{eq:projectedSubgradient}
-\vec{\theta}^{(p+1)}=\Proj^{\set{H}}(\vec{\theta}\p-\rho\p\vec{g}\p)
+\vec{\theta}[k]\pplusone=\Proj^{\set{H}}(\vec{\theta}[k]\p-\rho\p\vec{g}[k]\p),
+\end{equation}
+where ${\set{H} = \setbuild{\vec{\theta}}{I_{c}^{M}\vec{\theta}\preceq \vec{U}_{\max}\ \&\&\ \vec{\theta}\succeq\0 }}$, being ${c=\predhorz n_{u}}$, ${I_{c}^{M}=\kron{\1_{M}}{I_{c}}}$, ${\vec{U}_{\max}=\kron{\1_{\predhorz}}{\vec{u}_{\max}}}$, $(p)$ is a given step in the iterative process and $\vec{g}\p[k]$ is a sub-gradient of the objective of the function problem in~\eqref{eq:DOP_master} in step $(p)$.
+
+From~\cite{BoydEtAl2015} it is known that $-\lambdaik$ are sub-gradients of objectives $J_{i}$ and by consequence also of objective $J[k]$.
+Plugging it in the place of $\vec{g}[k]\p$ we can find an update of the form
+\begin{equation}
+  \label{eq:thetaNegot}
+\vec{\theta}[k]\pplusone=\Proj^{\set{H}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
+\end{equation}
+which henceforth is referred as the \emph{negotiation} phase.
+
+Once the \emph{negotiation} for a given time $k$ converges, the $\optuikk$ found in the last step of the negotiation are applied in their corresponding agent and then the receding horizon strategy is followed.
+
+The algorithm to find the optimal $\optthetai[k]$ can be summarized in algorithm~\ref{alg:quantityAlg}, and the exchange between coordinator and agents can be seen in Fig.~\ref{fig:dmpc_graph}.
+
+\begin{algorithm2e}[h]
+  \DontPrintSemicolon%
+  $p:=0$\;
+  Coordinator initializes $\vec{\theta}\p$ \;
+  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
+  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
+  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
+  $p:=p+1$
+}
+ \caption{Quantity decomposition based \acrlong{dmpc}.}\label{alg:quantityAlg}
+\end{algorithm2e}
+
+One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\vec{\lambda}_{i}[k]$.
+The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the negotiation can be driven by a malicious agent that can take advantage of this trust to favor itself at the expense of others or even destabilize the negotiation, as shown in~\cite{NogueiraEtAl2021}.
+
+Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy
+\begin{equation}\label{eq:cheating}
+\lambdaicheat=\gamma_{i}(\lambdai).
 \end{equation}
 
-Where ${\set{H} = \setbuild{\vec{\theta}}{\sum_{i=1}^{M}\vec{\theta}_{i}\preceq \vec{u}_{\max}\ \&\&\ \vec{\theta}_{i}\succeq\0,i }}$.
-\todo{The sum $\sum_{i=1}^{M}\vec{\theta}_{i}$ can also be represented by the matrix multiplication $I_{c}^{M}\vec{\theta}$.
-}
+\section{Towards a safe dMPC}\label{sec:TSM}
 
+Here we focus on the local problems in~\eqref{eq:DOP_local}, and study how their structure can contribute for a safe dMPC algorithm.
+
+\subsection{Local Problems --- Formal Analysis}\label{ssec:FA}
+
+The \emph{local problems} can be rewritten in a equivalent (same solution) constrained \qp{} form:
 \begin{equation}
-  \label{eq:thetaNegot}
-\thetai\pplusone=\thetai\p+\rho\left(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p)\right)
+\begin{matrix}
+\underset{\Uik}{\mathrm{minimize}}& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
+\mathrm{s.t.}&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\\
+&\Uik\succeq\0_{c},
+\end{matrix}
+\label{eq:quadratic_case}
 \end{equation}
+where we stack the input prediction sequences in ${\Uik}$, and
+${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
+The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
+The approach depends only on the fact that $H_{i}:\semidefpos$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
 
-where ${I_{c}^{M}=\kron{\1_{M,1}}{I_{c}}}$, with ${c=N_{p}\pi_{\uik}}$.
-Observe that ${I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{v}=\frac{\sum_{i=0}^{M}\vec{v}_{i}}{M}}$.
+We choose reference tracking just to illustrate the approach. With this objective, we have
+${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
+\begin{equation}
+\small\begin{matrix*}[l]
+ H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
+\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik),
+\end{matrix*}
+\label{eq:matrices}
+\end{equation}
+where $\Wik$ stacks the setpoint prediction sequence, which is supposed to be constant during the predicted horizon ${\Wik=\kron{\1_{\predhorz}}{\vec{w}_{i}[k]}}$ (no prior information).
+${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
+${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
 
-\no{\todo[delete probably]{
-  But observe in~\eqref{eq:projectedSubgradient} that if ${\vec{\theta}_{i}\p=\0}$, ${(\vec{\lambda}_{i}^{\star} (\vec{\theta}_{i}\p)-I_{c}^{M} \pseudoinv{I_{c}^{M}}\vec{\lambda}^{\star} (\vec{\theta}\p))}$ needs to be negative so $\vec{\theta}_{i}\pplusone$ becomes negative, which is impossible \question{is it though?} since $\lambdaik\succeq\0$, from the KKT optimality conditions, seen in~\cite[Chapter 5]{BoydVandenberghe2004}.
-}}
-
+\begin{figure}[t]
 % TODO(accacio): correct image
-\begin{figure}[h]
   \centering
@@
   \end{tikzpicture}
-  \caption{Graph representation of Coordinator exchangin with $M$ agents in a \dmpc{} framework.}\label{fig:dmpc_graph}
+  \caption{Graph representation of Coordinator exchanging with $M$ agents in a \dmpc{} framework.}\label{fig:dmpc_graph}
 \end{figure}
 
-\begin{algorithm2e}[h]
-  \DontPrintSemicolon%
-  $p:=0$\;
-  Coordinator initializes $\vec{\theta}\p$ \;
-  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
-  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
-  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
-  $p:=p+1$
-}
- \caption{Quantity decomposition based \acrlong{dmpc}.}\label{alg:quantityAlg}
-\end{algorithm2e}
-
-\subsection{Local Problems --- Formal Analysis}\label{ssec:FA}
-% \todo{As seen in \cite{NogueiraEtAl2021}},
-
+As seen in~\cite{BemporadEtAl2002} and in~\cite{BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
+Similarly, we can get an explicit solution for its dual variables $\vec{\lambda}_{i}[k]$, which is a \pwa{} function with respect to $\vec{\theta}_{i}[k]$:
 \begin{equation}
-\begin{matrix}
-\underset{\Uik}{\mathrm{minimize}}& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
-\mathrm{s.t.}&\Theta_{i}\Uik\preceq\vec{\theta}_{i}:\vec{\lambda}_{i}\\
-&\Uik\succeq\0
-\end{matrix}
-\label{eq:quadratic_case}
+\resizebox{.88\columnwidth}{!}{$
+  \begin{aligned}
+    \label{eq:lambdafuntheta}
+    \lambdaik=
+    \begin{cases}
+      -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetaik \preceq z_{1}[k] \\
+      \qquad\quad \vdots&\qquad\quad \vdots\\
+      -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetaik \preceq z_{N}[k] \\
+    \end{cases}
+  \end{aligned}
+  $}
 \end{equation}
 
-If we choose reference tracking as our, for instance, we have:
-\begin{equation}
-\small\begin{matrix*}[l]
- H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
-\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik)\\
-
-\end{matrix*}
-\label{eq:matrices}
-\end{equation}
-
-We stack in vectors $\Uik$ and $\Wik$
-the input and setpoint predictions sequences ${\uiseq}$ and ${\wiseq}$.
-${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
-$\bar{Q}_{i}$, $\bar{R}_{i}$, and $\Theta_{i}$ are block diagonal matrices built repeating $\predhorz$ times $Q_{i}$, $R_{i}$, and $\Gamma_i$ respectively.
-\todo{
-  Observe that $H_{i}:\defpos$, and that $\fik$ varies with $k$.
-}
-
-\todo{As seen in~\cite{BemporadEtAl2002} and in~\cite{BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
-Similarly, we can get an explicit  solution for its dual variables $\vec{\lambda}_{i}$, which is also a \pwa{} function with respect to $\vec{\theta}_{i}$:}
-\begin{equation}
-  \begin{aligned}
-    \label{eq:lambdafuntheta}
-    \lambdai=
-    \begin{cases}
-      -P_{i}^{1}\vec{\theta}_{i}-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
-      \qquad\quad \vdots&\qquad\quad \vdots\\
-      -P_{i}^{N}\vec{\theta}_{i}-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
-    \end{cases}
-  \end{aligned}
-\end{equation}
-The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active/inactive constraints of $\Theta_{i}\Uik\preceq\vec{\theta}_{i}$ for a given time $k$.
+The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active/inactive constraints of ${\bar{\Gamma}_{i}\Uik\preceq\thetaik}$ for a given time $k$.
 
@@
  \begin{remark}\label{re:N_different_zones}
-  Observe that depending on the constraints, some of these combinations can be redundant or even make the optimization infeasible.
+  Observe that depending on the constraints, some of these combinations can be redundant or even make the optimization infeasible (empty intersection).
  \end{remark}
  \begin{assumption}
-   We assume that none of these cases in Remark~\ref{re:N_different_zones} can occur, and we have $N$ zones.
+   We assume that none of these cases in Remark~\ref{re:N_different_zones} can occur, thus we have $N$ zones.
  \end{assumption}
 
-\todo{
-  The values of $P_{i}^{j}$ and $\vec{s}_{i}^{j}$, depend on $H_{i}$, $\fik$ and parts of $\Theta_{i}$ which represents the active constraints.
-  If we take for example when all constraints are active, we have ${P_{i}^{1}={(\Theta_{i}H_{i}^{-1}\Theta_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{1}[k]=P_{i}^{1}\Theta_{i}H_{i}^{-1}\fik}$.
-  And if all constraints are inactive, ${P_{i}^{N}=0_{\card{\Theta_{i}}\times\card{\Theta_{i}}}}$ and ${\vec{s}_{i}^{N}=\0}$.
-  Observe that in all functions, only $\vec{s}_{i}^{j}[k]$ depend on time $k$.
-}
+  $P_{i}^{j}$ and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$, depending on the active constraints.
+  For example, if we take the case when all constraints are active, we have ${P_{i}^{1}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{1}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
+  And if all constraints are inactive, ${P_{i}^{N}=0_{c\times c}}$ and ${\vec{s}_{i}^{N}=\0_{c}}$.\\
+  Observe that in all components of the function in~\eqref{eq:lambdafuntheta}, only $\vec{s}_{i}^{j}[k]$ depend on time $k$.
 
-\subsection{Attack in DMPC scheme}\label{ssec:attack}
+\subsection{The Attack}\label{ssec:attack}
+We suppose the malicious agent choose a function $\gamma_{i}(\cdot)$ to use throughout a given time $k$.
+\begin{assumption}
+  $\gamma_{i}(\cdot)$ does not change during the negotiation phase for a given time $k$.
+\end{assumption}
 
-\begin{equation}\label{eq:cheating}
-\lambdaicheat=\gamma_{i}(\lambdai)
-\end{equation}
-
-% \subsection{Detection and mitigation}
-\begin{assumption}
-  $\gamma_{i}(\cdot)$ is the same during the negotiation phase for a given time $k$ (it does not depend on $p$).
+In~\cite{NogueiraEtAl2021}, we have seen that one interpretation of $\lambdai$ is the dissatisfaction of an agent with the resource allocated for it, with $\lambdai=\0_{c}$ meaning total satisfaction.
+\begin{assumption}\label{ass:gamma_zero}
+  We suppose $\gamma(\cdot)=0\iff\lambdai=\0_{c}$
 \end{assumption}
@@
 \begin{assumption}
-  equation
-  T inversivel
-  We suppose the agent chooses a linear function such as \[ {\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i}} \] resulting in:
+  We suppose the agent chooses a linear function such as \[ {\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i}}, \]
+  where $\determinant{T}\neq 0$ (from assumption~\ref{ass:gamma_zero}).
+\end{assumption}
+
+Such attack function results in a \pwa{} solution for $\lambdaicheat[k]$:
 \begin{equation}
+  \resizebox{.88\columnwidth}{!}{$
     \label{eq:linear_cheating}
-    \tilde{\lambdai}=
+    \lambdaicheat[k]=
     \begin{cases}
@@
       \qquad\quad \vdots&\qquad\quad \vdots\\
-      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k] \\
+      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k],
     \end{cases}
+    $}
 \end{equation}
-and we define $\tilde{P}_{i}^{j}[k]=\Tik P_{i}^{j}$ and $\tilde{\vec{s}}_{i}^{j}[k]=\Tik\vec{s}_{i}^{j}[k]$.
+where we define $\tilde{P}_{i}^{j}[k]=\Tik P_{i}^{j}$ and $\tilde{\vec{s}}_{i}^{j}[k]=\Tik\vec{s}_{i}^{j}[k]$.
+Observe that as the function is applied only in $\lambdaik$ it does not affect the hyperplanes generating the zones.
+
+\todo{
+  Since $\bar{\Gamma}_{i}$ are block diagonal we can assure\pdfnote{proof?} there exists one zone where all constraints are active.
+}
+\begin{assumption}
+  We suppose that for every time $k$ there exists a zone for each agent $i$ where all constraints are active.
 \end{assumption}
-\todo{
-  Since $\Theta_{i}$ are diagonal we can assure\pdfnote{proof?} there exists one \question{zone?}zone where all constraints are active. We will fix $j=1$ for this zone and called it the $1$-zones.
-  We can estimate the values for this partition using:
-}
+
+Fixing the index for the zones where all constraints are active to $1$, we can called them the $1$-zones.
+We can estimate the parameters of their respective solution in~\eqref{eq:linear_cheating} with
 \begin{equation}
   \label{eq:lambdafuntheta_tilde}
-\tilde{\lambdai}=\gamma_{i}(\lambdai(\thetai))=-\widehat{\tilde{P}_{i}^{1}}[k]\vec{\theta}_{i}-\widehat{\tilde{\vec{s}}_{i}^{1}}[k]
+\lambdaicheat[k]=\gamma_{i}(\lambdai(\thetaik))=-\widehat{\tilde{P}_{i}^{1}}[k]\thetaik-\widehat{\tilde{\vec{s}}_{i}^{1}}[k].
 \end{equation}
@@
   \begin{cases}
-    -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{with probability}\ \pi_{1} \\
+    -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{if in zone } 1\\
     \qquad\quad \vdots&\qquad\quad \vdots\\
-    -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{with probability}\ \pi_{Z} \\
+    -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{if in zone } Z\\
   \end{cases}.

### 2022-02-10 article/article.tex
changes after meeting with Romain

@@
 \begin{abstract}% Abstract of not more than 250 words.
-  \todo[write abstract]{\blindtext}
+  \todo[write abstract]{
+  In distributed predictive control structures, communication among agents is required to achieve a consensus and approach an optimal global behavior.
+Such negotiation mechanisms are sensitive to attacks on these exchanges.
+  This paper proposes a monitoring scheme that detects and mitigates these attacks' effects in a resource allocation framework.
+  The performance of the proposed method is illustrated through simulations of the temperature control of multiple rooms under input power constraints.
+}
 \end{abstract}
@@
 \section{INTRODUCTION}
-\todo{Voilà ce que je propose en introduction. Il manque les références et
-peut-être un ou deux articles (j'ai laissé un endroit « libre »). Redis moi
-ce que tu en penses.
-}
+The information collected, processed, and transmitted within \cps{} impacts physical systems whose proper functioning is essential for society, like energy networks, industrial processes, water production plants, etc..
+Attacks in those critical infrastructures, such as Stuxnet, increased awareness for cyber-security of \cps{}.
 
-The cybersecurity of critical infrastructures (energy networks, industrial
-processes, water production plants, etc.) is a worldwide and sensitive
-issue. The information collected, transmitted, and processed within these
-infrastructures impact physical systems whose proper functioning is
-essential for society.
+Given the complexity, scale and even geographic location of these structures, it is necessary to adopt control structures that are decentralized rather than \todo{monolithic}.
+For these systems, a set of controllers are designed to ensure the safe and efficient operation of the facilities collectively.
 
-Given the complexity of these structures, it is necessary to adopt a control
-structure that is distributed rather than centralized. For these systems, a
-set of controllers are designed to ensure the safe and efficient operation
-of the facilities collectively. To do this, they exchange information with
-each other, allowing them to work collaboratively. This kind of approach can
-be the case for the voltage regulation of electrical networks [REF] or the
-production-consumption balance in these same networks [REF]. Multi-agent
-systems [REF] or distributed model predictive control [REF] are standard
-tools to design the controllers.
+This operation is done cooperatively, where the controllers exchange information with each other.
+This approach can be used, for example, for the voltage regulation of electrical networks~\citep{HansEtAl2019}, and the production-consumption balance in these same networks~\citep{HerreraEtAl2015}.
+Multi-agent systems~\citep{KantamneniEtAl2015} or \dmpc{}~\citep{MaestreEtAl2014} are standard tools to design these controllers.
 
-But attacking of one of these controllers, which then leaves the cooperative
-mode, can result in the violent destruction of the corrupted element. More
-subtly, it can lead to deviant behavior that is more difficult to detect. It
-aims to de-optimize the global operation or to lead the system slowly
-towards a state of non-functioning or instability [REF].
+When one of these controllers is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
+Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to \todo{slowly} lead the system towards a state of malfunction/instability~\todo[use other]{\citep{NogueiraEtAl2021}}.
 
-With this in mind, a few articles have focused on safety in distributed
-control structures.
+With this in mind, a few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as lagrange-based decomposition~\citep{VelardeEtAl2017}, and Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018}.
 
-The recent state of the art (REF Maetre 2021) highlights the complexity of
-the problems. A modification of a local goal, the hacking of communication,
+The authors in~\cite{VelardeEtAl2017} highlight the complexity of the problems. A modification of a local goal, the hacking of communication,
 or the failure of one of the agents are events that disrupt global behavior.
 
-[QUELQUES AUTRES REFERENCES A AJOUTER]
+Some strategies are presented to mitigate the effects of such event,
+for instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
+using secure scenarios based on reliable historical data~\cite{VelardeEtAl2017a}.
+The most recent~\cite{MaestreEtAl2021} presents ways to generate these scenarios.
 
 Our previous work~\citep{NogueiraEtAl2021} proposed a framework dedicated to
-sharing resource algorithms. The attack consists of communication corruption
-between a coordinator and constraint-free local agents. In this paper, we
-propose an extension of our approach by including local constraints. This
-extension profoundly changes the complexity, as the exchanges between the
-agents and the coordinator are no longer characterized by an affine function
-but a piecewise affine function. This leads not only to a combinatorial
-explosion but also to a parametric identification challenge. To overcome
-this problem, we propose using a learning method based on the Expectation
-Maximisation algorithm (REF), which allows us to estimate the corruption
+sharing resource algorithms (primal decomposition), and shown its vulnerabilities.
+More precisely, \fdi{} in the communication
+between a coordinator and local constraint-free agents bound by global equality constraints. The paper also proposed a safe algorithm based on data reconstruction to mitigate the effects.
+
+In this paper, we propose an extension of our approach by including local constraints and changing the global constraints.
+This extension profoundly changes the complexity, as the exchanges between the
+agents and the coordinator are no longer characterized by an affine functions
+but by \pwa{} functions.
+This leads not only to a combinatorial explosion but also to a parametric identification challenge.
+To overcome this problem, we propose using a learning method based on the \EM{} algorithm~\citep{DempsterEtAl1977}, which allows us to estimate the corruption
 mechanism and correct it if necessary.
 
-The paper is constructed as follows ...
-
-\todo{Some citations:
-  \cite{MaestreEtAl2021}
-
-}
-
-\section{Problem Statement}\label{sec:PS}
-In this section we present an algorithm to decompose a \mpc{} problem based on the primal decomposition \citep{BoydEtAl2015}, and a vulnerability of the decomposition that can be exploited by a malicious agent.
+\todo{The paper is constructed as follows}
 
@@
 $\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto set $\set{T}$.
-$\card{x}$ is the number of elements in $x$.
+% $\card{x}$ is the number of elements in $x$.
 $\kron{}{}$, and $\hadamard{}{}$ are the Kronecker and Hadamard products.
@@
 $A^{\dagger}$ is the generalized inverse ${{(A^{T}A)}^{-1}A^{T}}$.
-$\det{A}$ is the determinant of $A$.
+$\determinant{A}$ is the determinant of $A$.
 A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
@@
 $\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
-$\pdf{\cdot}$ is a probability density function. ${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.
+% $\pdf{\cdot}$ is a probability density function.
+${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.
+
+\section{Problem Statement}\label{sec:PS}
+In this section we present an algorithm to decompose a \mpc{} problem based on the primal decomposition \citep{BoydEtAl2015}, and a vulnerability of the decomposition that can be exploited by a malicious agent.
 
 \subsection{Model Predictive Control}\label{ssec:MPC}
-Let us consider a system composed of $M$ discrete-time linear time-invariant agents.
-The dynamics of an agent $i$ is ruled by:
+Let us consider a system composed of $M$ discrete-time linear time-invariant agents, whose dynamics are ruled by:
 \begin{equation}
@@
 \end{equation}
-with ${\umax:\R^{n_{u}}}$, and
-the $M$ agents are coupled by a global input constraint with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
+with ${\umax:\R^{n_{u}}}$.
+The $M$ agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
 \begin{equation}
@@
 The overall system is controlled by a \mpc{}, which predicts the dynamics for a finite prediction horizon $\predhorz$ and computes the optimal input by solving the following problem:
-\begin{problem}{Global \mpc{} Problem.}\label{pb:GOP}
-\begin{equation*}
+\begin{equation}\label{eq:GOP}
+\resizebox{.88\columnwidth}{!}{$
 \begin{matrix}
@@
   &\forall j\in \{1:\predhorz\}
-\end{aligned}\right.
+\end{aligned}\right.,
 
 \end{matrix}
+  $}
 % \label{eq:dmpcModLOP}
-\end{equation*}
+\end{equation}
 where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
 $\optglobobj$ denotes the optimal value of the objective function for the the optimal control sequences ${\optuiseq}$.
-The problem~\ref{pb:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a receding horizon strategy.
-\end{problem}
+The problem in~\eqref{eq:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a \rhs{}.
 
-Due to the monolithic aspect of problem~\ref{pb:GOP}, solving it at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive.
+Due to the monolithic aspect of~\eqref{eq:GOP}, solving it at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive.
 Another issue that can arise is the fact that the complete information (${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$) from all the subsystems is needed to solve the problem,
@@
 One can see that if it were not for the global constraints~\eqref{eq:global_constraint}, the problem would be easily decomposed in $M$ parts and solvable in parallel.
-So, we use the \emph{primal decomposition} method~\citep{BoydEtAl2015}, sometimes called \emph{quantity decomposition} or \emph{resource allocation}~\citep{Cohen1978}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.
+So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.
 
-\subsection{Primal Decomposition based \todo{dMPC}}\label{ssec:dMPC}
+\subsection{Primal Decomposition based dMPC}\label{ssec:dMPC}
 
-The technique consists of decomposing the monolithic \mpc{} problem~\ref{pb:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
+The technique consists of decomposing the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
 \begin{subequations}
@@
 
-The \emph{master problem}, finds the optimal allocation sequences $\optthetai$ by using a projected sub-gradient method~\citep{BoydVandenberghe2004}:
+The \emph{master problem} finds the optimal allocation sequences $\optthetai$ by using a projected sub-gradient method:
 \begin{equation}
@@
 
-From~\cite{BoydEtAl2015} it is known that $-\lambdaik$ are sub-gradients of objectives $J_{i}$ and by consequence also of objective $J[k]$.
-Plugging it in the place of $\vec{g}[k]\p$ we can find an update of the form
+From~\cite{BoydEtAl2015}, it is known that $-\lambdaik$ are sub-gradients of objective $J[k]$.
+Plugging it in $\vec{g}[k]\p$ we have
 \begin{equation}
@@
 \end{equation}
-which henceforth is referred as the \emph{negotiation} phase.
+\newcommand{\negotiation}{\emph{negotiation}}
+which henceforth is referred as the \negotiation.
 
-Once the \emph{negotiation} for a given time $k$ converges, the $\optuikk$ found in the last step of the negotiation are applied in their corresponding agent and then the receding horizon strategy is followed.
+Once the \negotiation{} for a given time $k$ converges, the $\optuikk$ found in the last step of the \negotiation{} are applied in their corresponding agent and then the \rhs{} is followed.
 
@@
 
-One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\vec{\lambda}_{i}[k]$.
-The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the negotiation can be driven by a malicious agent that can take advantage of this trust to favor itself at the expense of others or even destabilize the negotiation, as shown in~\cite{NogueiraEtAl2021}.
+One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
+An interpretation of $\lambdaik$ is the dissatisfaction of an agent with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
+The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent that can take advantage of this trust to favor itself at the expense of others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.
 
-Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy
-\begin{equation}\label{eq:cheating}
+Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy of the form
+\begin{equation*}
 \lambdaicheat=\gamma_{i}(\lambdai).
-\end{equation}
+\end{equation*}
 
@@
 The \emph{local problems} can be rewritten in a equivalent (same solution) constrained \qp{} form:
-\begin{equation}
+\begin{equation}\label{eq:quadratic_case}
 \begin{matrix}
@@
 \end{matrix}
-\label{eq:quadratic_case}
 \end{equation}
-where we stack the input prediction sequences in ${\Uik}$, and
+where ${\Uik}$ stacks the input prediction sequences, and
 ${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
@@
 
+\no{
   We choose reference tracking just to illustrate the approach. With this objective, we have
@@
 ${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
+}
 
@@
 
-As seen in~\cite{BemporadEtAl2002} and in~\cite{BorrelliEtAl2017}, the solution for a constrained linear quadratic regulation problem yields in a \pwa{} function.
-Similarly, we can get an explicit solution for its dual variables $\vec{\lambda}_{i}[k]$, which is a \pwa{} function with respect to $\vec{\theta}_{i}[k]$:
+As seen in~\cite{BemporadEtAl2002}, the solution for a linear constrained \qp{} problem yields in a \pwa{} function.
+Similarly, we can get an explicit solution for its dual variables $\vec{\lambda}_{i}[k]$, which is a \pwa{} function w.r.t. $\vec{\theta}_{i}[k]$:
+% \begin{equation}
+%   \resizebox{.88\columnwidth}{!}{$
+%     \begin{aligned}
+%       \label{eq:lambdafuntheta}
+%       \lambdaik=
+%       \begin{cases}
+%         -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G^{1}[k]\thetaik \preceq b^{1}[k] \\
+%         \qquad\quad \vdots&\qquad\quad \vdots\\
+%         -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G^{N}[k]\thetaik \preceq b^{N}[k] \\
+%       \end{cases}
+%     \end{aligned}
+%     $}
+% \end{equation}
 \begin{equation}
@@
     \lambdaik=
-    \begin{cases}
-      -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetaik \preceq z_{1}[k] \\
-      \qquad\quad \vdots&\qquad\quad \vdots\\
-      -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetaik \preceq z_{N}[k] \\
-    \end{cases}
+    % \begin{cases}
+      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
+    % \end{cases}
   \end{aligned}
@@
 \end{equation}
+with $n\in\{1\mathbin{:}N\}$.
 
-The halfspaces defined by the set of pairs $(G_{j}[k],z_{j}[k])$ represent a combination of active/inactive constraints of ${\bar{\Gamma}_{i}\Uik\preceq\thetaik}$ for a given time $k$.
+The halfspaces defined by the set of pairs $(G^{j}[k],b^{j}[k])$ represent a combination of active/inactive constraints of ${\bar{\Gamma}_{i}\Uik\preceq\thetaik}$ for a given time $k$, and vary w.r.t.
+$\vec{v}_{i}[k]$.
+
+The $P_{i}^{j}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$, and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$.
+The specific construction depends on the active constraints.
+For example, if we take the case when all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
+And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.\\
+\begin{remark}\label{rmk:P_constant}
+The $\vec{s}_{i}^{j}[k]$ depend on time $k$, while $P_{i}^{j}$ do not.
+\end{remark}
 
 \begin{challenge}\label{ch:partition_unknown}
-  It is important to note that since the actual state is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
+  It is important to note that since $\vec{v}_{i}$ is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
 \end{challenge}
@@
 
- As each constraint can be active or inactive, for a group of $n$ constraints, we can have at most $N=2^{n}$ different combinations of active/inactive sets and consequently $N$ regions (or zones).
- \begin{remark}\label{re:N_different_zones}
-  Observe that depending on the constraints, some of these combinations can be redundant or even make the optimization infeasible (empty intersection).
- \end{remark}
+ As each constraint can be active or inactive, for a given group of $n$ constraints, we can have at most $N=2^{n}$ different combinations of active/inactive sets and consequently $N$ regions (zones).
+
  \begin{assumption}
-   We assume that none of these cases in Remark~\ref{re:N_different_zones} can occur, thus we have $N$ zones.
+   We assume that none of the combinations of active/inactive constraints are redundant (no linear dependency), nor make the optimization infeasible (no empty intersection).
+   Thus we always have $N$ zones.
  \end{assumption}
-
-  $P_{i}^{j}$ and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$, depending on the active constraints.
-  For example, if we take the case when all constraints are active, we have ${P_{i}^{1}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{1}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
-  And if all constraints are inactive, ${P_{i}^{N}=0_{c\times c}}$ and ${\vec{s}_{i}^{N}=\0_{c}}$.\\
-  Observe that in all components of the function in~\eqref{eq:lambdafuntheta}, only $\vec{s}_{i}^{j}[k]$ depend on time $k$.
 
@@
 \begin{assumption}
-  $\gamma_{i}(\cdot)$ does not change during the negotiation phase for a given time $k$.
+  $\gamma_{i}(\cdot)$ does not change during the \negotiation{} phase for a given time $k$.
 \end{assumption}
 
-In~\cite{NogueiraEtAl2021}, we have seen that one interpretation of $\lambdai$ is the dissatisfaction of an agent with the resource allocated for it, with $\lambdai=\0_{c}$ meaning total satisfaction.
 \begin{assumption}\label{ass:gamma_zero}
-  We suppose $\gamma(\cdot)=0\iff\lambdai=\0_{c}$
+  We suppose $\gamma(\cdot)=\0_{c}\iff\lambdai=\0_{c}$
 \end{assumption}
@@
 \begin{assumption}
-  We suppose the agent chooses a linear function such as \[ {\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i}}, \]
+  The agent chooses a linear function
+  \begin{equation}
+\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i},
+\end{equation}
   where $\determinant{T}\neq 0$ (from assumption~\ref{ass:gamma_zero}).
@@
 Such attack function results in a \pwa{} solution for $\lambdaicheat[k]$:
+% \begin{equation}
+%   \resizebox{.88\columnwidth}{!}{$
+%     \label{eq:linear_cheating}
+%     \lambdaicheat[k]=
+%     \begin{cases}
+%       -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq b_{1}[k] \\
+%       \qquad\quad \vdots&\qquad\quad \vdots\\
+%       -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq b_{N}[k],
+%     \end{cases}
+%     $}
+% \end{equation}
 \begin{equation}
@@
     \lambdaicheat[k]=
-    \begin{cases}
-      -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq z_{1}[k] \\
-      \qquad\quad \vdots&\qquad\quad \vdots\\
-      -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq z_{N}[k],
-    \end{cases}
+    % \begin{cases}
+      -\tilde{P}_{i}^{n}[k]\thetaik-\tilde{\vec{s}}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
+    % \end{cases}
     $}
 \end{equation}
-where we define $\tilde{P}_{i}^{j}[k]=\Tik P_{i}^{j}$ and $\tilde{\vec{s}}_{i}^{j}[k]=\Tik\vec{s}_{i}^{j}[k]$.
+with $n\in\{1\mathbin{:}N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
 Observe that as the function is applied only in $\lambdaik$ it does not affect the hyperplanes generating the zones.
 
-\todo{
-  Since $\bar{\Gamma}_{i}$ are block diagonal we can assure\pdfnote{proof?} there exists one zone where all constraints are active.
-}
-\begin{assumption}
-  We suppose that for every time $k$ there exists a zone for each agent $i$ where all constraints are active.
-\end{assumption}
+We fix the index for these zones where all constraints are active to $1$, and we can call them the $1$-zones.
 
-Fixing the index for the zones where all constraints are active to $1$, we can called them the $1$-zones.
-We can estimate the parameters of their respective solution in~\eqref{eq:linear_cheating} with
+\subsection{Detection and mitigation}\label{ssec:DM}
+Supposing we have a sequence of $\thetaik$ in a zone $j$, we can estimate the parameters for these zones with
 \begin{equation}
   \label{eq:lambdafuntheta_tilde}
-\lambdaicheat[k]=\gamma_{i}(\lambdai(\thetaik))=-\widehat{\tilde{P}_{i}^{1}}[k]\thetaik-\widehat{\tilde{\vec{s}}_{i}^{1}}[k].
+\lambdaicheat[k]=\gamma_{i}(\lambdai(\thetaik))=-\widehat{\tilde{P}_{i}^{j}}[k]\thetaik-\widehat{\tilde{\vec{s}}_{i}^{j}}[k].
 \end{equation}
@@
 \begin{assumption}\label{ass:Pnominal}
-  The nominal value of $P_{i}^{1}$, denoted $\bar{P}_{i}^{1}$, is available from reliable attack-free historical data.
+  The nominal value of $P_{i}^{j}$ for this given zone $j$, denoted $\bar{P}_{i}^{j}$, is available from reliable attack-free historical data.
 \end{assumption}
 
-\todo{
-  Using this strategy, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{1}}[k]-\bar{P}_{i}^{1}\|_{F}}$, where ${\|\cdot\|_{F}}$ is the Frobenius norm.
+Since $P_{i}^{n}$ do not change w.r.t.\ time, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{j}}[k]-\bar{P}_{i}^{j}\|_{F}}$.
 Let ${d_{i}\in\{0,1\}}$ be an indicator that detects the attack in agent $i$.
 If the disturbance $E_{i}[k]$ respects an arbitrary bound
-}
 \begin{equation}
@@
 \end{equation}
-\todo{
 then  ${d_{i}=0}$, and no attack is detected. Otherwise, ${d_{i}=1}$, and a change in behavior of agent $i$ is detected.
-}
 
-\todo{
-  Using these assumptions, we can try to estimate the inverse of $T_{i}(k)$ as in
-}
-\begin{equation}\label{eq:5}
-\widehat{{T_{i}(k)}^{-1}}=\bar{P}_{i}^{1}{\widehat{\tilde{P}_{i}^{1}}[k]}^{-1},
-\end{equation}
-
-and from~\eqref{eq:lambdafuntheta}, we can derive a method to reconstruct $\vec{\lambda}_{i}$:
+Using all assumptions presented, we can estimate the inverse of $T_{i}(k)$ with
+\begin{equation*}
+\widehat{{T_{i}(k)}^{-1}}=\bar{P}_{i}^{j}{\widehat{\tilde{P}_{i}^{j}}[k]}^{-1},
+\end{equation*}
+if $\widehat{\tilde{P}_{i}^{j}}[k]$ is invertible, and $\widehat{\tilde{P}_{i}^{j}}[k]$ is only invertible when all constraints are active, that means, when ${j=1}$.
+\begin{assumption}
+  We suppose there exists a $1$-zone for each agent $i$ for every time $k$.
+\end{assumption}
+And from~\eqref{eq:lambdafuntheta}, we can derive a method to reconstruct $\vec{\lambda}_{i}$:
 \begin{equation}
@@
 \end{equation}
-\subsection{Estimating}
-\todo{
+
+Since we don't know how to generate points in the $1$-zones, we use the \EM{} algorithm which can potentially identify the parameters of all modes.
+
+\subsection{Expectation Maximization}
 In order to estimate $\widehat{\tilde{P}_{i}^{1}}$, we must have enough observed points in the $1$-zones.
 We propose to generate points surrounding arbitrary $\bar{\vec{\theta}_{i}}$ in the $1$-zones, and use them to estimate $\widehat{\tilde{P}_{i}^{1}}$.
-Now we have to find such points.
-}
 
@@
 
-\section{Expectation-Maximization Algorithm}\label{sec:expect-maxim-algo}
-
 The main objective of the \EM{} algorithm, seen in~\cite{DempsterEtAl1977}, is to find maximum likelihood estimators for models with latent variables.
@@
 For an observation ${o\in\set{O}=\{1\mathbin{:}O\}}$, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$.
-As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a expansion of the model referred as \emph{mixture of switching regressions} in~\cite{QuandtRamsey1978} and \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010}, which we will call \emph{mixture of affine regressions}, since our regressors have a linear term (matrices $\tilde{P}^{z}$) and a constant term (vectors $\tilde{\vec{s}}^{z}$):
+As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use an expansion of the model referred as \emph{mixture of switching regressions} in~\cite{QuandtRamsey1978} and \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010}, which we will call \emph{mixture of affine regressions}, since our regressors have a linear term (matrices $\tilde{P}^{z}$) and a constant term (vectors $\tilde{\vec{s}}^{z}$):
 % TODO(accacio): change subscript n
@@
 theta_0 and lambda_0).
-
 \begin{equation}
@@
 
-Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use in our mitigation scheme proposed in \S\ref{sec:attack}.
+Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use in our mitigation scheme proposed in \S\ref{ssec:DM}.
 
@@
 \todo{Simulated annealing as in \cite{OzerovFevotte2010}}
+
+\subsection{Safe dMPC Algorithm}\label{ssec:safe_algo}
+\SetKwBlock{negotPhase}{ Negotiation Phase:}{}
+\SetKwBlock{detectPhase}{ Detection Phase:}{}
+\begin{algorithm2e}[h]
+  \DontPrintSemicolon
+  \detectPhase{
+    Coordinator sends sequence of $\vec{\theta}_{i}^{o}$, $o\in\set{O}$ \;
+    Subsystems solve~\eqref{eq:DOP_local}, and send $\tilde{\vec{\lambda}}^{o}_{i}$, $o\in\set{O}$\;
+    Coordinator estimates $\widehat{\tilde{P}_{i}^{1}}[k]$ and $\widehat{\tilde{\vec{s}}_{i}^{1}}[k]$ with \EM{}\;
+    Coordinator computes $d_{i}$ using~\eqref{eq:2}\;
+  }
+  \negotPhase{
+  Coordinator initializes $\vec{\theta}^{(0)}$ \;
+  $p:=0$\;
+
+  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
+  Subsystems solve~\eqref{eq:DOP_local}, and send $\vec{\lambda}^{\star}_{i}(\vec{\theta}\p)$\;
+
+  Coordinator updates allocation~\eqref{eq:thetaNegot} using adequate versions of $\vec{\lambda}_{i}\p$ for each agent:
+  $\tilde{\vec{\lambda}}_{i}\p$, if $d_{i}=0$ and ${\vec{\lambda}_{i}}_{\mathrm{rec}}$, if ${d_{i}=1}$\;
+  $p:=p+1$
+ }}
+ \caption{Secure DMPC.}\label{alg:safeDMPC}
+\end{algorithm2e}
 

### 2022-02-11 code/update_parameters.m
change variables names to reflect article

@@ -1,12 +1,13 @@
-function [Phi, pi_new] = update_parameters(X, Y, Responsabilities)
+function [Phi, pi_new] = update_parameters(Theta, Lambda, Responsabilities)
 % UPDATE_PARAMETERS -
     modes=size(Responsabilities,1);
-    [n N]=size(X);
+    [n N]=size(Theta);
 
-    A=kron(ones(n,1),eye(n));
-    B=kron(eye(N),ones(1,n));
-    C=kron(repmat(eye(n),1,N),ones(n,1));
-    theta=sparse([(A*X*B).*C; kron(ones(1,N),eye(n))]');
+    Upsilon=kron(ones(n,1),eye(n));
+    Delta=kron(eye(N),ones(1,n));
+    G=kron(ones(1,N),eye(n));
+    Y=kron(G,ones(n,1));
+    Omega=sparse([(Upsilon*Theta*Delta).*Y; G]');
 
     N_k=sum(Responsabilities,2);
     pi_new(:)=N_k/N;
@@ -15,7 +16,7 @@ function [Phi, pi_new] = update_parameters(X, Y, Responsabilities)
         resp2=cellfun(@(x) x*eye(n),mat2cell(responsabilities',ones(1,N)),'UniformOutput',0);
         Gamma=sqrt(sparse(blkdiag(resp2{:})));
 
-        Phi(i,:)=-((Gamma*theta)\(Gamma*Y(:)));
+        Phi(i,:)=-((Gamma*Omega)\(Gamma*Lambda(:)));
     end
 
 end

### 2022-02-11 article/article.tex
version with 6 pages, Abstr/Intro/Concl TODO

@@
 Such negotiation mechanisms are sensitive to attacks on these exchanges.
-  This paper proposes a monitoring scheme that detects and mitigates these attacks' effects in a resource allocation framework.
-  The performance of the proposed method is illustrated through simulations of the temperature control of multiple rooms under input power constraints.
+  This paper proposes a monitoring scheme based on Expectation Maximization to detect and mitigate the effects of attacks in a resource allocation framework.
+  The performance is assessed through academical example of the temperature control of multiple rooms under input power constraints.
 }
@@
 This operation is done cooperatively, where the controllers exchange information with each other.
-This approach can be used, for example, for the voltage regulation of electrical networks~\citep{HansEtAl2019}, and the production-consumption balance in these same networks~\citep{HerreraEtAl2015}.
+This approach can be used, for example, for the voltage regulation of electrical networks, and the production-consumption balance in these same networks.
 Multi-agent systems~\citep{KantamneniEtAl2015} or \dmpc{}~\citep{MaestreEtAl2014} are standard tools to design these controllers.
 
-When one of these controllers is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
-Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to \todo{slowly} lead the system towards a state of malfunction/instability~\todo[use other]{\citep{NogueiraEtAl2021}}.
+A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018} and primal decomposition~\citep{NogueiraEtAl2021}.
 
-With this in mind, a few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as lagrange-based decomposition~\citep{VelardeEtAl2017}, and Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018}.
+The authors in these works highlight the complexity of the problems.
+A modification of a local goal, the hacking of communication, or the failure of one of the agents are events that disrupt global behavior.
+If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
+% Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to lead the system towards a state of malfunction/instability~\todo[use other]{}.
 
-The authors in~\cite{VelardeEtAl2017} highlight the complexity of the problems. A modification of a local goal, the hacking of communication,
-or the failure of one of the agents are events that disrupt global behavior.
+Some strategies are presented to mitigate the effects of such event.
+For instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
+using secure scenarios based on reliable historical data~\cite{MaestreEtAl2021}.
 
-Some strategies are presented to mitigate the effects of such event,
-for instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
-using secure scenarios based on reliable historical data~\cite{VelardeEtAl2017a}.
-The most recent~\cite{MaestreEtAl2021} presents ways to generate these scenarios.
-
-Our previous work~\citep{NogueiraEtAl2021} proposed a framework dedicated to
-sharing resource algorithms (primal decomposition), and shown its vulnerabilities.
-More precisely, \fdi{} in the communication
-between a coordinator and local constraint-free agents bound by global equality constraints. The paper also proposed a safe algorithm based on data reconstruction to mitigate the effects.
+Our previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.
 
@@
 
-\todo{The paper is constructed as follows}
+\todo{
+This paper is organized as follows.
+First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based DMPC and its vulnerabilities.
+Then in Section~\ref{sec:TSM} we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
+Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with academic simulation and we assess its performance.
+Finally, in Section~\ref{sec:CC}, we conclude, and we give an outlook of future works.
+}
 
@@
 In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
-$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto set $\set{T}$.
+$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
 % $\card{x}$ is the number of elements in $x$.
-$\kron{}{}$, and $\hadamard{}{}$ are the Kronecker and Hadamard products.
+$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
 $\vectorize{A}$ vectorizes matrix $A$.
 ${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
-$[A,B]$ and $[A;B]$ are horizontal and vertical concatenation of matrices.
+$[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
 ${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
@@
 $\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
-$\preceq$ is the component-wise less or equal for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X'=X}}$.
+$\preceq$ is the component-wise inequality for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X\T=X}}$.
 ${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
 $I_{c}$ is the identity matrix of size $c$.
-$A^{\dagger}$ is the generalized inverse ${{(A^{T}A)}^{-1}A^{T}}$.
+$A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}$.
 $\determinant{A}$ is the determinant of $A$.
@@
 \subsection{Model Predictive Control}\label{ssec:MPC}
-Let us consider a system composed of $M$ discrete-time linear time-invariant agents, whose dynamics are ruled by:
+Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\mathbin{:}\nsubsystems\}}$, modeled by
 \begin{equation}
@@
 with ${\umax:\R^{n_{u}}}$.
-The $M$ agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
+The agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
 \begin{equation}
   \label{eq:global_constraint}
-  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax.
+  \sum_{i\in\set{M}}\Gamma_{i}\uik\preceq\umax.
 \end{equation}
 
-The overall system is controlled by a \mpc{}, which predicts the dynamics for a finite prediction horizon $\predhorz$ and computes the optimal input by solving the following problem:
+The overall system is controlled by a \mpc{}, which
+computes the optimal input for a finite prediction horizon ${\set{H}=\{1\mathbin{:}\predhorz\}}$ by solving the following problem:
 \begin{equation}\label{eq:GOP}
@@
 \begin{matrix}
-\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
+\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits_{i\in\set{M}} \overbrace{\sum_{j\in\set{H}}\norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
 \mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
@@
 \begin{aligned}
-  &\forall i\in \{1:\nsubsystems\}\\
-  &\forall j\in \{1:\predhorz\}
+  &\forall i\in\set{M}\\
+  &\forall j\in\set{H}
 \end{aligned}\right.,
@@
 
-Due to the monolithic aspect of~\eqref{eq:GOP}, solving it at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive.
-Another issue that can arise is the fact that the complete information (${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$) from all the subsystems is needed to solve the problem,
+Solving~\eqref{eq:GOP} at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive.
+Another issue that can arise is the fact that the complete information \todo[verify]{(${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$)} from all the subsystems is needed to solve the problem,
 which can be viewed negatively in a confidentiality point of view.
 
-One can see that if it were not for the global constraints~\eqref{eq:global_constraint}, the problem would be easily decomposed in $M$ parts and solvable in parallel.
 So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.
@@
       \small
-      \begin{aligned}
-        J_{i}^{\star}(\thetaik)&=\underset{\uiseq}{\mathrm{minimize}} \obji\\
-        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
-        &\quad\Gamma_{i}\uik\preceq\qik:\dik\\
-      \end{aligned}
+        J_{i}^{\star}(\thetaik)=
+        \begin{matrix}
+        \underset{\uiseq}{\mathrm{minimize}}&\obji\\
+        \mathrm{subject~ to} &\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
+        &\Gamma_{i}\uik\preceq\qik:\dik\\
+      \end{matrix}
     \right\}
@@
     \begin{aligned}
-      &\forall i\in \{1\mathbin{:}\nsubsystems\}\\
-      &\forall j\in \{1\mathbin{:}\predhorz\}
+      &\forall i\in\set{M}\\
+      &\forall j\in\set{H}
     \end{aligned}
@@
   \end{equation}
+
   \begin{equation}
@@
     \begin{aligned}
-      J^{\star}&=\underset{\qiseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\qik)\\
-      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\qik\preceq\umax,\ \forall j\in\{1\mathbin{:}\predhorz\}\\
-      &\quad \qik[k]\succeq\0_{n_{u}}\text{, } \forall i\in\{1\mathbin{:}M\},\forall j\in\{1\mathbin{:}\predhorz\}
+      J^{\star}=
+      \begin{matrix}
+        \underset{\qiseq}{\mathrm{minimize}} &\sum_{i\in\set{M}} J^{\star}_i(\qik)\\
+        \mathrm{subject~ to} &
+        \left.
+          \begin{aligned}
+            \quad \sum_{i\in\set{M}}\qik\preceq\umax\\
+            \quad \qik[k]\succeq\0_{n_{u}},       \forall i\in\set{M}
+          \end{aligned}
+        \right\}
+        \begin{aligned}
+          \forall j\in\set{H}
+        \end{aligned}
+      \end{matrix}
     \end{aligned}
@@
 \end{subequations}
-The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $[k]$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}.
+The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}.
 This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$.
-The sequence of this variables variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.
+The sequence of this variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.
 
@@
   \label{eq:projectedSubgradient}
-\vec{\theta}[k]\pplusone=\Proj^{\set{H}}(\vec{\theta}[k]\p-\rho\p\vec{g}[k]\p),
+\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p-\rho\p\vec{g}[k]\p),
 \end{equation}
-where ${\set{H} = \setbuild{\vec{\theta}}{I_{c}^{M}\vec{\theta}\preceq \vec{U}_{\max}\ \&\&\ \vec{\theta}\succeq\0 }}$, being ${c=\predhorz n_{u}}$, ${I_{c}^{M}=\kron{\1_{M}}{I_{c}}}$, ${\vec{U}_{\max}=\kron{\1_{\predhorz}}{\vec{u}_{\max}}}$, $(p)$ is a given step in the iterative process and $\vec{g}\p[k]$ is a sub-gradient of the objective of the function problem in~\eqref{eq:DOP_master} in step $(p)$.
+where ${\set{S} = \setbuild{\vec{\theta}}{I_{c}^{M}\vec{\theta}\preceq \vec{U}_{\max}\ \&\ \vec{\theta}\succeq\0 }}$, being ${c=\predhorz n_{u}}$, ${I_{c}^{M}=\kron{\1_{M}}{I_{c}}}$, ${\vec{U}_{\max}=\kron{\1_{\predhorz}}{\vec{u}_{\max}}}$, $(p)$ is a given step in the iterative process and $\vec{g}\p[k]$ is a sub-gradient of the objective function of problem in~\eqref{eq:DOP_master} in step $(p)$.
 
@@
   \label{eq:thetaNegot}
-\vec{\theta}[k]\pplusone=\Proj^{\set{H}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
+\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
 \end{equation}
@@
 }
- \caption{Quantity decomposition based \acrlong{dmpc}.}\label{alg:quantityAlg}
+ \caption{Quantity decomposition based \acrshort{dmpc}.}\label{alg:quantityAlg}
 \end{algorithm2e}
@@
 One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
-An interpretation of $\lambdaik$ is the dissatisfaction of an agent with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
-The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent that can take advantage of this trust to favor itself at the expense of others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.
+An interpretation of $\lambdaik$ is the dissatisfaction of agent $i$ with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
+The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent which can take advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.
 
-Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy of the form
+Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy
 \begin{equation*}
@@
 
-The \emph{local problems} can be rewritten in a equivalent (same solution) constrained \qp{} form:
-\begin{equation}\label{eq:quadratic_case}
-\begin{matrix}
-\underset{\Uik}{\mathrm{minimize}}& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
-\mathrm{s.t.}&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\\
-&\Uik\succeq\0_{c},
-\end{matrix}
-\end{equation}
+The \emph{local problems}~\eqref{eq:DOP_local} can be rewritten in a equivalent (same solution) constrained \qp{} form:
+\todo[Correct alignment maybe]{
+  \begin{subequations}\label{eq:quadratic_case}
+    \begin{align}
+      \label{eq:4}
+      \underset{\Uik}{\mathrm{minimize}}&& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
+      \mathrm{s.t.}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
+                                        &&\Uik\succeq\0_{c},
+    \end{align}
+  \end{subequations}
+}
 where ${\Uik}$ stacks the input prediction sequences, and
@@
 }
-
+\todo[make a better figure if time permits]{
 \begin{figure}[t]
-% TODO(accacio): correct image
   \centering
-  \begin{tikzpicture}[node distance=1.7cm]
+  \begin{tikzpicture}[node distance=1.5cm]
     \tikzset{
       agent/.style={draw,circle,minimum width=1cm,},
-      coord/.style={draw,circle,minimum width=1.7cm,very thick}
+      coord/.style={draw,circle,minimum width=1cm,very thick}
     }
@@
 
-    \node[agent] (a1) {A$_{1}$};
+    \node[agent] (a1) {a$_{1}$};
     \node (a2) [right of=a1]  {$\dots$};
-    \node[agent] (am) [right of=a2] {A$_{M}$};
-    \node[coord] (c) [above of=a2] {Coord};
+    \node[agent] (am) [right of=a2] {a$_{M}$};
+    \node[coord] (c) [above of=a2] {C};
 
-    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}$} (c) ;
-    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}$} (a1);
+    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}[k]$} (c) ;
+    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}[k]$} (a1);
 
-    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}$} (c);
-    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}$} (am);
+    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}[k]$} (c);
+    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}[k]$} (am);
 
   \end{tikzpicture}
-  \caption{Graph representation of Coordinator exchanging with $M$ agents in a \dmpc{} framework.}\label{fig:dmpc_graph}
+  \caption{Exchange between coordinator (C) and agents $a_{i}$.}\label{fig:dmpc_graph}
 \end{figure}
+}
 
-As seen in~\cite{BemporadEtAl2002}, the solution for a linear constrained \qp{} problem yields in a \pwa{} function.
-Similarly, we can get an explicit solution for its dual variables $\vec{\lambda}_{i}[k]$, which is a \pwa{} function w.r.t. $\vec{\theta}_{i}[k]$:
+We can get an explicit solution for the dual variables $\vec{\lambda}_{i}[k]$, which are \acrfull{pwa} functions w.r.t. $\vec{\theta}_{i}[k]$:
 % \begin{equation}
@@
 \resizebox{.88\columnwidth}{!}{$
-  \begin{aligned}
-    \label{eq:lambdafuntheta}
+  \begin{aligned}\label{eq:lambdafuntheta}
     \lambdaik=
@@
 
-The halfspaces defined by the set of pairs $(G^{j}[k],b^{j}[k])$ represent a combination of active/inactive constraints of ${\bar{\Gamma}_{i}\Uik\preceq\thetaik}$ for a given time $k$, and vary w.r.t.
-$\vec{v}_{i}[k]$.
+The halfspaces defined by the set of pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.
 
-The $P_{i}^{j}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$, and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$.
-The specific construction depends on the active constraints.
-For example, if we take the case when all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
-And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.\\
+The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$.
+The specific construction depends on the active constraints in~\eqref{eq:constraints}.
+For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
+And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.
+
 \begin{remark}\label{rmk:P_constant}
@@
 
-\begin{challenge}\label{ch:partition_unknown}
+\begin{challenge}\label{ch:zone_unknown}
   It is important to note that since $\vec{v}_{i}$ is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
 \end{challenge}
-\begin{challenge}\label{ch:zone_unknown}
+\begin{challenge}\label{ch:s_unknown}
  For the same reason, the values of $\vec{s}_{i}[k]$ are also unknown.
@@
 
- As each constraint can be active or inactive, for a given group of $n$ constraints, we can have at most $N=2^{n}$ different combinations of active/inactive sets and consequently $N$ regions (zones).
+ As each constraint can be active or inactive, for a given group of $t$ constraints, we can have at most $N=2^{t}$ different combinations of active sets.
 
@@
 \subsection{The Attack}\label{ssec:attack}
-We suppose the malicious agent choose a function $\gamma_{i}(\cdot)$ to use throughout a given time $k$.
+We suppose the malicious agent chooses a function $\gamma_{i}(\cdot)$ to use throughout a given time $k$.
 \begin{assumption}
   $\gamma_{i}(\cdot)$ does not change during the \negotiation{} phase for a given time $k$.
-\end{assumption}
-
-\begin{assumption}\label{ass:gamma_zero}
-  We suppose $\gamma(\cdot)=\0_{c}\iff\lambdai=\0_{c}$
 \end{assumption}
@@
 \end{equation}
-  where $\determinant{T}\neq 0$ (from assumption~\ref{ass:gamma_zero}).
+  where $\determinant{T}\neq 0$.
 \end{assumption}
@@
 with $n\in\{1\mathbin{:}N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
-Observe that as the function is applied only in $\lambdaik$ it does not affect the hyperplanes generating the zones.
+Observe that as the function is applied only in $\lambdaik$, it does not affect the hyperplanes generating the zones.
 
@@
 \subsection{Detection and mitigation}\label{ssec:DM}
-Supposing we have a sequence of $\thetaik$ in a zone $j$, we can estimate the parameters for these zones with
+Supposing we have a sequence of $\thetaik$ in a zone $j$, we can estimate the parameters for this zone with
 \begin{equation}
@@
 Since $P_{i}^{n}$ do not change w.r.t.\ time, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{j}}[k]-\bar{P}_{i}^{j}\|_{F}}$.
-Let ${d_{i}\in\{0,1\}}$ be an indicator that detects the attack in agent $i$.
-If the disturbance $E_{i}[k]$ respects an arbitrary bound
+Let ${d_{i}}$ be an indicator
 \begin{equation}
   \label{eq:2}
-  E_{i}[k]\leq\epsilon_{P},
+  d_{i}=\indicator{E_{i}[k]\geq\epsilon_{P}},
 \end{equation}
-then  ${d_{i}=0}$, and no attack is detected. Otherwise, ${d_{i}=1}$, and a change in behavior of agent $i$ is detected.
+which detects the attack in agent $i$
+if the disturbance $E_{i}[k]$ disrespects an arbitrary bound $\epsilon_{P}$.
 
-Using all assumptions presented, we can estimate the inverse of $T_{i}(k)$ with
-\begin{equation*}
+If an attack is detected, we can estimate the inverse of $T_{i}(k)$ with
+\begin{equation}
 \widehat{{T_{i}(k)}^{-1}}=\bar{P}_{i}^{j}{\widehat{\tilde{P}_{i}^{j}}[k]}^{-1},
-\end{equation*}
+\end{equation}
 if $\widehat{\tilde{P}_{i}^{j}}[k]$ is invertible, and $\widehat{\tilde{P}_{i}^{j}}[k]$ is only invertible when all constraints are active, that means, when ${j=1}$.
@@
 \begin{equation}
-  \label{eq:lambdareconstruction}
+  \label{eq:lambda_reconstruction}
   {\vec{\lambda}_{i}}_{\mathrm{rec}}=\widehat{{T_{i}[k]}^{-1}} \tilde{\vec{\lambda}_{i}}
@@
 
-Since we don't know how to generate points in the $1$-zones, we use the \EM{} algorithm which can potentially identify the parameters of all modes.
+In order to estimate $\widehat{\tilde{P}_{i}^{1}}[k]$, we must have enough observed points in the $1$-zones, so
+we propose to generate points surrounding arbitrary $\bar{\vec{\theta}_{i}}$ in the $1$-zones.
 
+\begin{assumption}\label{ass:}
+  Given the unconstrained solution of~\ref{eq:quadratic_case} ${ {\optuncUik}=-H_{i}^{-1}\fik}$, we suppose
+$\bar{\Gamma}_{i}\optuncUik\succeq\0_{c}$ for all $k$, that means, the points $\thetai=\0_{c}$ are in the $1$-zones.
+\end{assumption}
+
+\glsreset{EM}
+Unfortunately, since we don't know the hyperplanes that separates the different zones, we don't know how close of $\bar{\theta}_{i}=\0_{c}$ we need to generate our points.
+So we generate points arbitrarily close to $\bar{\theta}_{i}$ and we use the \EM{} algorithm which can potentially identify the parameters of all $N$ modes.
 \subsection{Expectation Maximization}
-In order to estimate $\widehat{\tilde{P}_{i}^{1}}$, we must have enough observed points in the $1$-zones.
-We propose to generate points surrounding arbitrary $\bar{\vec{\theta}_{i}}$ in the $1$-zones, and use them to estimate $\widehat{\tilde{P}_{i}^{1}}$.
+As the method is the same for all agents and repeated each time $k$, we drop the subscript $i$ and the time dependency $[k]$ to simplify the notation.
 
-% TODO(accacio): change unconstrained notation
-\todo{
-  Let $\optuncUik$ be the solution of the unconstrained version of~\ref{eq:quadratic_case}.
-  It is known that its solution is ${ {\optuncUik}=-H_{i}^{-1}\fik}$.
-  For the constrained solution we know that the constraints will be inactive if they respect the inequalities $\Theta_{i}\optuncUik\preceq\thetai$.
-  If they regard the inequalities, the constrained value is calculated as is, if not it should be projected onto the viable region.
-  This way, for all constraints to be active, all inequalities should be disrespected in $\Theta_{i}\optuncUik\preceq\thetai$, thus \[\thetai\prec\Theta_{i}\optuncUik.\]
-  % Since $H_{i}:\defpos$, then $H_{i}^{-1}:\defpos$,
-  If we choose $\bar{\theta}_{i}=\0$, the constraints will be active when the  will only  that can assure that all constraints are active, then we can estimate ${\widehat{\tilde{P}_{i}^{1}}[k]}$ around this point.
-Unfortunately, since we don't know the hyperplanes that separates the different zones, some points generated for the estimation can belong to other zones.
-  So we use the \EM{} algorithm to estimate the parameter for the $1$-zones, as it can clusterize the points and estimate parameters at the same time.
+Given a set of parameters $\set{P}$, a set of observable data $\set{B}$ and a set of non-observable data $\set{U}$, the main objective of the \EM{} algorithm is to find estimators of $\set{P}$ which maximize the log marginal likelihood of the observed data ${\ln\probability{\set{B};\set{P}}}$, for models with latent variables in $\set{U}$.
+
+Since maximizing ${\ln\probability{\set{B};\set{P}}}$ does not have an analytical solution, the algorithm solves the optimization problem in a iterative manner.
+So, rather than finding the $\set{P}$ that maximizes ${\ln\probability{\set{B};\set{P}}}$, we find the $\set{P}$ that maximizes the expectation of the complete data log-likelihood $\ln\probability{\set{B},\set{U};\set{P}}$ w.r.t.
+the posterior probabilities ${\probability{\set{U}|\set{B};\set{P}}}$
+calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$.
+This provides a convergence-guaranteed iterative algorithm ensuring at each iteration a monotonic increase of the log-marginal likelihood.
+
+The algorithm can be divided into 2 steps presented in algorithm~\ref{alg:em}.
+
+\SetKwBlock{Estep}{ E step:}{}
+\SetKwBlock{Mstep}{ M step:}{}
+\begin{algorithm2e}[h]
+  \DontPrintSemicolon%
+  Initialize parameters $\set{P}_{\mathrm{new}}$\;
+  \Repeat{$\set{P}_{\mathrm{cur}}$ converge}{
+    $\set{P}_{\mathrm{cur}}=\set{P}_{\mathrm{new}}$\;
+    \Estep{
+      Evaluate ${\probability{\set{U}|\set{B};\set{P}_{\mathrm{cur}}}}$\;
     }
+    \Mstep{
+      Reestimate parameters using:
+      \begin{equation}
+      \resizebox{.68\columnwidth}{!}{$
+        \label{eq:m_step}
+        \set{P}_{\mathrm{new}}=\arg\underset{\set{P}}{\max}\ \expectation[{\probability{\set{U}|\set{B};\set{P}_{\mathrm{cur}}}}]{\ln\probability{\set{B},\set{U};\set{P}}}
+        $}
+      \end{equation}
+    }
+}
+ \caption{Expectation Maximization}\label{alg:em}
+\end{algorithm2e}
 
-The main objective of the \EM{} algorithm, seen in~\cite{DempsterEtAl1977}, is to find maximum likelihood estimators for models with latent variables.
+For a group of $O$ exchanges between an agent and the coordinator, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$, ${o\in\set{O}=\{1\mathbin{:}O\}}$, which can be organized in ${\random{\Theta},\random{\Lambda}\in\R^{c\times O}}$.
+${(\random{\Theta},\random{\Lambda})}$ is our set of observable data $\set{B}$.
 
-We will use our estimation problem to illustrate how the algorithm works, and
-we will concentrate on the estimation of the parameters for a single agent in a single step $k$, so we drop the subscript $i$ and the time dependency $[k]$ to simplify the notation.
+As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a multidimensional expansion of the model referred as \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010} to map the relation between $\random{\Theta}$ and $\random{\Lambda}$.
+We will call the model \emph{mixture of affine regressions}, since our regressors have linear and constant terms:
+% TODO(accacio): change subscript n
+% \begin{equation}\label{eq:linear_cheating_random}
+%   \randomvec{\lambda}_{o}=
+%   \begin{cases}
+%     -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{if in zone } 1\\
+%     \qquad\quad \vdots&\qquad\quad \vdots\\
+%     -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{if in zone } Z\\
+%   \end{cases}.
+% \end{equation}
+\begin{equation}\label{eq:linear_cheating_random}
+  \randomvec{\lambda}_{o}=
+    -\tilde{P}^{z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{z}\text{, if in zone } z,\\
+\end{equation}
+with  ${z\in\set{Z}=\{1\mathbin{:}Z\}}$.
 
-As in~\eqref{eq:lambdafuntheta},  any response variable $\tilde{\vec{\lambda}}$, is a function of an input $\vec{\theta}$, which belongs to a unknown zone ${\set{Z}=\{1\mathbin{:}Z\}}$, Challenge~\ref{ch:zone_unknown}.
-The relationship between the input and the response is given by a set of parameters we want to estimate ${\set{P}=\setbuild{(\tilde{P}^{z},\tilde{\vec{s}}^{z})}{z\in\set{Z}}}$.
+As \eqref{eq:linear_cheating_random} is a \pwa{} function whose modes depend on the unknown zone ${z\in\set{Z}}$,
+we associate to each couple $(    \randomvec{\lambda}_{o}, \randomvec{\theta}_{o})$ a latent unobserved random variable ${\random{z}_{o}}$ that indicates in which zone in $\set{Z}$ the observable variables were obtained.
+These variables are organized in ${\random{Z}:\R^{1\times O}}$ which is our set of latent variables $\set{U}$.
 
-For an observation ${o\in\set{O}=\{1\mathbin{:}O\}}$, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$.
-As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use an expansion of the model referred as \emph{mixture of switching regressions} in~\cite{QuandtRamsey1978} and \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010}, which we will call \emph{mixture of affine regressions}, since our regressors have a linear term (matrices $\tilde{P}^{z}$) and a constant term (vectors $\tilde{\vec{s}}^{z}$):
-% TODO(accacio): change subscript n
-\begin{equation}
-  \label{eq:linear_cheating_random}
-  \randomvec{\lambda}_{o}=
-  \begin{cases}
-    -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{if in zone } 1\\
-    \qquad\quad \vdots&\qquad\quad \vdots\\
-    -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{if in zone } Z\\
-  \end{cases}.
-\end{equation}
+The latent variable $\random{z}_{o}$ follows a categorical prior distribution, with associated probabilities ${\Pi=\{\pi^{z}|z\in\set{Z}\}}$ such
+\[\probability{\random{z}_{o}=z}=\pi^{z} \in [0,1], \qquad \sum_{z\in\set{Z}} \pi^{z} = 1.
+\]
+As parameters to estimate, we have ${\set{P}=\setbuild{\set{P}^{z}}{z\in\set{Z}}}$, with ${\set{P}^{z}=(\tilde{P}^{z},\tilde{\vec{s}}^{z},\pi^{z})}$.
 
-Each couple of observed input and response variables $(    \randomvec{\lambda}_{o}, \randomvec{\theta}_{o})$ is associated in this model to a latent unobserved random variable ${\random{z}_{o}\in\set{Z}}$ that indicates from which of the $Z$ affine regression models in~\eqref{eq:linear_cheating_random} the response variable was obtained.
-The latent variable $\random{z}_{o}$ follows a categorical prior distribution, with associated probabilities ${\Pi=\{\pi_{1},\dots,\pi_{Z}\}}$:
-   \begin{equation}
-     \label{eq:prob_zo_equal_z}
-\probability{\random{z}_{o}=z}=\pi_{z} \in [0,1], \qquad \sum_{z=1}^{Z} \pi_{z} = 1.
-\end{equation}
-Since $\vec{\theta}$ is our input,
-we consider a non-informative improper probability density function~\cite{ChristensenEtAl2010}
-% SL: BTW, maybe we should use different notations for "the probability that some discrete random variable (e.g.
-z_0) is equal to something, and the probability density function of a continuous random variable (e.g.
-theta_0 and lambda_0).
-\begin{equation}
-  \label{eq:theta_almost_surely}
+Since $\vec{\theta}$ is our input, we consider a non-informative improper probability density function~\citep{ChristensenEtAl2010}
+\[
   \probability{\randomvec{\theta}_{o}} \,\propto \,1.
-\end{equation}
+\]
 
@@
 where, following~\eqref{eq:linear_cheating_random}, the mean vector is defined by
-\[f(\randomvec{\theta}_{o};(P,\vec{s}))=-{P}\randomvec{\theta}_{o}-{\vec{s}},\]
+\({f(\randomvec{\theta}_{o};(P,\vec{s},\pi))=-{P}\randomvec{\theta}_{o}-{\vec{s}},}\)
 and the covariance matrix ${\Sigma^{z}}$ tends to $0$.
 
-This mixture of affine regression model is represented as a probabilistic graphical model in Fig.~\ref{fig:model}, and it corresponds to the following factorization of the complete-data (i.e., observed and latent data) likelihood, \cite{Bishop2006}:
+\no{
+This mixture of affine regression model corresponds to the following factorization of the complete-data (i.e., observed and latent data) likelihood, \cite{Bishop2006}:
 \begin{align}\label{eq:completedataLikelihood}
-  \probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}= \prod_{o=1}^{O}\prod_{z=1}^{Z}\big[&\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z;\set{P}^{z}} \nonumber \\
+  \probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}= \prod_{\set{O}}\prod_{\set{Z}}\big[&\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z;\set{P}^{z}} \nonumber \\
 & \times \probability{\random{z}_{o}=z}\probability{\randomvec{\theta}_{o}}\big]^{\indicator{\random{z}_{o}=z}},
 \end{align}
+}
 
-Ideally, we would like to estimate the unknown model parameters $\set{P}$ by maximizing the log-marginal likelihood $\ln \probability{\random{\Theta},\random{\Lambda};\set{P}}$, computed by marginalizing the latent variables $\random{Z}$ in~\eqref{eq:completedataLikelihood}.
-Unfortunately, this optimization problem does not admit a closed-form analytical solution.
-Instead, we can exploit the latent-variable structure of the model to derive an \EM{} algorithm,~\cite{DempsterEtAl1977}, which is a convergence-guaranteed iterative algorithm ensuring at each iteration a monotonic increase of the log-marginal likelihood.
-
-\begin{figure}[b]
-  \centering
-  \begin{tikzpicture}
-    \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
-    \node at (1.2,-1.3) {O};
-    \node (pi) at (-2.5,.5) {$\Pi$};
-    \node (phi) at (2.5,-1.) {$\set{P}$};
-
-    \graph [edge quotes={fill=white,inner sep=1pt},
-    clockwise=3,nodes={circle,draw,rotate=-60,minimum width=1cm}] {
-      a/"${\randomvec{\theta}_{o}}$"[rotate=60,fill=lightgray],b/"${\random{\vec{\lambda}}_{o}}$"[rotate=60,fill=lightgray],c/"${\random{z}}_{o}$"[rotate=60];
-      {a,c} ->[-latex] b;
-      (pi) ->[thick,{Circle[length=3.pt]}-latex] c;
-      (phi) ->[thick,{Circle[length=3.pt]}-latex] b;
-      };
-  \end{tikzpicture}
-  \caption{Graph representation of model proposed.}\label{fig:model}
-\end{figure}
-
-Now we can formalize the \EM{} problem.
-\begin{problem}{Expectation Maximization Problem}\label{pb:EM}
-
-  Given a set of observed data $(\random{\Theta},\random{\Lambda})$, estimate the unknown latent variables $\random{Z}$ and parameter set $\set{P}$ by iteratively maximizing \todo{a function of} the complete-data log likelihood $\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}$.
-\end{problem}
-
-Using~\eqref{eq:prob_zo_equal_z},~\eqref{eq:theta_almost_surely}, and~\eqref{eq:multivariate_gaussian} in~\eqref{eq:completedataLikelihood} yields {\color{blue}the following expression of the complete-data log-likelihood:}
-\begin{equation}\label{eq:completedataLogLikelihood_complete}
-\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}=  \sum_{o=1}^{O}\sum_{z=1}^{Z}{\indicator{\random{z}_{o}=z}}
-  \alpha_{zo},
-\end{equation}
+We can calculate the posterior probabilities  $\zeta_{zo}(\set{P})=\probability{\random{z}_{o}=z|\randomvec{\lambda}_{o},\randomvec{\theta}_{o};\set{P}}$, also called \emph{responsibilities}:
+\begin{align}
+  \label{eq:responsibilities}
+\zeta_{zo}(\set{P})&=\frac{\pi_{z}{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}{\sum\limits_{j=1}^{Z}\pi_{j}\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{j}),{\Sigma^{j}})},
+\end{align}
+and then calculate the expectation of ${\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}$ with respect to ${\zeta_{zo}(\set{P}_{\mathrm{cur}})}$ \citep[Chapter 9]{Bishop2006}
+\begin{align}
+  \label{eq:completedataLogLikelihood_expectation}
+\expectation[{\zeta_{zo}(\set{P}_{\mathrm{cur}})}]{\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}&= \sum_{o\in\set{O}}\sum_{z\in\set{Z}}  \zeta_{zo}(\set{P}_{\mathrm{cur}})\alpha_{zo},
+\end{align}
 where ${\alpha_{zo}=\ln{\pi_{z}}+\ln{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}$.
 
-But as mentioned, we do not observe the complete data $(\random{\Theta},\random{\Lambda},\random{Z})$, instead we observe only $(\random{\Theta},\random{\Lambda})$ and all the information on $\random{Z}$ given these observations is carried by the posterior probabilities  ${\zeta(z_{zo};\set{P})=\probability{\random{z}_{o}=z|\randomvec{\lambda}_{o},\randomvec{\theta}_{o};\set{P}}}$, also called \emph{responsibilities}, which can be calculated as
-\todo[Change Notation?]{Change Notation?}
-% SL: the notation z_{zo} is a bit confusing and I am not sure it was defined.
-It's actually related to the fact that you use z_0 as the latent variable and $z$ as the values it can take in \mathcal{Z}.
-Using another variable for the latter may help clarifying.
+\begin{remark}
+The $\Sigma^{z}$ used in~\eqref{eq:responsibilities} are updated using a technique called \emph{Simulated annealing}~\citep{OzerovFevotte2010}, where they are initialized with arbitrarily big values indicating the uncertainty of the parameters and it is reduced as the estimations converge.
+\end{remark}
 
-\begin{align}
-  \label{eq:responsibilities}
-\zeta(z_{zo};\set{P})&=\frac{\probability{\random{z}_{o}=z}\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o};\set{P}^{z}}}{\sum\limits_{j=1}^{Z}\probability{\random{z}_{o}=j}\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o};\set{P}^{j}}}\nonumber\\
-  &=\frac{\pi_{z}{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}{\sum\limits_{j=1}^{Z}\pi_{j}\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{j}),{\Sigma^{j}})}.
-\end{align}
-
-Observe that by taking
-\todo[Change Notation?]{Change Notation?}
-\begin{equation}\label{eq:argmaxz}
-\arg\underset{z}{\max}\ {\zeta_{zo}(;\set{P})},
-\end{equation}
-we can get the most probable $z$-zone which generated the observation $\randomvec{\lambda}_{o}$.
-
-So, rather than using the complete data log-likelihood~\eqref{eq:completedataLogLikelihood_complete}, which we do not have, since the latent variables are not observed by definition, the \EM{} algorithm uses its expectation with respect to the posterior calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$:
-\begin{align}
-  \label{eq:completedataLogLikelihood_expectation}
- Q\left(\set{P},\set{P}_{\mathrm{cur}}\right) &= \expectation[{\zeta(z_{zo};\set{P}_{\mathrm{cur}})}]{\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}
-\end{align}
-and since \[\expectation[{\zeta(z_{zo};\set{P}_{\mathrm{cur}})}]{\indicator{\random{z}_{o}=z}}=\zeta(z_{zo};\set{P}_{\mathrm{cur}}),\]
-we can rewrite~\eqref{eq:completedataLogLikelihood_expectation}:
-\begin{align}
-  \label{eq:completedataLogLikelihood_expectation_developed}
-Q\left(\set{P},\set{P}_{\mathrm{cur}}\right) &= \sum_{o=1}^{O}\sum_{z=1}^{Z}  \zeta(z_{zo};\set{P}_{\mathrm{cur}})\alpha_{zo},
-\end{align}
-\todo{\begin{remark}
-For the first iteration, the model parameters must be initialized.
-A discussion about the initialization of the parameters is beyond the scope of this article, and we refer the reader to~\cite{Bishop2006},~\cite{DempsterEtAl1977} and their references.
-\end{remark}}
-
-Then we can find a new estimate of $\set{P}$ that maximizes $Q(\set{P},\set{P}_{\mathrm{cur}})$:
-\begin{equation} \label{eq:Mstep}
-  \set{P}_{\mathrm{new}}=\underset{\set{P}}{\argmax}\ Q(\set{P},\set{P}^{\mathrm{cur}}).
-\end{equation}
-For the next iteration of the algorithm, this new estimate $\set{P}_{\mathrm{new}}$ will be used as $\set{P}^{\mathrm{cur}}$.
-
-With this information we can describe the \EM{} algorithm in two parts, first calculate the responsibilities $\zeta(z_{zo};\set{P}_{\mathrm{cur}})$, and then update the parameters.
-Algorithm~\ref{alg:em} adapted from~\mbox{\cite[Chapter 9]{Bishop2006}} describes the steps.
-
-\begin{algorithm2e}[h]
-  \DontPrintSemicolon%
-  Initialize parameters $\set{P}_{\mathrm{new}}$\;
-  \Repeat{$\set{P}_{\mathrm{cur}}$ converge}{
-    $\set{P}_{\mathrm{cur}}=\set{P}_{\mathrm{new}}$\;
-  \textbf{E step} Evaluate responsibilities~\eqref{eq:responsibilities}\;
-  \textbf{M step} Reestimate parameters~\eqref{eq:Mstep}\;
-}
- \caption{Expectation Maximization}\label{alg:em}
-\end{algorithm2e}
-
-Here we introduce a variable ${\vec{\phi}^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$ so we can calculate the new estimates of $\set{P}$ from~\eqref{eq:Mstep}.
-Using the KKT conditions~\cite{BoydVandenberghe2004}, we can find an optimal solution for the problem in~\eqref{eq:Mstep}, by
+Here we introduce a variable ${\vec{\phi}^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$.
+Using the KKT conditions, we can find an optimal solution for the problem in~\eqref{eq:m_step}, by
 taking the gradients of~\eqref{eq:completedataLogLikelihood_expectation} with respect to vectors $\vec{\phi}^{z}$ and making them vanish.
 Because of the multidimensional nature of the problem, some matrix operations are needed to synthesize the results.
-After those operations, we have a matricial solution that yields the optimal estimates $\vec{\phi}_{z}^{\mathrm{new}}$:
+After those operations, we have a matricial solution that yields the optimal estimates $\vec{\phi}^{z}_{\mathrm{new}}$:
 \begin{equation}
   \label{eq:mstepestimation}
-  \vec{\phi}_{z}^{\mathrm{new}}=\pseudoinv{(\Xi_{z}\random{\Omega})}\Xi_{z}\vectorize{\random{\Lambda}},
+  \vec{\phi}^{z}_{\mathrm{new}}=\pseudoinv{(\Xi^{z}\random{\Omega})}\Xi^{z}\vectorize{\random{\Lambda}},
 \end{equation}
@@
 with matrices
-${\Upsilon=\kron{\1_{n}^{T}}{I_{n}}}$,
-${\Delta=\kron{I_{N}}{\1_{n}^{T}}}$,
-${Y=\kron{G}{\1_{n}}}$,
-${G=\kron{\1_{N}^{T}}{I_{n}}}$,
+${\Upsilon=\kron{\1_{c}^{T}}{I_{c}}}$,
+${\Delta=\kron{I_{O}}{\1_{c}^{T}}}$,
+${Y=\kron{G}{\1_{c}}}$,
+${G=\kron{\1_{O}^{T}}{I_{c}}}$,
 and
-\[{\Xi_{z}={\diag(\sqrt{{\zeta(z_{z1};\set{P}_{\mathrm{cur}})}}I_{n},\cdots,\sqrt{{\zeta(z_{zO};\set{P}_{\mathrm{cur}})}}I_{n})}}.\]
-% SL: I find it surprising that you have square roots of the responsabilities in this equation. Are you sure it is correct?
+\[{\Xi^{z}={\diag(\sqrt{{\zeta(z_{z1};\set{P}_{\mathrm{cur}})}}I_{c},\cdots,\sqrt{{\zeta(z_{zO};\set{P}_{\mathrm{cur}})}}I_{c})}}.\]
+Doing the same for $\pi^{z}$ we get
+\begin{equation}
+  \label{eq:3}
+  \pi^{z}=\sum\limits_{o\in\set{O}}\tfrac{\zeta_{zo}(\set{P}_{\mathrm{cur}})}{O}.
+\end{equation}
 
@@
 Instead of affecting the observed data to a zone with 100\% of certainty (\emph{hard assignment}), \EM{} uses the responsibilities (\emph{soft assignment}) for each zone.
-Then, when we solve the \textbf{M Step}, rather than using only the data assigned to update the parameters of a certain zone, we take into account all data, whose contribution to the update is weighted by their associated responsibilities.
 
 Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use in our mitigation scheme proposed in \S\ref{ssec:DM}.
+Observe that not necessarily the $z$ indices correspond to the indices $n$ in~\eqref{eq:lambdafuntheta}.
+So, to find the $z$ that corresponds to ${n=1}$, we take the observation $o$ for $\bar{\theta}=\0_{c}$, which we know belongs to the $1$-zone, and we find the most probable $z$ with
+\begin{equation}\label{eq:argmaxz}
+\arg\underset{z}{\max}\ {\zeta_{zo}(\set{P})}.
+\end{equation}
 
-\todo {
-  Observe that not necessarily the parameters estimated associated to a value z corresponds to the z shown in
-}
-% \begin{equation*}
-%   p(\randomvec{x})=\sum_{i=0}^{N}\pi_{i} \mathcal{N}(\random{y};f_{i}(\randomvec{x}),\Sigma_{i})
-% \end{equation*}
-
-% \begin{equation}
-%   \mathcal{N}(\random{y}; {\mu}_{i},\Sigma_{i})=\frac{1}{{\sqrt{2\pi\sigma_{i}^{2}}}}e^{-\frac{(\random{y}-{\mu}_{i})^{2}}{2\sigma^{2}}}
-% \end{equation}
-
-% for a $D-$dimensional gaussian we have:
-% \begin{equation}
-%   \mathcal{N}(\randomvec{y}; \vec{\mu}_{i},\Sigma_{i})=\frac{1}{{{(2\pi)}^{D/2}}{|\Sigma_{i}|^{\frac{1}{2}}}}e^{-\frac{1}{2}\norm{\randomvec{y}-\vec{\mu}_{i}}^{2}_{\Sigma_{i}^{-1}}}
-% \end{equation}
-
-% \begin{figure}[b]
-%   \centering
-%   \scriptsize \def\svgwidth{0.49\textwidth}
-%   \includegraphics[width=\columnwidth]{pwagaussian.pdf}
-%   \caption{Example of Gaussian Probability density functions used in mixture for identification of a 2D Piecewise Affine function with 2 modes, where \todo[verify colormap]{black represents 0 probability and white maximum }}\label{fig:pwagaussian}
-% \end{figure}
-\todo{Simulated annealing as in \cite{OzerovFevotte2010}}
+\begin{remark}
+A discussion about the initialization and update of other parameters is beyond the scope of this article, so we refer the reader to~\cite{Bishop2006}, and~\cite{DempsterEtAl1977}.
+\end{remark}
 
 \subsection{Safe dMPC Algorithm}\label{ssec:safe_algo}
+Integrating the EM to our detection and mitigation mechanism we can propose in Algorithm~\ref{alg:safeDMPC} a secure \dmpc{}, divided into two phases: detection and negotiation.
 \SetKwBlock{negotPhase}{ Negotiation Phase:}{}
@@
   \negotPhase{
-  Coordinator initializes $\vec{\theta}^{(0)}$ \;
-  $p:=0$\;
-
-  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
-  Subsystems solve~\eqref{eq:DOP_local}, and send $\vec{\lambda}^{\star}_{i}(\vec{\theta}\p)$\;
-
-  Coordinator updates allocation~\eqref{eq:thetaNegot} using adequate versions of $\vec{\lambda}_{i}\p$ for each agent:
-  $\tilde{\vec{\lambda}}_{i}\p$, if $d_{i}=0$ and ${\vec{\lambda}_{i}}_{\mathrm{rec}}$, if ${d_{i}=1}$\;
-  $p:=p+1$
- }}
+  Apply Algorithm~\ref{alg:quantityAlg} with adequate versions of $\vec{\lambda}_{i}\p$:\;
+  \quad$\tilde{\vec{\lambda}}_{i}\p$, if $d_{i}=0$ and ${\vec{\lambda}_{i}}_{\mathrm{rec}}$, if ${d_{i}=1}$~\eqref{eq:lambda_reconstruction}\;
+ }
  \caption{Secure DMPC.}\label{alg:safeDMPC}
@@
 
-\section{Simulation}
+\section{Example: Temperature Control}\label{sec:App_dMPC}
+The system consists of $4$~distinct~rooms (I to IV), which we want to control the air temperatures inside each one of them.
+The systems are modeled as continuous-time linear \mbox{time-invariant} systems using the \mbox{3R-2C} model~\citep{NogueiraEtAl2021}, with parameters shown in Tables~\ref{tab:modelParamMeaning} and~\ref{tab:modelParam}, and state space matrices
+\begin{equation*}
+  \label{eq:room_system}
+  \begin{matrix}
+\resizebox{.98\columnwidth}{!}{$
+  A_{\mathrm{c}_{i}}=\left[
+    \begin{matrix}
+      -\frac{1}{C^{\text{walls}}_{i}R^{\text{oa/ia}}_{i}}-\frac{1}{C^{\text{walls}}_{i}R^{\text{iw/ia}}_{i}}& \frac{1}{C^{\text{walls}}_{i}R^{\text{iw/ia}}_{i}}\\
+      \frac{1}{C^{\text{air}}_{i}R^{\text{iw/ia}}_{i}} &-\frac{1}{C^{\text{air}}_{i}Ro_{i}}-\frac{1}{C^{\text{air}}_{i}R^{\text{iw/ia}}_{i}}
+    \end{matrix}\right]
+    $}
+  \\
+  \begin{matrix}
+    B_{\mathrm{c}_{i}}=\left[
+      \begin{matrix}  \frac{10}{C^{\text{walls}}_{i}}& 0\end{matrix}
+    \right]\T&C_{\mathrm{c}_{i}}=\left[\begin{matrix}1 & 0\end{matrix}\right]
+  \end{matrix}
+  \end{matrix},
+\end{equation*}
+where the states are ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$, which represent the mean temperatures of the air and walls inside room $i$, and the input of the system is $\vec{u}_{i}$, which is the heating power
+for the corresponding room.
+
+The inputs are constraint by ${\sum_{i=1}^{4}\vec{u}_{i}(k)\preceq 4\mathrm{kW}}$.
+
+\begin{table}[b]
+  \centering
+  \caption{Model Parameters}\label{tab:modelParamMeaning}
+  \begin{tabular}[b]{cl}
+    \toprule
+    Symbol&Meaning\\
+    \midrule
+    $C^{\text{air}}_{i}$&Heat Capacity of Inside Air\\
+    $C^{\text{walls}}_{i}$&Heat Capacity of External Walls\\
+    $R^{\text{oa/ia}}_{i}$&Resistance Between Inside and Outside Air (from windows)\\
+    $R^{\text{iw/ia}}_{i}$&Resistance Between Inside Air and Inside Walls\\
+    $R^{\text{ow/oa}}_{i}$&Resistance Between Outside Air and Outside Walls\\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+  \begin{table}[b]
+  \centering
+  \caption{
+\todo{
+  change values
+}
+    Value of parameters for each agent}\label{tab:modelParam}
+  \begin{tabular}[t]{cccccc} \toprule
+    Symbol& I & II & III & IV &Unit\\
+    \midrule
+    $C^{\text{walls}}$    &$5$   & $4$ & $4.5$ &$4.7$ &$10^{4}\mathrm{J/K}$ \\
+    $C^{\text{air}}$               &$8$   & $7$ &$9$&$6$  &$10^{4}\mathrm{J/K}$  \\
+    $R^{\text{oa/ia}}$               &$5$   & $6$& $4$& $5$  &$10^{-3}\mathrm{K/W}$ \\
+    $R^{\text{iw/ia}}$               &$2.5$ & $2.3$&$2$&$2.2$ &$10^{-4}\mathrm{K/W}$\\
+    $R^{\text{ow/oa}}$               &$0.5$ & $1$ &$0.8$&$0.9$  & $10^{-4}\mathrm{K/W}$ \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+The subsystems are discretized using the zero-order hold discretization method with sampling time
+${T_{s}=0.25\mathrm{h}}$
+and the quantity decomposition-based DMPC is implemented using prediction horizon ${\predhorz=4}$.
+
+Three scenarios are simulated for a period of 12.5 hours:
+
+  \begin{enumerate}
+    \item Nominal behavior.
+    \item \mbox{Agent I presents non-cooperative behavior} ${T_{I}[k]}$ for ${k\geq25}$, without correction.
+    \item \mbox{Agent I presents non-cooperative behavior} ${T_{I}[k]}$ for ${k\geq25}$, with correction, ${\epsilon_{P}=10^{-4}}$.
+  \end{enumerate}
+
+In Fig.~\ref{fig:response3Scenarios}, first, we compare the air temperature in room I
+with its reference (${w_{I}=25.5^{\circ}}$C), and then the
+decision variable ${E_{I}(k)}$ with the threshold $\epsilon_{P}$.
+Observe that in the nominal behavior (in magenta), the reference $w_{I}$ is not reached
+due to power constraints the systems are under.
+
 \begin{figure}[h]
@@
  \includegraphics[width=\columnwidth]{airtemp_roomI/__ErrorWX_command_normErrH.pdf}
-  \caption{Air temperature in room I and the Frobenius norm of the decision variable $E_{I}[k]$ for different scenarios: nominal (N), selfish behavior without correction (S),
+  \caption{Air temperature in room I and the decision variable $E_{I}[k]$ for different scenarios: nominal (N), selfish behavior without correction (S),
 and selfish behavior with correction.}\label{fig:response3Scenarios}
 \end{figure}
-\begin{figure}[h]
-  \centering
- \includegraphics[width=\columnwidth]{airtemp_roomI/control.pdf}
-  \caption{Control applied in all rooms for the 3 different scenarios: nominal, selfish behavior without correction, and selfish behavior with correction (C).}\label{fig:response3Scenarios}
-\end{figure}
 
+As expected, the decision variable lies under the threshold $\epsilon_{P}=10^{-4}$ with values of order ${E_{I}^{N}(k)\approx10^{-10}}$.
+
+When the agent presents a selfish behavior (in orange), the tracking error
+${w_{I}-y_{I}}$ is reduced but insufficient to attain the reference.
+In this case, the detection variable surpasses $\epsilon_{P}$,
+${E_{I}^{S}\approx200}$, indicating the change of behavior.
+
+When the correction is activated in the system, we see that the temperatures approach their nominal value $y_{I}^{N}$.
+We can also illustrating the good
+performances of our proposition by comparing the inputs in Fig.~\ref{fig:control_3Scenarios}.
+When I is selfish its control increases while the other decreases, unless the correction is activated and it approaches the nominal.
+
+We can also evaluate the performance of the proposed mechanism by comparing the global objective function calculated using the total period of simulation ${N=50}$ for the 3 scenarios (Table~\ref{tab:costsGlobalLocal}).
+When agent I is selfish, we see the increase in the objective, degrading the overall performance.
+When the correction mechanism is activated, the difference between costs is minimal, $|\frac{J^{C}-J^{N}}{J^{N}}|\leq10^{-8}$.
+%%%%%
 % TODO(accacio): Verify
@@
     \midrule
-    I & -11468 &  -9074 & -11468 \\
-    II & -6536 &  -7151 & -6536 \\
-    III & -3192 &  -4052 & -3192 \\
-    IV & -2849 &  -3694 & -2849 \\
-    Global & -24044 &  -23971 & -24044 \\
-
+    % I & -11468 &  -9074 & -11468 \\
+    % II & -6536 &  -7151 & -6536 \\
+    % III & -3192 &  -4052 & -3192 \\
+    % IV & -2849 &  -3694 & -2849 \\
+    % Global & -24044 &  -23971 & -24044 \\
+    Global & -24044.180111 &  -23971.164122 & -24044.180027\\
     \bottomrule
@@
 \end{table}
+% \enlargethispage{-.8cm}
 
-\section{Conclusion}
+\section{CONCLUSION AND FUTURE WORKS}\label{sec:CC}
+\todo{
+In this paper, an algorithm for monitoring and correcting exchanges between agents in a resource-sharing system has been proposed.
+The algorithm exploits the particular structure of exchanges, part of which must be constant over time.
+The first phase consists of identifying this constant part and checking if an attacker has modified it.
+From this identification, it is possible to reconstruct the original mechanism and find the centralized optimality.
+This principle should be generalized to other types of decomposition structures, and this is what we plan to do in the near future.
+}
 
-A conclusion section is not required. Although a conclusion may review
-the main points of the paper, do not replicate the abstract as the
-conclusion. A conclusion might elaborate on the importance of the work
-or suggest applications and extensions.
+\begin{figure}[h]
+  \centering
+ \includegraphics[width=\columnwidth]{airtemp_roomI/control.pdf}
+  \caption{Control applied in all rooms for the 3 scenarios.}\label{fig:control_3Scenarios}
+\end{figure}
 
@@
 \bibliography{bibliography}
+% \bibliographystyle{ifacconf-harvard}
 % \todo[Change bibliography to the one used in the aux]{}

### 2022-02-11 code/generate_figures.py

@@ -89,6 +89,11 @@ print("II &", round(nominalII),"& ",round(selfishII),"&",round(correctedII),"\\\
 print("III &", round(nominalIII),"& ",round(selfishIII),"&",round(correctedIII),"\\\\")
 print("IV &", round(nominalIV),"& ",round(selfishIV),"&",round(correctedIV),"\\\\")
 print("Global &", round(nominal_global),"& ",round(selfish_global),"&",round(corrected_global),"\\\\")
+print("Global &", round(nominal_global)/round(nominal_global),"& ",round(selfish_global)/round(nominal_global),"&",round(corrected_global)/round(nominal_global),"\\\\")
+print("Global &", (nominal_global),"& ",(selfish_global),"&",(corrected_global),"\\\\")
+print((corrected_global-nominal_global)/nominal_global)
+print((selfish_global-nominal_global)/nominal_global)
+
 
 # print("Nominal")
 # print(np.sum((Wt[0:simK,0]-xt[0:simK,0])*(Wt[0:simK,0]-xt[0:simK,0])+uHist[0,-1,:,0]*uHist[0,-1,:,0])) # J1
@@ -124,36 +129,39 @@ print("Global &", round(nominal_global),"& ",round(selfish_global),"&",round(cor
 #       +np.sum((correctWt[0:simK,3]-correctxt[0:simK,3])*(correctWt[0:simK,3]-correctxt[0:simK,3])+correctuHist[0,-1,:,3]*correctuHist[0,-1,:,3])) # J4
 
 # NOTE(accacio): Detection
-# fig, axs = plt.subplots(2, 1)
+fig, axs = plt.subplots(2, 1)
 
-# axs[0].plot(np.arange(0,simK+1),np.matlib.repmat(nominal_Wt[0],simK+1,1),'-',drawstyle='steps-post')
-# axs[0].plot(np.arange(0,simK+1),nominal_xt[0,0:simK+1,0],'-',drawstyle='steps-post')
-# axs[0].plot(np.arange(0,simK+1),selfish_xt[0,0:simK+1,0],'-',drawstyle='steps-post')
-# axs[0].scatter(np.arange(0,simK+1),corrected_xt[0,0:simK+1,0],s=15,color='black')
[…]

### 2022-02-11 code/update_parameters.m

@@ -1,19 +1,18 @@
 function [Phi, pi_new] = update_parameters(Theta, Lambda, Responsabilities)
 % UPDATE_PARAMETERS -
     modes=size(Responsabilities,1);
-    [n N]=size(Theta);
+    [c O]=size(Theta);
 
-    Upsilon=kron(ones(n,1),eye(n));
-    Delta=kron(eye(N),ones(1,n));
-    G=kron(ones(1,N),eye(n));
-    Y=kron(G,ones(n,1));
+    Upsilon=kron(ones(c,1),eye(c));
+    Delta=kron(eye(O),ones(1,c));
+    G=kron(ones(1,O),eye(c));
+    Y=kron(G,ones(c,1));
     Omega=sparse([(Upsilon*Theta*Delta).*Y; G]');
 
-    N_k=sum(Responsabilities,2);
-    pi_new(:)=N_k/N;
+    pi_new(:)=sum(Responsabilities,2)/O;
     for i=1:modes
         responsabilities=Responsabilities(i,:);
-        resp2=cellfun(@(x) x*eye(n),mat2cell(responsabilities',ones(1,N)),'UniformOutput',0);
+        resp2=cellfun(@(x) x*eye(c),mat2cell(responsabilities',ones(1,O)),'UniformOutput',0);
         Gamma=sqrt(sparse(blkdiag(resp2{:})));
 
         Phi(i,:)=-((Gamma*Omega)\(Gamma*Lambda(:)));

### 2022-02-12 article/article.tex
Intro, conclusion, abstract

@@
 \begin{abstract}% Abstract of not more than 250 words.
-  \todo[write abstract]{
-  In distributed predictive control structures, communication among agents is required to achieve a consensus and approach an optimal global behavior.
-Such negotiation mechanisms are sensitive to attacks on these exchanges.
-  This paper proposes a monitoring scheme based on Expectation Maximization to detect and mitigate the effects of attacks in a resource allocation framework.
+  Controlling large scale systems sometimes requires decentralized computation.
+  Communication among agents is crucial to achieve consensus and optimal global behavior.
+  These negotiation mechanisms are sensitive to attacks on those exchanges.
+  This paper proposes an algorithm based on Expectation Maximization to mitigate the effects of attacks in a resource allocation based distributed model predictive control.
   The performance is assessed through academical example of the temperature control of multiple rooms under input power constraints.
-}
 \end{abstract}
@@
 
-Given the complexity, scale and even geographic location of these structures, it is necessary to adopt control structures that are decentralized rather than \todo{monolithic}.
+Given the complexity, scale and even geographic location of these structures, it is necessary to adopt control structures that are decentralized rather than monolithic.
 For these systems, a set of controllers are designed to ensure the safe and efficient operation of the facilities collectively.
@@
 This operation is done cooperatively, where the controllers exchange information with each other.
-This approach can be used, for example, for the voltage regulation of electrical networks, and the production-consumption balance in these same networks.
-Multi-agent systems~\citep{KantamneniEtAl2015} or \dmpc{}~\citep{MaestreEtAl2014} are standard tools to design these controllers.
+This approach can be used, i.e., for the voltage regulation of electrical networks, or the production-consumption balance in these same networks, water distribution and many other applications.
+Multi-agent systems~\citep{KantamneniEtAl2015} and \dmpc{}~\citep{MaestreEtAl2014} are standard approaches to design these controllers.
 
@@
 
-Our previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.
+A previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.
 
@@
 
-\todo{
 This paper is organized as follows.
-First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based DMPC and its vulnerabilities.
+First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based \dmpc{} and its vulnerabilities.
 Then in Section~\ref{sec:TSM} we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
 Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with academic simulation and we assess its performance.
-Finally, in Section~\ref{sec:CC}, we conclude, and we give an outlook of future works.
-}
+Finally, in Section~\ref{sec:CC}, we conclude with an outlook of future works.
 
@@
 The \emph{local problems}~\eqref{eq:DOP_local} can be rewritten in a equivalent (same solution) constrained \qp{} form:
-\todo[Correct alignment maybe]{
   \begin{subequations}\label{eq:quadratic_case}
@@
       \underset{\Uik}{\mathrm{minimize}}&& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
-      \mathrm{s.t.}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
+      \mathrm{subject~ to}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
                                         &&\Uik\succeq\0_{c},
@@
   \end{subequations}
-}
 where ${\Uik}$ stacks the input prediction sequences, and
@@
 The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
-The approach depends only on the fact that $H_{i}:\semidefpos$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
+The approach depends only on the fact that ${H_{i}:\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
 
@@
 The system consists of $4$~distinct~rooms (I to IV), which we want to control the air temperatures inside each one of them.
-The systems are modeled as continuous-time linear \mbox{time-invariant} systems using the \mbox{3R-2C} model~\citep{NogueiraEtAl2021}, with parameters shown in Tables~\ref{tab:modelParamMeaning} and~\ref{tab:modelParam}, and state space matrices
+The systems are modeled as continuous-time linear \mbox{time-invariant} systems using the \mbox{3R-2C} model~\citep{NogueiraEtAl2021}, with parameters shown in Tables~\ref{tab:modelParamMeaning} and~\ref{tab:modelParam}, and dynamics
+\begin{equation}
+\begin{matrix}
+  \label{eq:systems}
+\dot{\vec{x}}_{i}(t)&=&{A_{c}}_{i}\vec{x}_{i}(t) &+& {B_{c}}_{i}\vec{u}_{i}(t)\\
+\vec{y}_{i}(t)&=&{C_{c}}_{i}\vec{x}_{i}(t) &&
+\end{matrix},
+\end{equation}
+with
 \begin{equation*}
@@
 \end{equation*}
-where the states are ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$, which represent the mean temperatures of the air and walls inside room $i$, and the input of the system is $\vec{u}_{i}$, which is the heating power
-for the corresponding room.
-
-The inputs are constraint by ${\sum_{i=1}^{4}\vec{u}_{i}(k)\preceq 4\mathrm{kW}}$.
+where ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$, which represents the mean temperatures of the air~(${x_A}_i$) and walls inside room~$i$~(${x_W}_i$), and $\vec{u}_{i}$ is the input (the heating power)
+for the corresponding room. The inputs are constraint by ${\sum_{i=1}^{4}\vec{u}_{i}(t)\preceq 4\mathrm{kW}}$.
 
@@
   \caption{
-\todo{
-  change values
-}
-    Value of parameters for each agent}\label{tab:modelParam}
+    Parameters for each agent}\label{tab:modelParam}
   \begin{tabular}[t]{cccccc} \toprule
@@
     \midrule
-    $C^{\text{walls}}$    &$5$   & $4$ & $4.5$ &$4.7$ &$10^{4}\mathrm{J/K}$ \\
-    $C^{\text{air}}$               &$8$   & $7$ &$9$&$6$  &$10^{4}\mathrm{J/K}$  \\
-    $R^{\text{oa/ia}}$               &$5$   & $6$& $4$& $5$  &$10^{-3}\mathrm{K/W}$ \\
-    $R^{\text{iw/ia}}$               &$2.5$ & $2.3$&$2$&$2.2$ &$10^{-4}\mathrm{K/W}$\\
-    $R^{\text{ow/oa}}$               &$0.5$ & $1$ &$0.8$&$0.9$  & $10^{-4}\mathrm{K/W}$ \\
+    $C^{\text{walls}}$    &$5.4$&$4.9$&$4.7$&$4.7$ &$10^{4}\mathrm{J/K}$ \\
+    $C^{\text{air}}$               &$7.5$ &$8.4 $&$8.2$ &$7.7$&$10^{4}\mathrm{J/K}$  \\
+    $R^{\text{oa/ia}}$               &$5.2$&$4.6$&$4.9$&$5.4$&$10^{-3}\mathrm{K/W}$ \\
+    $R^{\text{iw/ia}}$               &$2.3$&$2.4$&$2.3$&$2.9$&$10^{-4}\mathrm{K/W}$\\
+    $R^{\text{ow/oa}}$               &$1.5$&$0.6$&$0.7$&$0.7$& $10^{-4}\mathrm{K/W}$ \\
     \bottomrule
@@
 
-In Fig.~\ref{fig:response3Scenarios}, first, we compare the air temperature in room I
+In Fig.~\ref{fig:response3Scenarios}, first, we compare the air temperature in room~I
 with its reference (${w_{I}=25.5^{\circ}}$C), and then the
@@
 When the agent presents a selfish behavior (in orange), the tracking error
-${w_{I}-y_{I}}$ is reduced but insufficient to attain the reference.
+${w_{I}-y_{I}}$ is reduced, almost attaining the reference.
 In this case, the detection variable surpasses $\epsilon_{P}$,
@@
 performances of our proposition by comparing the inputs in Fig.~\ref{fig:control_3Scenarios}.
-When I is selfish its control increases while the other decreases, unless the correction is activated and it approaches the nominal.
+When room~I is selfish, its control increases while the other decreases, unless the correction is activated and it approaches the nominal.
 
-We can also evaluate the performance of the proposed mechanism by comparing the global objective function calculated using the total period of simulation ${N=50}$ for the 3 scenarios (Table~\ref{tab:costsGlobalLocal}).
-When agent I is selfish, we see the increase in the objective, degrading the overall performance.
+We can also evaluate the performance of the proposed mechanism by comparing the global objective function calculated using the total period of simulation (${N=50}$) for the 3 scenarios (Table~\ref{tab:costsGlobalLocal}).
+When agent~I is selfish, we see the increase in the objective, degrading the overall performance.
 When the correction mechanism is activated, the difference between costs is minimal, $|\frac{J^{C}-J^{N}}{J^{N}}|\leq10^{-8}$.
@@
 % TODO(accacio): Verify
-\begin{table}[t]
+\begin{table}[h]
   \centering
@@
 \section{CONCLUSION AND FUTURE WORKS}\label{sec:CC}
-\todo{
-In this paper, an algorithm for monitoring and correcting exchanges between agents in a resource-sharing system has been proposed.
-The algorithm exploits the particular structure of exchanges, part of which must be constant over time.
-The first phase consists of identifying this constant part and checking if an attacker has modified it.
-From this identification, it is possible to reconstruct the original mechanism and find the centralized optimality.
-This principle should be generalized to other types of decomposition structures, and this is what we plan to do in the near future.
-}
+In this paper, we proposed an algorithm for monitoring and correcting exchanges between agents in a resource-sharing framework.
+The first phase of the algorithm consists of identifying the attacker by exploiting the exchange structure.
+After this identification, if necessary, it is possible to reconstruct the original mechanism and recover nominal optimality by inverting the attack.
+This principle should be generalized to cases when the attack is not completely invertible, reconstructing by parts the original mechanism.
+Also, other types of decomposition structures and attack models need to be explored, and this is what we plan to do in the near future.
 

### 2022-02-14 article/article.tex
change simulation, add data, Submitted to NecSys

@@
 \draft% show todos in red
-% \final% give error if there is todos
+\final% give error if there is todos
 % \comments% blank right column to make comments in pen & paper
@@
 \begin{abstract}% Abstract of not more than 250 words.
-  Controlling large scale systems sometimes requires decentralized computation.
-  Communication among agents is crucial to achieve consensus and optimal global behavior.
+  Controlling large-scale systems sometimes requires decentralized computation.
+  Communication among agents is crucial to achieving consensus and optimal global behavior.
   These negotiation mechanisms are sensitive to attacks on those exchanges.
   This paper proposes an algorithm based on Expectation Maximization to mitigate the effects of attacks in a resource allocation based distributed model predictive control.
-  The performance is assessed through academical example of the temperature control of multiple rooms under input power constraints.
+  The performance is assessed through an academic example of the temperature control of multiple rooms under input power constraints.
 \end{abstract}
@@
 \begin{keyword}
-  Model predictive and optimization-based control,
-  Distributed control and estimation,
-Large scale optimization problems
+%   Model predictive and optimization-based control,
+%   Distributed control and estimation,
+% Large scale optimization problems
+Distributed constrained control and MPC;
+Cyber security in networked control systems;
+Decentralized Control and Large-Scale Systems
 \end{keyword}
@@
 \section{INTRODUCTION}
-The information collected, processed, and transmitted within \cps{} impacts physical systems whose proper functioning is essential for society, like energy networks, industrial processes, water production plants, etc..
+The information collected, processed, and transmitted within \cps{} impacts physical systems whose proper functioning is essential for society, like energy networks, industrial processes, water distribution, and others.
 Attacks in those critical infrastructures, such as Stuxnet, increased awareness for cyber-security of \cps{}.
 
-Given the complexity, scale and even geographic location of these structures, it is necessary to adopt control structures that are decentralized rather than monolithic.
-For these systems, a set of controllers are designed to ensure the safe and efficient operation of the facilities collectively.
+Given these structures' complexity, scale, and geographic distribution, it is necessary to adopt decentralized control structures rather than monolithic ones.
+For these systems, a set of controllers are designed to collectively ensure the safe and efficient operation of the facilities.
 
-This operation is done cooperatively, where the controllers exchange information with each other.
-This approach can be used, i.e., for the voltage regulation of electrical networks, or the production-consumption balance in these same networks, water distribution and many other applications.
+This operation is done cooperatively, where the controllers exchange information.
+This approach can be used, i.e., for the voltage regulation of electrical networks, the production-consumption balance in these same networks, water distribution, and many other applications.
 Multi-agent systems~\citep{KantamneniEtAl2015} and \dmpc{}~\citep{MaestreEtAl2014} are standard approaches to design these controllers.
 
-A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018} and primal decomposition~\citep{NogueiraEtAl2021}.
+A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018}, and primal decomposition~\citep{NogueiraEtAl2021}.
 
@@
 A modification of a local goal, the hacking of communication, or the failure of one of the agents are events that disrupt global behavior.
-If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
+If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to a deviant behavior that is more difficult to detect.
 % Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to lead the system towards a state of malfunction/instability~\todo[use other]{}.
 
-Some strategies are presented to mitigate the effects of such event.
-For instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
+Some strategies are presented to mitigate the effects of such an event.
+For instance, dismissing extreme values as in~\cite{VelardeEtAl2017} or
 using secure scenarios based on reliable historical data~\cite{MaestreEtAl2021}.
 
-A previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.
+\cite{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.
 
-In this paper, we propose an extension of our approach by including local constraints and changing the global constraints.
-This extension profoundly changes the complexity, as the exchanges between the
-agents and the coordinator are no longer characterized by an affine functions
-but by \pwa{} functions.
-This leads not only to a combinatorial explosion but also to a parametric identification challenge.
-To overcome this problem, we propose using a learning method based on the \EM{} algorithm~\citep{DempsterEtAl1977}, which allows us to estimate the corruption
-mechanism and correct it if necessary.
+In this paper, we propose an extension of their approach by including local constraints and changing the global constraints.
+This extension profoundly changes the complexity, as the exchanges between the agents and the coordinator are no longer characterized by affine functions but by \pwa{} functions.
+This fact leads not only to a combinatorial explosion but also to a parametric identification challenge.
+To overcome this problem, we propose using a learning method based on the \EM{} algorithm~\citep{DempsterEtAl1977}, which allows us to estimate the corruption mechanism and correct it if necessary.
 
 This paper is organized as follows.
-First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based \dmpc{} and its vulnerabilities.
-Then in Section~\ref{sec:TSM} we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
-Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with academic simulation and we assess its performance.
+First, Section~\ref{sec:PS} introduces the primal decomposition-based \dmpc{} and its vulnerabilities.
+Then in Section~\ref{sec:TSM}, we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
+Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with an academic simulation and assess its performance.
 Finally, in Section~\ref{sec:CC}, we conclude with an outlook of future works.
@@
 \emph{Notation:}
-In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
+In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and ${\norm{\vec{v}}_{Y}=\norm{Y^{\frac{1}{2}}\vec{v}}}$ represent the $\ell_{2}$, Frobenius and weighted norms.
 $\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
@@
 $\vectorize{A}$ vectorizes matrix $A$.
-${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
-$[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
+${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ means ${i=1}$.
+% $[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
 ${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
@@
 $\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
-$\preceq$ is the component-wise inequality for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X\T=X}}$.
-${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
-$I_{c}$ is the identity matrix of size $c$.
-$A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}$.
-$\determinant{A}$ is the determinant of $A$.
+$\symmetric^{n}$, and $\defpos^{n}$ ($\semidefpos^{n}$) are symmetric and positive (semi-)definite matrices of size $n$.
+${A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}}$.
+% $\determinant{A}$ is the determinant of $A$.
 A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
-$\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
-$\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
+% $\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
+$\expectation{\random{x}}$ is the expected value of random variable $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
 % $\pdf{\cdot}$ is a probability density function.
@@
 \section{Problem Statement}\label{sec:PS}
-In this section we present an algorithm to decompose a \mpc{} problem based on the primal decomposition \citep{BoydEtAl2015}, and a vulnerability of the decomposition that can be exploited by a malicious agent.
+In this section, we present a \dmpc{} algorithm based on the primal decomposition \citep{BoydEtAl2015} and a vulnerability of the decomposition, which can be exploited.
 
@@
 with ${\umax:\R^{n_{u}}}$.
-The agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
+The agents are coupled by global input constraints with weighting matrices ${\Gamma_{i}:\semidefpos^{n_{u}}}$:
 \begin{equation}
@@
 where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
-$\optglobobj$ denotes the optimal value of the objective function for the the optimal control sequences ${\optuiseq}$.
+$\optglobobj$ denotes the optimal value of the objective function for the optimal control sequences ${\optuiseq}$.
 The problem in~\eqref{eq:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a \rhs{}.
 
-Solving~\eqref{eq:GOP} at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive.
-Another issue that can arise is the fact that the complete information \todo[verify]{(${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$)} from all the subsystems is needed to solve the problem,
-which can be viewed negatively in a confidentiality point of view.
+Since the prediction horizon and the number of agents can be large, solving~\eqref{eq:GOP} at each time $k$ can be rather challenging due to computational costs.
+Another issue that can arise is the fact that the complete information ${\set{I}_i=\{A_i,B_i,C_i,Q_i,R_i,\Gamma_i,\vec{v}_i\}}$ from all the subsystems is needed to solve the problem,
+which can be viewed negatively from a confidentiality point of view.
 
-So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.
+So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with which we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.
 
@@
 
-The technique consists of decomposing the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
+The technique decomposes the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local} (solvable in parallel by each subsystem) called \emph{local problems}, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
 \begin{subequations}
@@
 \end{subequations}
-The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}.
+The modified \mpc{} problems~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, which correspond to the quantity of the total resource $\vec{u}_{\max}$ allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quantity decomposition}.
 This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$.
-The sequence of this variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.
+The sequences ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.
 
@@
 \newcommand{\negotiation}{\emph{negotiation}}
-which henceforth is referred as the \negotiation.
+which henceforth is referred to as the \negotiation.
 
@@
 
-The algorithm to find the optimal $\optthetai[k]$ can be summarized in algorithm~\ref{alg:quantityAlg}, and the exchange between coordinator and agents can be seen in Fig.~\ref{fig:dmpc_graph}.
+The algorithm to find the optimal $\optthetai[k]$ can be summarized in Algorithm~\ref{alg:quantityAlg}, and the exchange between coordinator and agents can be seen in Fig.~\ref{fig:dmpc_graph}.
+
+\begin{figure}[h]
+  \centering
+    \resizebox{.47\columnwidth}{!}{
+          \begin{tikzpicture}[font=\small,thick,node distance=.0cm and .5cm,
+          mpcSmall/.style={
+            rectangle,
+            align=center,
+            fill=supelecRed!10,
+            minimum width=1.5cm,
+            minimum height=1cm},
+          coordinator/.style={
+            rectangle,
+            align=center,
+            fill=mpc_grey,
+            minimum height=3.4cm,
+            minimum width=1.5cm,
+          },
+          superv/.style={
+            rectangle,
+            minimum height=1.4cm,
+            minimum width=.3cm,
+            fill=mpc_yellow,
+          },
+          negotiation/.style={
+            rectangle,
+            minimum height=3.0cm,
+            minimum width=.45cm,
+            fill=mpc_green,
+          },
+          ]
+
+          \node[draw,
+          mpcSmall,
+          fill=mpc_blue,
+          ] (block1) {MPC 1};
+
+          \node[draw,
+          mpcSmall,
+          fill=none,
+          draw=white,
+          below=of block1,
+          ] (mult) {\vdots};
+
+          \node[draw,
+          mpcSmall,
+          fill=mpc_blue,
+          below=of mult,
+          ] (blockM) {MPC M};
+
+          \node[draw,
+          coordinator,
+          right=2cm of mult,
+          ] (coordinator) {};
+          \node at ($(coordinator)+(.25,0)$) {\rotatebox{90}{Coordinator}};
+
+          \node[draw,negotiation] (negot) at ({$(coordinator.center)+(0,0)$} -|{$(coordinator.west)+(0.25,)$}) {};
+          \node  at (negot) {\rotatebox{90}{Negotiation}};
+
+          \draw[latex-,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
+          \draw[-latex,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};
+
+          \draw[latex-,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
+          \draw[-latex,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
+
+        \end{tikzpicture}
+        }
+  \caption{Exchange between coordinator and agents.}\label{fig:dmpc_graph}
+\end{figure}
 
@@
 }
- \caption{Quantity decomposition based \acrshort{dmpc}.}\label{alg:quantityAlg}
+ \caption{Quantity decomposition-based \acrshort{dmpc}.}\label{alg:quantityAlg}
 \end{algorithm2e}
 
-One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
+One can observe that each agent only sends $\lambdaik$ instead of using $\mathcal{I}_{i}$.
 An interpretation of $\lambdaik$ is the dissatisfaction of agent $i$ with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
-The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent which can take advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.
+The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations.
+However, if false data is injected, the \negotiation{} can be driven by a malicious agent, taking advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.
 
@@
       \label{eq:4}
-      \underset{\Uik}{\mathrm{minimize}}&& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
-      \mathrm{subject~ to}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
-                                        &&\Uik\succeq\0_{c},
+      \underset{\Uik}{\mathrm{minimize}}&&\small J_{i}(\vec{\theta}_{i})=&\frac{1}{2} \Uik^T H_i\Uik +{\fik}^T\Uik\\
+      \mathrm{subject~ to}&&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
+                                        &&&\Uik\succeq\0_{c},
     \end{align}
@@
 ${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
-The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
+The values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
 The approach depends only on the fact that ${H_{i}:\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
@@
 \no{
-  We choose reference tracking just to illustrate the approach. With this objective, we have
+  We choose reference tracking to illustrate the approach. With this objective, we have
 ${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
@@
 ${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
-}
-\todo[make a better figure if time permits]{
-\begin{figure}[t]
-  \centering
-  \begin{tikzpicture}[node distance=1.5cm]
-    \tikzset{
-      agent/.style={draw,circle,minimum width=1cm,},
-      coord/.style={draw,circle,minimum width=1cm,very thick}
-    }
-    % \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
-    % \node at (1.2,-1.3) {O};
-    % \node (pi) at (-2.5,.5) {$\pi$};
-    % \node (phi) at (2.5,-1.) {$\phi_{c}$};
-
-    \node[agent] (a1) {a$_{1}$};
-    \node (a2) [right of=a1]  {$\dots$};
-    \node[agent] (am) [right of=a2] {a$_{M}$};
-    \node[coord] (c) [above of=a2] {C};
-
-    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}[k]$} (c) ;
-    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}[k]$} (a1);
-
-    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}[k]$} (c);
-    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}[k]$} (am);
-
-  \end{tikzpicture}
-  \caption{Exchange between coordinator (C) and agents $a_{i}$.}\label{fig:dmpc_graph}
-\end{figure}
 }
@@
 
-The halfspaces defined by the set of pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.
+The halfspaces defined by the pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.
 
@@
 For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
-And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.
+Futhermore, if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.
+
+% \begin{figure}[t]
+%   \centering
+%           \begin{tikzpicture}[font=\small,thick,node distance=.0cm and .5cm,
+%           mpcSmall/.style={
+%             rectangle,
+%             align=center,
+%             fill=supelecRed!10,
+%             minimum width=2.cm,
+%             minimum height=1cm},
+%           coordinator/.style={
+%             rectangle,
+%             align=center,
+%             fill=mpc_grey,
+%             minimum width=1cm,
+%             minimum height=3.5cm
+%           } ]
+
+%           \node[draw,
+%           mpcSmall,
+%           fill=mpc_blue,
+%           ] (block1) {MPC 1};
+
+%           \node[draw,
+%           mpcSmall,
+%           fill=none,
+%           draw=white,
+%           below=of block1,
+%           ] (mult) {\vdots};
+
+%           \node[draw,
+%           mpcSmall,
+%           fill=mpc_blue,
+%           below=of mult,
+%           ] (blockM) {MPC M};
+
+%           \node[draw,
+%           coordinator,
+%           right=2cm of mult,
+%           ] (coordinator) {\rotatebox{90}{Coordinator}};
+
+%           \draw[latex-,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\vec{\lambda_{1}}[k]$};
+%           \draw[-latex,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{1}}[k]$};
+
+%           \draw[latex-,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\vec{\lambda_{M}}[k]$};
+%           \draw[-latex,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
+
+%         \end{tikzpicture}
+%   \caption{Exchange between coordinator and agents.}\label{fig:dmpc_graph}
+% \end{figure}
 
@@
  \begin{assumption}
-   We assume that none of the combinations of active/inactive constraints are redundant (no linear dependency), nor make the optimization infeasible (no empty intersection).
+   We assume that none of the active/inactive constraints combinations are redundant (no linear dependency), nor make the optimization infeasible (no empty intersection).
    Thus we always have $N$ zones.
@@
 \end{equation}
-  where $\determinant{T}\neq 0$.
+  where $T$ is invertible.
 \end{assumption}
@@
 with $n\in\{1\mathbin{:}N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
-Observe that as the function is applied only in $\lambdaik$, it does not affect the hyperplanes generating the zones.
+Observe that as the function is applied only in $\lambdaik$, it does not affect the zones' hyperplanes.
 
-We fix the index for these zones where all constraints are active to $1$, and we can call them the $1$-zones.
+We fix the index for these zones where all constraints are active to $1$, now referenced to as the $1$-zones.
 
@@
 \begin{assumption}
-  We suppose there exists a $1$-zone for each agent $i$ for every time $k$.
+  We suppose a $1$-zone exists for each agent $i$ for every time $k$.
 \end{assumption}
-And from~\eqref{eq:lambdafuntheta}, we can derive a method to reconstruct $\vec{\lambda}_{i}$:
+Moreover, from~\eqref{eq:lambdafuntheta}, we can reconstruct $\vec{\lambda}_{i}$:
 \begin{equation}
@@
 \glsreset{EM}
-Unfortunately, since we don't know the hyperplanes that separates the different zones, we don't know how close of $\bar{\theta}_{i}=\0_{c}$ we need to generate our points.
-So we generate points arbitrarily close to $\bar{\theta}_{i}$ and we use the \EM{} algorithm which can potentially identify the parameters of all $N$ modes.
+Unfortunately, since we do not know the hyperplanes separating the different zones, we do not know how close of $\bar{\vec{\theta}}_{i}=\0_{c}$ we need to generate our points.
+So we generate points arbitrarily close to $\bar{\vec{\theta}}_{i}$ and use the \EM{} algorithm, which can potentially identify the parameters of all $N$ modes.
 \subsection{Expectation Maximization}
@@
 
-Given a set of parameters $\set{P}$, a set of observable data $\set{B}$ and a set of non-observable data $\set{U}$, the main objective of the \EM{} algorithm is to find estimators of $\set{P}$ which maximize the log marginal likelihood of the observed data ${\ln\probability{\set{B};\set{P}}}$, for models with latent variables in $\set{U}$.
+Given a set of parameters $\set{P}$, a set of observable data $\set{B}$ and a set of non-observable data $\set{U}$, the main objective of the \EM{} algorithm is to find estimators of $\set{P}$ that maximize the log marginal likelihood of the observed data ${\ln\probability{\set{B};\set{P}}}$, for models with latent variables in $\set{U}$.
 
-Since maximizing ${\ln\probability{\set{B};\set{P}}}$ does not have an analytical solution, the algorithm solves the optimization problem in a iterative manner.
+Since maximizing ${\ln\probability{\set{B};\set{P}}}$ does not have an analytical solution, the algorithm solves the optimization problem iteratively.
 So, rather than finding the $\set{P}$ that maximizes ${\ln\probability{\set{B};\set{P}}}$, we find the $\set{P}$ that maximizes the expectation of the complete data log-likelihood $\ln\probability{\set{B},\set{U};\set{P}}$ w.r.t.
@@
 calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$.
-This provides a convergence-guaranteed iterative algorithm ensuring at each iteration a monotonic increase of the log-marginal likelihood.
-
-The algorithm can be divided into 2 steps presented in algorithm~\ref{alg:em}.
+These steps provide a convergence-guaranteed iterative algorithm ensuring a monotonic increase of the log-marginal likelihood at each iteration (Algorithm~\ref{alg:em}).
 
@@
   Initialize parameters $\set{P}_{\mathrm{new}}$\;
-  \Repeat{$\set{P}_{\mathrm{cur}}$ converge}{
+  \Repeat{$\set{P}_{\mathrm{cur}}$ converges}{
     $\set{P}_{\mathrm{cur}}=\set{P}_{\mathrm{new}}$\;
@@
         \label{eq:m_step}
-        \set{P}_{\mathrm{new}}=\arg\underset{\set{P}}{\max}\ \expectation[{\probability{\set{U}|\set{B};\set{P}_{\mathrm{cur}}}}]{\ln\probability{\set{B},\set{U};\set{P}}}
+        \set{P}_{\mathrm{new}}=\arg\underset{\set{P}}{\max\!.}\ \expectation[{\probability{\set{U}|\set{B};\set{P}_{\mathrm{cur}}}}]{\ln\probability{\set{B},\set{U};\set{P}}}
         $}
@@
 
-As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a multidimensional expansion of the model referred as \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010} to map the relation between $\random{\Theta}$ and $\random{\Lambda}$.
+As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose using a multidimensional expansion of the model referred to as \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010} to map the relation between $\random{\Theta}$ and $\random{\Lambda}$.
 We will call the model \emph{mixture of affine regressions}, since our regressors have linear and constant terms:
@@
 \begin{remark}
-The $\Sigma^{z}$ used in~\eqref{eq:responsibilities} are updated using a technique called \emph{Simulated annealing}~\citep{OzerovFevotte2010}, where they are initialized with arbitrarily big values indicating the uncertainty of the parameters and it is reduced as the estimations converge.
+The $\Sigma^{z}$ used in~\eqref{eq:responsibilities} are updated using a technique called \emph{Simulated annealing}~\citep{OzerovFevotte2010}, where they are initialized with arbitrarily significant values indicating the uncertainty of the parameters and it is reduced as the estimations converge.
 \end{remark}
@@
 Here we introduce a variable ${\vec{\phi}^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$.
-Using the KKT conditions, we can find an optimal solution for the problem in~\eqref{eq:m_step}, by
+We can find an optimal $\vec{\phi}^{z}$ for the problem in~\eqref{eq:m_step}, by
 taking the gradients of~\eqref{eq:completedataLogLikelihood_expectation} with respect to vectors $\vec{\phi}^{z}$ and making them vanish.
@@
 Doing the same for $\pi^{z}$ we get
-\begin{equation}
+\begin{equation*}
   \label{eq:3}
   \pi^{z}=\sum\limits_{o\in\set{O}}\tfrac{\zeta_{zo}(\set{P}_{\mathrm{cur}})}{O}.
-\end{equation}
+\end{equation*}
 
-As we can see,~\eqref{eq:mstepestimation} is the solution of a weighted Least-Squares, having the responsibilities as weights, adjusting the contribution of all observations to the regression model of index $z$.
+As we can see,~\eqref{eq:mstepestimation} is the solution of a weighted Least-Squares, with responsibilities as weights, which adjust the contribution of all observations to the regression models.
 We can see some similarities to the K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
-Instead of affecting the observed data to a zone with 100\% of certainty (\emph{hard assignment}), \EM{} uses the responsibilities (\emph{soft assignment}) for each zone.
+Instead of affecting the observed data to a zone with 100\% of certainty (\emph{hard assignment}), \EM{} uses each zone's responsibilities (\emph{soft assignment}).
 
-Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use in our mitigation scheme proposed in \S\ref{ssec:DM}.
-Observe that not necessarily the $z$ indices correspond to the indices $n$ in~\eqref{eq:lambdafuntheta}.
-So, to find the $z$ that corresponds to ${n=1}$, we take the observation $o$ for $\bar{\theta}=\0_{c}$, which we know belongs to the $1$-zone, and we find the most probable $z$ with
-\begin{equation}\label{eq:argmaxz}
-\arg\underset{z}{\max}\ {\zeta_{zo}(\set{P})}.
-\end{equation}
+Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use them in our mitigation scheme proposed in \S\ref{ssec:DM}.
+Since the $z$ indices do not necessarily correspond to the indices $n$ in~\eqref{eq:lambdafuntheta} (due to initialization of $\set{P}$),
+to find the $z$ that corresponds to ${n=1}$, we take the observation $o$ for $\bar{\vec{\theta}}=\0_{c}$, which we know belongs to the $1$-zone, and we find the most probable $z$ with
+\begin{equation*}\label{eq:argmaxz}
+\arg\underset{z}{\max\!.}\ {\zeta_{zo}(\set{P})}.
+\end{equation*}
 
 \begin{remark}
-A discussion about the initialization and update of other parameters is beyond the scope of this article, so we refer the reader to~\cite{Bishop2006}, and~\cite{DempsterEtAl1977}.
+A discussion about the initialization and update of other parameters is beyond the scope of this article, so we refer the reader to~\cite{Bishop2006}.
 \end{remark}
@@
 Integrating the EM to our detection and mitigation mechanism we can propose in Algorithm~\ref{alg:safeDMPC} a secure \dmpc{}, divided into two phases: detection and negotiation.
+The new algorithm corresponds to adding a supervision layer for each agent as in Fig.~\ref{fig:safe_dmpc_graph}
 \SetKwBlock{negotPhase}{ Negotiation Phase:}{}
@@
 The system consists of $4$~distinct~rooms (I to IV), which we want to control the air temperatures inside each one of them.
-The systems are modeled as continuous-time linear \mbox{time-invariant} systems using the \mbox{3R-2C} model~\citep{NogueiraEtAl2021}, with parameters shown in Tables~\ref{tab:modelParamMeaning} and~\ref{tab:modelParam}, and dynamics
+The systems are modeled as continuous-time linear \mbox{time-invariant} systems using the \mbox{3R-2C} model, with parameters in Tables~\ref{tab:modelParamMeaning} and~\ref{tab:modelParam}, and dynamics
 \begin{equation}
 \begin{matrix}
-  \label{eq:systems}
+  \label{eq:systems_cont}
 \dot{\vec{x}}_{i}(t)&=&{A_{c}}_{i}\vec{x}_{i}(t) &+& {B_{c}}_{i}\vec{u}_{i}(t)\\
@@
 \end{equation*}
-where ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$, which represents the mean temperatures of the air~(${x_A}_i$) and walls inside room~$i$~(${x_W}_i$), and $\vec{u}_{i}$ is the input (the heating power)
+where ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$. ${x_A}_i$ and ${x_W}_i$ are the mean temperatures of the air and walls inside room~$i$. $\vec{u}_{i}$ is the input (the heating power)
 for the corresponding room. The inputs are constraint by ${\sum_{i=1}^{4}\vec{u}_{i}(t)\preceq 4\mathrm{kW}}$.
+
+\begin{figure}[t]
+  \centering
+  \resizebox{.6\columnwidth}{!}{
+          \begin{tikzpicture}[font=\small,thick,node distance=.0cm and .5cm,
+          mpcSmall/.style={
+            rectangle,
+            align=center,
+            fill=supelecRed!10,
+            minimum width=1.5cm,
+            minimum height=1cm},
+          coordinator/.style={
+            rectangle,
+            align=center,
+            fill=mpc_grey,
+            minimum width=2.5cm,
+            minimum height=4cm
+          },
+          superv/.style={
+            rectangle,
+            minimum height=1.5cm,
+            minimum width=.35cm,
+            fill=mpc_yellow,
+          },
+          negotiation/.style={
+            rectangle,
+            minimum height=3.5cm,
+            minimum width=.5cm,
+            fill=mpc_green,
+          },
+          ]
+
+          \node[draw,
+          mpcSmall,
+          fill=mpc_blue,
+          ] (block1) {MPC 1};
+
+          \node[draw,
+          mpcSmall,
+          fill=none,
+          draw=white,
+          below=of block1,
+          ] (mult) {\vdots};
+
+          \node[draw,
+          mpcSmall,
+          fill=mpc_blue,
+          below=of mult,
+          ] (blockM) {MPC M};
+
+          \node[draw,
+          coordinator,
+          right=2cm of mult,
+          ] (coordinator) {};
+          \node at ($(coordinator)+(.8,0)$) {\rotatebox{90}{Coordinator}};
+
+          \node[draw,superv] (superv1) at ({$(block1)+(0,0)$} -|{$(coordinator.west)+(.2,)$}) {};
+          \node  at (superv1) {\rotatebox{90}{Superv. 1}};
+
+          \node[draw,superv] (supervM) at ({$(blockM)+(0,0)$} -|{$(coordinator.west)+(.2,)$}) {};
+          \node  at (supervM) {\rotatebox{90}{Superv. M}};
+
+          \node[draw,negotiation] (negot) at ({$(coordinator.center)+(0,0)$} -|{$(superv1.east)+(1.0,)$}) {};
+          \node  at (negot) {\rotatebox{90}{Negotiation}};
+
+          \draw[latex-,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
+          \draw[-latex,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};
+
+          \draw[latex-latex,very thick] (supervM.east) -- ( {$(supervM.east)$} -| negot.west) node [above,midway] {};
+          \draw[latex-latex,very thick] (superv1.east) -- ( {$(superv1.east)$} -| negot.west) node [above,midway] {};
+
+          \draw[latex-,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
+          \draw[-latex,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
+
+        \end{tikzpicture}
+        }
+  \caption{Exchange between agents in secure \dmpc{}.}\label{fig:safe_dmpc_graph}
+\end{figure}
 
@@
     $C^{\text{walls}}_{i}$&Heat Capacity of External Walls\\
-    $R^{\text{oa/ia}}_{i}$&Resistance Between Inside and Outside Air (from windows)\\
-    $R^{\text{iw/ia}}_{i}$&Resistance Between Inside Air and Inside Walls\\
-    $R^{\text{ow/oa}}_{i}$&Resistance Between Outside Air and Outside Walls\\
+    $R^{\text{iw/ia}}_{i}$&Resist. Between Inside Air and Inside Walls\\
+    $R^{\text{ow/oa}}_{i}$&Resist. Between Outside Air and Outside Walls\\
+    $R^{\text{oa/ia}}_{i}$&Resist. Between Inside and Out.\ Air (from windows)\\
     \bottomrule
@@
 
-The subsystems are discretized using the zero-order hold discretization method with sampling time
-${T_{s}=0.25\mathrm{h}}$
-and the quantity decomposition-based DMPC is implemented using prediction horizon ${\predhorz=4}$.
+The subsystems are discretized using zero-order hold discretization method with sampling time
+${T_{s}=\input{sampling.tex}\mathrm{h}}$
+and the quantity decomposition-based \dmpc{} is implemented using prediction horizon ${\predhorz=\input{prediction_horizon.tex}}$.
 
-Three scenarios are simulated for a period of 12.5 hours:
+Three scenarios are simulated for a period of $\input{simulation_time.tex}\mathrm{h}$:
 
@@
     \item Nominal behavior.
-    \item \mbox{Agent I presents non-cooperative behavior} ${T_{I}[k]}$ for ${k\geq25}$, without correction.
-    \item \mbox{Agent I presents non-cooperative behavior} ${T_{I}[k]}$ for ${k\geq25}$, with correction, ${\epsilon_{P}=10^{-4}}$.
+    \item \mbox{Agent I presents non-cooperative behavior} for ${k\geq25}$.
+    \item \mbox{Agent I presents non-cooperative behavior} for ${k\geq25}$, with correction.
   \end{enumerate}
@@
 In Fig.~\ref{fig:response3Scenarios}, first, we compare the air temperature in room~I
-with its reference (${w_{I}=25.5^{\circ}}$C), and then the
+with its reference (${w_{I}=\input{reference}^{\circ}}$C), and then the
 decision variable ${E_{I}(k)}$ with the threshold $\epsilon_{P}$.
-Observe that in the nominal behavior (in magenta), the reference $w_{I}$ is not reached
-due to power constraints the systems are under.
-
+Observe that the reference $w_{I}$ is not reached in the nominal behavior (in magenta),
+due to power constraints by which the systems are influenced.
+As expected, the decision variable lies under the threshold ${\epsilon_{P}=10^{-4}}$ with values of order ${E_{I}^{N}(k)\approx10^{-10}}$.
 \begin{figure}[h]
   \centering
- \includegraphics[width=\columnwidth]{airtemp_roomI/__ErrorWX_command_normErrH.pdf}
-  \caption{Air temperature in room I and the decision variable $E_{I}[k]$ for different scenarios: nominal (N), selfish behavior without correction (S),
+ \includegraphics[width=.90\columnwidth,trim=0 .3cm 0 .2cm,clip]{airtemp_roomI/__ErrorWX_command_normErrH.pdf}
+  \caption{Air temperature in room I and the decision variable $E_{I}[k]$ for three scenarios: nominal (N), selfish behavior (S),
 and selfish behavior with correction.}\label{fig:response3Scenarios}
 \end{figure}
-
-As expected, the decision variable lies under the threshold $\epsilon_{P}=10^{-4}$ with values of order ${E_{I}^{N}(k)\approx10^{-10}}$.
 
@@
 
-When the correction is activated in the system, we see that the temperatures approach their nominal value $y_{I}^{N}$.
-We can also illustrating the good
-performances of our proposition by comparing the inputs in Fig.~\ref{fig:control_3Scenarios}.
-When room~I is selfish, its control increases while the other decreases, unless the correction is activated and it approaches the nominal.
+When the correction is activated in the system, the temperatures approach their nominal value $y_{I}^{N}$.
+We can also illustrate the good performances of our proposition by comparing the inputs in Fig.~\ref{fig:control_3Scenarios}.
+When room~I is selfish, its control increases while other rooms' decrease. When the correction is activated, it approaches the nominal.
 
-We can also evaluate the performance of the proposed mechanism by comparing the global objective function calculated using the total period of simulation (${N=50}$) for the 3 scenarios (Table~\ref{tab:costsGlobalLocal}).
-When agent~I is selfish, we see the increase in the objective, degrading the overall performance.
-When the correction mechanism is activated, the difference between costs is minimal, $|\frac{J^{C}-J^{N}}{J^{N}}|\leq10^{-8}$.
+We can also evaluate the performance of the proposed mechanism by comparing the objective functions calculated using the period of simulation ${N=50}$ for the three scenarios (Table~\ref{tab:costsGlobalLocal}).
+When agent~I is selfish, we see the decrease in its objective ($\approx\!-40\%$), degrading the overall performance ($\approx\!+10\%$).
+When the correction mechanism is activated, the absolute percentual error is ${|\frac{J^{C}-J^{N}}{J^{N}}|\leq10^{-8}}$.
 %%%%%
@@
   \centering
-  \caption{Comparison of costs $J_{i}$}\label{tab:costsGlobalLocal}
-  \begin{tabular}[t]{cccc} \toprule
-    Agent  & Nominal & Selfish & Selfish + Correction\\
+  \caption{Objective functions $J_{i}$ (\% error)}\label{tab:costsGlobalLocal}
+  \begin{tabular}[t]{cccc}
+    \toprule
+    Agent  & Nominal & Selfish & + Correction\\
     \midrule
-    % I & -11468 &  -9074 & -11468 \\
-    % II & -6536 &  -7151 & -6536 \\
-    % III & -3192 &  -4052 & -3192 \\
-    % IV & -2849 &  -3694 & -2849 \\
-    % Global & -24044 &  -23971 & -24044 \\
-    Global & -24044.180111 &  -23971.164122 & -24044.180027\\
+    \input{table_costs_all_rooms_error.tex}\\
+    % \input{table_costs_only_global.tex}\\
     \bottomrule
@@
 \end{table}
-% \enlargethispage{-.8cm}
 
@@
 In this paper, we proposed an algorithm for monitoring and correcting exchanges between agents in a resource-sharing framework.
-The first phase of the algorithm consists of identifying the attacker by exploiting the exchange structure.
+The first phase of the algorithm identifies the attacker by exploiting the exchange structure.
 After this identification, if necessary, it is possible to reconstruct the original mechanism and recover nominal optimality by inverting the attack.
-This principle should be generalized to cases when the attack is not completely invertible, reconstructing by parts the original mechanism.
-Also, other types of decomposition structures and attack models need to be explored, and this is what we plan to do in the near future.
+This principle should be generalized to cases when the attack is not entirely invertible, reconstructing by parts the original mechanism.
+Also, other decomposition structures and attack models need to be explored, which we plan to do shortly.
 
@@
   \centering
- \includegraphics[width=\columnwidth]{airtemp_roomI/control.pdf}
+ \includegraphics[width=.90\columnwidth,trim=0 .1cm 0 .3cm,clip]{airtemp_roomI/control.pdf}
   \caption{Control applied in all rooms for the 3 scenarios.}\label{fig:control_3Scenarios}

### 2022-02-14 code/decompositionNagents.m

@@ -10,6 +10,7 @@ curly = @(x, varargin) x{varargin{:}}; %
 
 %= Optimization settings
 options = optimset('Display', 'off');
+warning off
 % options = [];
 rand('seed',2);
 
@@ -19,8 +20,8 @@ Te=.25; %= Sampling
 Np=4;   %= Prediction horizon
 
 simK = 50;    %= Simulation horizon
-negotP = 100; %= max # of iteration for each negotiation
-err_theta=1e-2; %= err to test theta convergence in each negotiation
+negotP = 200; %= max # of iteration for each negotiation
+err_theta=1e-3; %= err to test theta convergence in each negotiation
 
 % TODO(accacio): change name of these variables
 chSetpoint_list = 0; %= Change setpoint?
@@ -82,6 +83,7 @@ Cmat_fun=@(sys, n) ...
 %= H and f, such $\frac{1}{2}\vec{U}^TH\vec{U}+f^T\vec{U}$
 H_fun=@(Cmat,Q,R) round(Cmat'*Q*Cmat+R*eye(size(Q)),10);
 f_fun=@(Cmat,Mmat,Q,xt,Wt) Cmat'*Q*(Mmat*xt-Wt);
+c_fun=@(Mmat,Q,Xt,Wt) Xt'*Mmat'*Q*Mmat*Xt-2*Wt'*Q*Mmat*Xt+Wt'*Q*Wt;
 
 %= Gains Q and R for $\sum_{j=1}^n \|v\|^2_{Q}+\|u\|^2_{R}$
 for i=M:-1:1 % make it backward to "preallocate"
@@ -112,10 +114,10 @@ X0(:,1,3) = [15 3.1]';
 X0(:,1,4) = [17. 5.7]';
 
 %= Setpoint
-Wt = [X0(1,1)*1.5;
-      X0(1,2)*1.05;
-      X0(1,3)*1.05;
-      X0(1,4)*1.05;
+Wt = [X0(1,1)*1.50;
+      X0(1,2)*1.20;
+      X0(1,3)*1.20;
+      X0(1,4)*1.20;
      ];
 Wt_final = [Wt(1)*1.05;
             Wt(2)*1.05;
@@ -183,8 +185,9 @@ for k=1:simK
 
     %= Get value of f[k]
     for i=M:-1:1
[…]

### 2022-02-14 code/generate_figures.py

@@ -54,47 +54,75 @@ corrected_err=corrected['norm_err']
 corrected_u=corrected['uHist'][0]
 
 # ktotal=mat['ktotal'][0][0].astype(int)
-nominal_J=nominal['J']
+nominal_J=2*nominal['J']+nominal['cHist']
 nominal_sumJ=np.sum(nominal_J,axis=0)
 
-selfish_J=selfish['J']
+selfish_J=2*selfish['J']+selfish['cHist']
 selfish_sumJ=np.sum(selfish_J,axis=0)
 
-corrected_J=corrected['J']
+corrected_J=2*corrected['J']+corrected['cHist']
 corrected_sumJ=np.sum(corrected_J,axis=0)
 
-# print(nominal_sumJ)
-nominalI=nominal_sumJ[0]
-nominalII=nominal_sumJ[1]
-nominalIII=nominal_sumJ[2]
-nominalIV=nominal_sumJ[3]
+nominal_I=nominal_sumJ[0]
+nominal_II=nominal_sumJ[1]
+nominal_III=nominal_sumJ[2]
+nominal_IV=nominal_sumJ[3]
 nominal_global=np.sum(nominal_sumJ)
 
-selfishI=selfish_sumJ[0]
-selfishII=selfish_sumJ[1]
-selfishIII=selfish_sumJ[2]
-selfishIV=selfish_sumJ[3]
+selfish_I=selfish_sumJ[0]
+selfish_II=selfish_sumJ[1]
+selfish_III=selfish_sumJ[2]
+selfish_IV=selfish_sumJ[3]
 selfish_global=np.sum(selfish_sumJ)
 
 
-correctedI=corrected_sumJ[0]
-correctedII=corrected_sumJ[1]
-correctedIII=corrected_sumJ[2]
-correctedIV=corrected_sumJ[3]
+corrected_I=corrected_sumJ[0]
+corrected_II=corrected_sumJ[1]
+corrected_III=corrected_sumJ[2]
+corrected_IV=corrected_sumJ[3]
 corrected_global=np.sum(corrected_sumJ)
+with open('../article/reference.tex','w') as f:
+    print(nominal_Wt[0][0],file=f)
+with open('../article/sampling.tex','w') as f:
+    print(nominal['Te'][0][0],file=f)
+with open('../article/simulation_time.tex','w') as f:
+    print(nominal['Te'][0][0]*nominal['simK'][0][0],file=f)
[…]

### 2022-02-15 article/article.tex
change wrong superscript and submit

@@
 The specific construction depends on the active constraints in~\eqref{eq:constraints}.
-For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
+For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{\text{ac}}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
 Futhermore, if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.

### 2022-02-16 code/calculate_responsibilities.m
correct grammar error

@@ -1,5 +1,5 @@
-function Responsabilities = calculate_responsabilities(x, y, Phi, Sigma, Pi)
-    %% modes x N responsability matrix
+function Responsibilities = calculate_responsibilities(x, y, Phi, Sigma, Pi)
+    %% modes x N responsibility matrix
     modes=size(Phi,1);
     [n, N]=size(x);
 
@@ -24,6 +24,6 @@ function Responsabilities = calculate_responsabilities(x, y, Phi, Sigma, Pi)
         respNum(:,i)=sparse(Pi(1,i)*(1/(2*pi)^(n/2))*(1/sqrt(det(Sigma(:,:,i))))*exp(-1/2*max(err_diag'*SigmaInvMat*err_diag))');
     end
     normalizing_constant=sum(respNum,2);
-    Responsabilities =(respNum./normalizing_constant)';
+    Responsibilities =(respNum./normalizing_constant)';
 
 end

### 2022-02-16 code/decompositionNagents.m

@@ -229,12 +229,12 @@ for k=1:simK
         Phi_init=20*rand(modes,n^2+n);
 
         %= Estimate normal behavior
-        % [Phi,Responsabilities,~, ~] = emgm_Nestimate (X,Y,[],modes,emMaxIter,maxErr);
-        [Phi,Responsabilities,~, ~] = emgm_estimate (X,Y,Phi_init,modes,emMaxIter,maxErr);
+        % [Phi,Responsibilities,~, ~] = emgm_Nestimate (X,Y,[],modes,emMaxIter,maxErr);
+        [Phi,Responsibilities,~, ~] = emgm_estimate (X,Y,Phi_init,modes,emMaxIter,maxErr);
         % NOTE(accacio): only if values have zero
         index_of_zero=find(sum(em_theta==zeros(size(em_theta)))==n);
         % index_of_zero=1;
-        [~, z_hat_zero]=max(Responsabilities(:,index_of_zero)); %#ok
+        [~, z_hat_zero]=max(Responsibilities(:,index_of_zero)); %#ok
         zero_params=Phi(z_hat_zero,:);
         H_est(:,:,k,i)=reshape(zero_params(1:n^2),n,n)';
         f_est(:,k,i)=zero_params(n^2+1:end)';

### 2022-02-16 code/emgm_estimate.m

@@ -1,4 +1,4 @@
-function [Phi,Responsabilities,pi_new,Sigma] = emgm_estimate(X,Y,phi_init,modes,emMaxIter,maxErr)
+function [Phi,Responsibilities,pi_new,Sigma] = emgm_estimate(X,Y,phi_init,modes,emMaxIter,maxErr)
 % EMGM_NESTIMATE - ESTIMATE N DIMENSIONAL
 
     Pi=repmat(1/modes,1,modes);
@@ -35,11 +35,11 @@ function [Phi,Responsabilities,pi_new,Sigma] = emgm_estimate(X,Y,phi_init,modes,
 
     for emInd=1:emMaxIter
 
-        Responsabilities=calculate_responsabilities(X,Y,Phi,Sigma,Pi);
+        Responsibilities=calculate_responsibilities(X,Y,Phi,Sigma,Pi);
 
-        [Phi, pi_new] = update_parameters(X, Y, Responsabilities);
+        [Phi, pi_new] = update_parameters(X, Y, Responsibilities);
 
-        [~,z_hat]=max(Responsabilities,[],1);
+        [~,z_hat]=max(Responsibilities,[],1);
         for i=1:modes
             z_i=find(z_hat==i);
             clusterSize(i)=size(z_i,2);

### 2022-02-16 code/update_parameters.m

@@ -1,6 +1,6 @@
-function [Phi, pi_new] = update_parameters(Theta, Lambda, Responsabilities)
+function [Phi, pi_new] = update_parameters(Theta, Lambda, Responsibilities)
 % UPDATE_PARAMETERS -
-    modes=size(Responsabilities,1);
+    modes=size(Responsibilities,1);
     [c O]=size(Theta);
 
     Upsilon=kron(ones(c,1),eye(c));
@@ -9,10 +9,10 @@ function [Phi, pi_new] = update_parameters(Theta, Lambda, Responsabilities)
     Y=kron(G,ones(c,1));
     Omega=sparse([(Upsilon*Theta*Delta).*Y; G]');
 
-    pi_new(:)=sum(Responsabilities,2)/O;
+    pi_new(:)=sum(Responsibilities,2)/O;
     for i=1:modes
-        responsabilities=Responsabilities(i,:);
-        resp2=cellfun(@(x) x*eye(c),mat2cell(responsabilities',ones(1,O)),'UniformOutput',0);
+        responsibilities=Responsibilities(i,:);
+        resp2=cellfun(@(x) x*eye(c),mat2cell(responsibilities',ones(1,O)),'UniformOutput',0);
         Gamma=sqrt(sparse(blkdiag(resp2{:})));
 
         Phi(i,:)=-((Gamma*Omega)\(Gamma*Lambda(:)));

### 2022-02-22 article/article.tex
correct arrows in figures

@@
 
-          \draw[latex-,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
-          \draw[-latex,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};
+          \draw[-latex,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
+          \draw[latex-,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};
 
-          \draw[latex-,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
-          \draw[-latex,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
+          \draw[-latex,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
+          \draw[latex-,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
 
@@
 
-          \draw[latex-,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
-          \draw[-latex,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};
+          \draw[-latex,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
+          \draw[latex-,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};
 
@@
 
-          \draw[latex-,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
-          \draw[-latex,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
+          \draw[-latex,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
+          \draw[latex-,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};
 

### 2022-05-09 article/article.tex
use \in instead of :

@@
 \end{equation}
-with $\vec{x}_{i}[k]:\R^{n_{x}}$, $\vec{u}_{i}[k]:\R^{n_{u}}$ and $\vec{y}_{i}[k]:\R^{n_{y}}$.
+with ${\vec{x}_{i}[k]\in R^{n_{x}}}$, ${\vec{u}_{i}[k]\in R^{n_{u}}}$ and ${\vec{y}_{i}[k]\in R^{n_{y}}}$.
 
@@
 \end{equation}
-with ${\umax:\R^{n_{u}}}$.
-The agents are coupled by global input constraints with weighting matrices ${\Gamma_{i}:\semidefpos^{n_{u}}}$:
+with ${\umax\in R^{n_{u}}}$.
+The agents are coupled by global input constraints with weighting matrices ${\Gamma_{i}\in\semidefpos^{n_{u}}}$:
 \begin{equation}
@@
 \end{equation}
-where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
+where ${Q_{i}\in\semidefpos}$, ${R_{i}\in\defpos}$, and $\vik$ is a control objective.
 $\optglobobj$ denotes the optimal value of the objective function for the optimal control sequences ${\optuiseq}$.
@@
 The values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
-The approach depends only on the fact that ${H_{i}:\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
+The approach depends only on the fact that ${H_{i}\in\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
 
@@
 
- As each constraint can be active or inactive, for a given group of $t$ constraints, we can have at most $N=2^{t}$ different combinations of active sets.
+ As each constraint can be active or inactive, for a given group of $t$ constraints, we can have at most ${N=2^{t}}$ different combinations of active sets.
 
@@
 we associate to each couple $(    \randomvec{\lambda}_{o}, \randomvec{\theta}_{o})$ a latent unobserved random variable ${\random{z}_{o}}$ that indicates in which zone in $\set{Z}$ the observable variables were obtained.
-These variables are organized in ${\random{Z}:\R^{1\times O}}$ which is our set of latent variables $\set{U}$.
+These variables are organized in ${\random{Z}\in\R^{1\times O}}$ which is our set of latent variables $\set{U}$.
 

### 2022-05-09 article/article.tex
function to colon operation

@@
 $\vectorize{A}$ vectorizes matrix $A$.
-${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ means ${i=1}$.
+${n\until i\until j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\until j}$ means ${i=1}$.
 % $[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
@@
 \subsection{Model Predictive Control}\label{ssec:MPC}
-Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\mathbin{:}\nsubsystems\}}$, modeled by
+Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\until \nsubsystems\}}$, modeled by
 \begin{equation}
@@
 The overall system is controlled by a \mpc{}, which
-computes the optimal input for a finite prediction horizon ${\set{H}=\{1\mathbin{:}\predhorz\}}$ by solving the following problem:
+computes the optimal input for a finite prediction horizon ${\set{H}=\{1\until \predhorz\}}$ by solving the following problem:
 \begin{equation}\label{eq:GOP}
@@
 \end{equation}
-with $n\in\{1\mathbin{:}N\}$.
+with $n\in\{1\until N\}$.
 
@@
 \end{equation}
-with $n\in\{1\mathbin{:}N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
+with $n\in\{1\until N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
 Observe that as the function is applied only in $\lambdaik$, it does not affect the zones' hyperplanes.
@@
 
-For a group of $O$ exchanges between an agent and the coordinator, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$, ${o\in\set{O}=\{1\mathbin{:}O\}}$, which can be organized in ${\random{\Theta},\random{\Lambda}\in\R^{c\times O}}$.
+For a group of $O$ exchanges between an agent and the coordinator, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$, ${o\in\set{O}=\{1\until O\}}$, which can be organized in ${\random{\Theta},\random{\Lambda}\in\R^{c\times O}}$.
 ${(\random{\Theta},\random{\Lambda})}$ is our set of observable data $\set{B}$.
@@
 \end{equation}
-with  ${z\in\set{Z}=\{1\mathbin{:}Z\}}$.
+with  ${z\in\set{Z}=\{1\until Z\}}$.
 

### 2022-05-09 article/article.tex
v needs to be linear at least

@@
 ${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
-The values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
-The approach depends only on the fact that ${H_{i}\in\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
+The values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$.
+The approach presented in this paper works for linear control objectives such as ${\vec{v}_{i}[k]=\vec{y}_{i}[k]}$ and ${\vec{v}_{i}[k]=\vec{y}_{i}[k]-\vec{w}_{i}[k]}$.
+The approach depends only on the fact that ${H_{i}\in\semidefpos}$ does not vary w.r.t. $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.
 

### 2022-05-09 code/generate_figures.py
print cheating matrix to latex file

@@ -13,6 +13,20 @@ import matplotlib.font_manager as font_manager
 from matplotlib import rc
 from matplotlib import rcParams
 
+# shamelessly copied from https://stackoverflow.com/a/17131750/9781176 and modified
+def smallmatrix(a):
+    """Returns a LaTeX smallmatrix
+
+    :a: numpy array
+    :returns: LaTeX bmatrix as a string
+    """
+    if len(a.shape) > 2:
+        raise ValueError('bmatrix can at most display two dimensions')
+    lines = str(a).replace('[', '').replace(']', '').splitlines()
+    rv = [r'\begin{smallmatrix}']
+    rv += ['  ' + ' & '.join(l.split()) + r'\\' for l in lines]
+    rv +=  [r'\end{smallmatrix}']
+    return '\n'.join(rv)
 
 # plt.rcParams['text.usetex'] = True
 # plt.rcParams['font.family'] = 'serif'
@@ -52,6 +66,9 @@ corrected_Wt=corrected['Wt']
 corrected_xt=corrected['xt']
 corrected_err=corrected['norm_err']
 corrected_u=corrected['uHist'][0]
+
+with open('../article/triche_matrix.tex', 'w') as f:
+    print(smallmatrix(corrected['T'][:,:,0]),file=f)
 
 # ktotal=mat['ktotal'][0][0].astype(int)
 nominal_J=2*nominal['J']+nominal['cHist']

### 2022-05-10 article/article.tex
add cheating matrix value
correct some reviewer's remarks

@@
 % $\determinant{A}$ is the determinant of $A$.
-A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
+A vector $\vec{v}_{i}$, corresponds to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
 % $\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
@@
 
-The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$.
+The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{n}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$.
 The specific construction depends on the active constraints in~\eqref{eq:constraints}.
@@
 \begin{remark}\label{rmk:P_constant}
-The $\vec{s}_{i}^{j}[k]$ depend on time $k$, while $P_{i}^{j}$ do not.
+The $\vec{s}_{i}^{n}[k]$ depend on time $k$, while $P_{i}^{n}$ do not.
 \end{remark}
@@
 \begin{challenge}\label{ch:zone_unknown}
-  It is important to note that since $\vec{v}_{i}$ is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
+  It is important to note that since $\vec{v}_{i}[k]$ is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
 \end{challenge}
@@
 
-We fix the index for these zones where all constraints are active to $1$, now referenced to as the $1$-zones.
+We fix the index for the zones where all constraints are active to $1$, now referenced to as zones $1$ or the $1$-zones.
 
@@
 \begin{assumption}\label{ass:Pnominal}
-  The nominal value of $P_{i}^{j}$ for this given zone $j$, denoted $\bar{P}_{i}^{j}$, is available from reliable attack-free historical data.
+  The nominal value of $P_{i}^{j}$ for this given $j$-zone, denoted $\bar{P}_{i}^{j}$, is available from reliable attack-free historical data.
 \end{assumption}
 
-Since $P_{i}^{n}$ do not change w.r.t.\ time, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{j}}[k]-\bar{P}_{i}^{j}\|_{F}}$.
-Let ${d_{i}}$ be an indicator
+Since $P_{i}^{j}$ do not change w.r.t.\ time, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{j}}[k]-\bar{P}_{i}^{j}\|_{F}}$.
+Let ${\mathfrak{D}_{i}}$ be an indicator
 \begin{equation}
   \label{eq:2}
-  d_{i}=\indicator{E_{i}[k]\geq\epsilon_{P}},
+  \mathfrak{D}_{i}=\indicator{E_{i}[k]\geq\epsilon_{P}},
 \end{equation}
@@
 \begin{assumption}
-  We suppose a $1$-zone exists for each agent $i$ for every time $k$.
+  For each agent $i$, at every time $k$, there is a corresponding $1$-zone.
 \end{assumption}
@@
   \Repeat{$\set{P}_{\mathrm{cur}}$ converges}{
-    $\set{P}_{\mathrm{cur}}=\set{P}_{\mathrm{new}}$\;
+    $\set{P}_{\mathrm{cur}}\gets\set{P}_{\mathrm{new}}$\;
     \Estep{
@@
 with  ${z\in\set{Z}=\{1\until Z\}}$.
+\remark{Observe that the indices $z$, do not necessarily correspond to the original indices $n$.}
 
@@
     Coordinator estimates $\widehat{\tilde{P}_{i}^{1}}[k]$ and $\widehat{\tilde{\vec{s}}_{i}^{1}}[k]$ with \EM{}\;
-    Coordinator computes $d_{i}$ using~\eqref{eq:2}\;
+    Coordinator computes $\mathfrak{D}_{i}$ using~\eqref{eq:2}\;
   }
@@
   Apply Algorithm~\ref{alg:quantityAlg} with adequate versions of $\vec{\lambda}_{i}\p$:\;
-  \quad$\tilde{\vec{\lambda}}_{i}\p$, if $d_{i}=0$ and ${\vec{\lambda}_{i}}_{\mathrm{rec}}$, if ${d_{i}=1}$~\eqref{eq:lambda_reconstruction}\;
+  \quad$\tilde{\vec{\lambda}}_{i}\p$, if $\mathfrak{D}_{i}=0$ and ${\vec{\lambda}_{i}}_{\mathrm{rec}}$, if ${\mathfrak{D}_{i}=1}$~\eqref{eq:lambda_reconstruction}\;
  }
@@
 where ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$. ${x_A}_i$ and ${x_W}_i$ are the mean temperatures of the air and walls inside room~$i$. $\vec{u}_{i}$ is the input (the heating power)
-for the corresponding room. The inputs are constraint by ${\sum_{i=1}^{4}\vec{u}_{i}(t)\preceq 4\mathrm{kW}}$.
+for the corresponding room. The inputs are constrained by ${\sum_{i=1}^{4}\vec{u}_{i}(t)\preceq 4\mathrm{kW}}$.
 
@@
 
+  For scenarios (2) and (3), Agent I uses
+  $$T_{I}=\left[\input{triche_matrix}\right].$$
+
 In Fig.~\ref{fig:response3Scenarios}, first, we compare the air temperature in room~I

### 2022-05-16 article/article.tex
correct captions and vectors

@@
 }
- \caption{Quantity decomposition-based \acrshort{dmpc}.}\label{alg:quantityAlg}
+ \caption{Quantity decomposition negotiation.}\label{alg:quantityAlg}
 \end{algorithm2e}
@@
     % \begin{cases}
-      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
+      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq \vec{b}^{n}[k],
     % \end{cases}
@@
 
-The halfspaces defined by the pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.
+The halfspaces defined by the pairs $(G^{n}[k],\vec{b}^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.
 
@@
   \caption{Air temperature in room I and the decision variable $E_{I}[k]$ for three scenarios: nominal (N), selfish behavior (S),
-and selfish behavior with correction.}\label{fig:response3Scenarios}
+and selfish behavior with correction (C).}\label{fig:response3Scenarios}
 \end{figure}

### 2022-05-16 code/generate_figures.py
change fonts for poster figures

@@ -31,8 +31,9 @@ def smallmatrix(a):
 # plt.rcParams['text.usetex'] = True
 # plt.rcParams['font.family'] = 'serif'
 # plt.rcParams['font.sans-serif'] = 'cm'
-
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+serif_font = {'family': 'serif', 'serif': ['Computer Modern']};
+arial_font = {'family': 'sans-serif'};
+rc('font', **serif_font)
 rc('text', usetex=True)
 import tikzplotlib
 
@@ -205,8 +206,15 @@ axs[1].set_xlim([1, simK])
 axs[1].set_xlabel('Time (k)',usetex=True,fontsize=16)
 
 fig.tight_layout()
-plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH" +  ".pdf",bbox_inches='tight')
-plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH" +  ".png",bbox_inches='tight')
+rc('font', **serif_font)
+plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH" +  ".pdf",bbox_inches='tight',transparent=True)
+plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH" +  ".png",bbox_inches='tight',transparent=True)
+
+# plt.rcParams.update({'font.size': 50})
+# fig.tight_layout()
+rc('font', **arial_font)
+plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH_poster" +  ".pdf",bbox_inches='tight',transparent=True)
+plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH_poster" +  ".png",bbox_inches='tight',transparent=True)
 
 # NOTE(accacio): control
 fig, axs = plt.subplots(3, 1)
@@ -236,8 +244,12 @@ axs[2].legend(( 'I', 'II','III','IV'),loc='upper right',ncol=4,fontsize=13)
 axs[2].set_xlabel('Time (k)',usetex=True,fontsize=16)
 
 fig.tight_layout()
-plt.savefig("../img/airtemp_roomI" + "/control" +  ".pdf",bbox_inches='tight')
-plt.savefig("../img/airtemp_roomI" + "/control" +  ".png",bbox_inches='tight')
+rc('font', **serif_font)
[…]

### 2022-05-17 article/article.tex
remove white box and capitalize title

@@
 \title{\LARGE \bf
-  Expectation-Maximization based defense mechanism for distributed Model Predictive Control
+  Expectation-Maximization Based Defense Mechanism for Distributed Model Predictive Control
 }
@@
           fill=none,
-          draw=white,
+          draw=none,
           below=of block1,

### 2022-06-23 code/generate_figures.py
add image and update images

@@ -135,7 +135,30 @@ print("III &", round(nominal_III,2),"(",nominal_III/nominal_III, ")& ",round(sel
 print("IV &", round(nominal_IV,2),"(",nominal_IV/nominal_IV, ")& ",round(selfish_IV,2),"(", round(selfish_IV/nominal_IV,2),")&",(corrected_IV)," (",round(corrected_IV/nominal_IV,2) ,")\\\\")
 print("Global &", round(nominal_global,2),"(",nominal_global/nominal_global, ")& ",round(selfish_global,2),"(", round(selfish_global/nominal_global,2),")&",(corrected_global)," (",round(corrected_global/nominal_global,2) ,")\\\\")
 
+fig, axs = plt.subplots(1, 1,figsize=(7, 3),facecolor=(.0, .0, .0, .0))
 
+nominal_bar=(round(nominal_I,2),round(nominal_II,2),round(nominal_III,2),round(nominal_IV,2),round(nominal_global,2))
+selfish_bar=(round(selfish_I,2),round(selfish_II,2),round(selfish_III,2),round(selfish_IV,2),round(selfish_global,2))
+corrected_bar=(round(corrected_I,2),round(corrected_II,2),round(corrected_III,2),round(corrected_IV,2),round(corrected_global,2))
+
+barHeight=0.25
+br_nominal = np.arange(5)
+br_selfish = [x - barHeight for x in br_nominal]
+br_corrected = [x - barHeight for x in br_selfish]
+
+plt.barh(br_nominal, nominal_bar, height=barHeight,label='Nominal') # nominal
+plt.barh(br_selfish, selfish_bar, height=barHeight,label='Selfish') # nominal
+plt.barh(br_corrected, corrected_bar, height=barHeight,label='+ Correction') # nominal
+plt.yticks([r - barHeight for r in range(5)],
+           ['I', 'II', 'III', 'IV', 'Global'],fontsize=20)
+plt.xticks(fontsize = 20)
+
+plt.ylabel('Agents', fontweight ='bold', fontsize = 25)
+plt.xlabel('Objective function', fontweight ='bold', fontsize = 25)
+plt.legend(fontsize=20)
+
+plt.savefig("../img/barplot_results_poster" +  ".pdf",bbox_inches='tight',facecolor=fig.get_facecolor())
[…]

### 2022-06-24 code/generate_figures.py
correct tikz errors update figures

@@ -137,9 +137,10 @@ print("Global &", round(nominal_global,2),"(",nominal_global/nominal_global, ")&
 
 fig, axs = plt.subplots(1, 1,figsize=(7, 3),facecolor=(.0, .0, .0, .0))
 
-nominal_bar=(round(nominal_I,2),round(nominal_II,2),round(nominal_III,2),round(nominal_IV,2),round(nominal_global,2))
-selfish_bar=(round(selfish_I,2),round(selfish_II,2),round(selfish_III,2),round(selfish_IV,2),round(selfish_global,2))
-corrected_bar=(round(corrected_I,2),round(corrected_II,2),round(corrected_III,2),round(corrected_IV,2),round(corrected_global,2))
+nominal_bar=(round(nominal_global,2),round(nominal_IV,2),round(nominal_III,2),round(nominal_II,2),round(nominal_I,2))
+selfish_bar=(round(selfish_global,2),round(selfish_IV,2),round(selfish_III,2),round(selfish_II,2),round(selfish_I,2))
+corrected_bar=(round(corrected_global,2),round(corrected_IV,2),round(corrected_III,2),round(corrected_II,2),round(corrected_I,2))
+labels=[ 'Global','IV', 'III', 'II', 'I']
 
 barHeight=0.25
 br_nominal = np.arange(5)
@@ -150,7 +151,7 @@ plt.barh(br_nominal, nominal_bar, height=barHeight,label='Nominal') # nominal
 plt.barh(br_selfish, selfish_bar, height=barHeight,label='Selfish') # nominal
 plt.barh(br_corrected, corrected_bar, height=barHeight,label='+ Correction') # nominal
 plt.yticks([r - barHeight for r in range(5)],
-           ['I', 'II', 'III', 'IV', 'Global'],fontsize=20)
+           labels,fontsize=20)
 plt.xticks(fontsize = 20)
 
 plt.ylabel('Agents', fontweight ='bold', fontsize = 25)

### 2022-06-24 code/generate_figures.py
yet another version

@@ -238,16 +238,40 @@ plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH" +  ".png",bbo
 # fig.tight_layout()
 # rc('font', **arial_font)
 
-axs[0].set_xticks(np.arange(0,simK+1,4))
+
+fig, axs = plt.subplots(2, 1,facecolor=(.0, .0, .0, .0))
+
+axs[0].plot(np.arange(0,simK+1),np.matlib.repmat(nominal_Wt[0],simK+1,1),'-',drawstyle='steps-post')
+axs[0].plot(np.arange(0,simK+1),nominal_xt[0,0:simK+1,0],'-',color='magenta',drawstyle='steps-post')
+axs[0].plot(np.arange(0,simK+1),selfish_xt[0,0:simK+1,0],'-',drawstyle='steps-post')
+axs[0].scatter(np.arange(0,simK+1),corrected_xt[0,0:simK+1,0],s=15,color='black')
+axs[0].set_title('Air temperature in room I ($^oC$)',fontsize=20)
+axs[0].set_ylim([15, 27])
+
+
+axs[1].plot(np.arange(0,simK+1),np.matlib.repmat(nominal_Wt[1],simK+1,1),'-',drawstyle='steps-post')
+axs[1].plot(np.arange(0,simK+1),nominal_xt[0,0:simK+1,1],'-',color='magenta',drawstyle='steps-post')
+axs[1].plot(np.arange(0,simK+1),selfish_xt[0,0:simK+1,1],'-',drawstyle='steps-post')
+axs[1].scatter(np.arange(0,simK+1),corrected_xt[0,0:simK+1,1],s=15,color='black')
+axs[1].set_title('Air temperature in room II ($^oC$)',fontsize=20)
+axs[1].set_ylim([15, 27])
+
+axs[0].set_xlim([1, simK])
+axs[1].set_xlim([1, simK])
+
+axs[1].set_xlabel('Time (k)',usetex=True,fontsize=16)
+
+# axs[0].set_xticks(np.arange(0,simK+1,4))
+axs[0].set_xticks([])
 axs[1].set_xticks(np.arange(0,simK+1,4))
 axs[0].tick_params(axis='both', which='major', labelsize=20)
 axs[1].tick_params(axis='both', which='major', labelsize=20)
[…]

### 2022-06-27 code/generate_figures.py
modifications after discussion with Paul/Zhighang

@@ -239,13 +239,13 @@ plt.savefig("../img/airtemp_roomI" + "/__ErrorWX_command_normErrH" +  ".png",bbo
 # rc('font', **arial_font)
 
 
-fig, axs = plt.subplots(2, 1,facecolor=(.0, .0, .0, .0))
+fig, axs = plt.subplots(2, 1,facecolor=(.0, .0, .0, .0),figsize=(6,5))
 
 axs[0].plot(np.arange(0,simK+1),np.matlib.repmat(nominal_Wt[0],simK+1,1),'-',drawstyle='steps-post')
 axs[0].plot(np.arange(0,simK+1),nominal_xt[0,0:simK+1,0],'-',color='magenta',drawstyle='steps-post')
 axs[0].plot(np.arange(0,simK+1),selfish_xt[0,0:simK+1,0],'-',drawstyle='steps-post')
 axs[0].scatter(np.arange(0,simK+1),corrected_xt[0,0:simK+1,0],s=15,color='black')
-axs[0].set_title('Air temperature in room I ($^oC$)',fontsize=20)
+axs[0].set_title('Air temperature in house I ($^oC$)',fontsize=20)
 axs[0].set_ylim([15, 27])
 
 
@@ -253,7 +253,7 @@ axs[1].plot(np.arange(0,simK+1),np.matlib.repmat(nominal_Wt[1],simK+1,1),'-',dra
 axs[1].plot(np.arange(0,simK+1),nominal_xt[0,0:simK+1,1],'-',color='magenta',drawstyle='steps-post')
 axs[1].plot(np.arange(0,simK+1),selfish_xt[0,0:simK+1,1],'-',drawstyle='steps-post')
 axs[1].scatter(np.arange(0,simK+1),corrected_xt[0,0:simK+1,1],s=15,color='black')
-axs[1].set_title('Air temperature in room II ($^oC$)',fontsize=20)
+axs[1].set_title('Air temperature in house II ($^oC$)',fontsize=20)
 axs[1].set_ylim([15, 27])
 
 axs[0].set_xlim([1, simK])
@@ -266,9 +266,8 @@ axs[0].set_xticks([])
 axs[1].set_xticks(np.arange(0,simK+1,4))
 axs[0].tick_params(axis='both', which='major', labelsize=20)
 axs[1].tick_params(axis='both', which='major', labelsize=20)
-axs[0].set_title('Air temperature in room I ($^oC$)',fontsize=20)
 
[…]
